### 2018-03-15 doc/ms.tex
Minor revisions to the code; started writing companion paper
Added functions to calculate the surface gravity given M, R, both for GR
and in the Newtonian case; also a function to calculate the redshift

Moved the plot section in the compare method for the ObservedBurst to the
end, to make it easier to read.

Updated the __init__ method for KeplerBurst, to ensure that the set of GR
parameters are more complete. Not sure this is really required

%%
%% Beginning of file 'ms.tex'
%%
%% Copied from sample61.tex by dkg 15.3.2018
%%
%% This is the paper describing the use and design of the concord suite of software
%% for burst observation-model comparisons
%%
%% See https://bitbucket.org/minbar/concord
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
%% All of these external packages should already be present in the modern TeX 
%% distributions.  If not they can also be obtained at www.ctan.org.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style  which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.1
\documentclass{aastex61}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be envoked like this:
%%
%% \documentclass[argument]{aastex61}
%% 
%% where the arguement options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can envoke to allow other stylistic
%% actions. The available options are:
%%
%%  astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%  tighten      : Makes baselineskip slightly smaller, only works with 
%%                 the twocolumn substyle.
%%  times        : uses times font instead of the default
%%  linenumbers  : turn on lineno package.
%%  trackchanges : required to see the revision mark up and print its output
%%  longauthor   : Do not use the more compressed footnote style (default) for 
%%                 the author/collaboration/affiliations. Instead print all
%%                 affiliation information after each name. Creates a much
%%                 long author list but may be desirable for short author papers
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex61}

%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003

%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2

%\received{July 1, 2016}
%\revised{September 27, 2016}
%\accepted{\today}

%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
\submitjournal{ApJ}

%% Mark up commands to limit the number of authors on the front page.
%% Note that in AASTeX v6.1 a \collaboration call (see below) counts as
%% an author in this case.
%
%\AuthorCollaborationLimit=3
%
%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist 
%% collaboration" on the front page of this example manuscript.
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{Burst model-observation comparisons}
\shortauthors{Galloway et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

\begin{document}

\title{Thermonuclear burst model-observation comparisons with {\sc concord}}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.1 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for 
%% calling author and affilations. The change was necessary to implement 
%% autoindexing of affilations which prior was a manual process that could 
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% arguement which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.1 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus 
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple 
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.1 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

\correspondingauthor{Duncan K. Galloway}
\email{duncan.galloway@monash.edu}

\author[0000-0002-6558-5121]{Duncan K. Galloway}
\affil{School of Physics \& Astronomy \\
Monash University \\
Clayton, VIC 3800, Australia}

%\author{August Muench}
%\affiliation{American Astronomical Society \\
%2000 Florida Ave., NW, Suite 300 \\
%Washington, DC 20009-1231, USA}
%\collaboration{(AAS Journals Data Scientists collaboration)}
%
%\author{Butler Burton}
%\affiliation{National Radio Astronomy Observatory}
%\affiliation{AAS Journals Associate Editor-in-Chief}
%\nocollaboration
%
%\author{Amy Hendrickson}
%\altaffiliation{Creator of AASTeX v6.1}
%\affiliation{TeXnology Inc.}
%\collaboration{(LaTeX collaboration)}
%
%\author{Julie Steffen}
%\affiliation{AAS Director of Publishing}
%\affiliation{American Astronomical Society \\
%2000 Florida Ave., NW, Suite 300 \\
%Washington, DC 20009-1231, USA}
%
%\author{Jeff Lewandowski}
%\affiliation{IOP Senior Publisher for the AAS Journals}
%\affiliation{IOP Publishing, Washington, DC 20005}

%% Note that the \and command from previous versions of AASTeX is now
%% depreciated in this version as it is no longer necessary. AASTeX 
%% automatically takes care of all commas and "and"s between authors names.

%% AASTeX 6.1 has the new \collaboration and \nocollaboration commands to
%% provide the collaboration status of a group of authors. These commands 
%% can be used either before or after the list of corresponding authors. The
%% argument for \collaboration is the collaboration identifier. Authors are
%% encouraged to surround collaboration identifiers with ()s. The 
%% \nocollaboration command takes no argument and exists to indicate that
%% the nearby authors are not part of surrounding collaborations.

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}

% AAS Journals, the Astrophysical Journal (ApJ), the
%Astrophysical Journal Letters (ApJL), and Astronomical Journal (AJ), all
%have a 250 word limit for the abstract.  If you exceed this length the
%Editorial office will ask you to shorten it.

Thermonuclear bursts arise from unstable ignition of accumulated fuel on the surface of neutron stars in low-mass X-ray binaries. Comparison of observed bursts with numerical models is a key factor in determining the properties of the host neutron star.
%
Here we describe some commonly-used approaches for determining system parameters, including composition of the burst fuel, and introduce a new suite of software tools, {\sc concord}, intended to make comprehensive model-observation comparisons practical.

\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{editorials, notices --- 
miscellaneous --- catalogs --- surveys}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.

%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

\section{Introduction} \label{sec:intro}

%\latex\ \footnote{\url{http://www.latex-project.org/}} is a document markup
%language that is particularly well suited for the publication of
%mathematical and scientific articles \citep{lamport94}. \latex\ was written
%in 1985 by Leslie Lamport who based it on the \TeX\ typesetting language
%which itself was created by Donald E. Knuth in 1978.  In 1988 a suite of
%\latex\ macros were developed to investigate electronic submission and
%publication of AAS Journal articles \citep{1989BAAS...21..780H}.  Shortly
%afterwards, Chris Biemesdefer merged these macros and more into a \latex\
%2.08 style file called \aastex.  These early \aastex\ versions introduced
%many common commands and practices that authors take for granted today.
%Substantial revisions
%were made by Lee Brotzman and Pierre Landau when the package was updated to
%v4.0.  AASTeX v5.0, written in 1995 by Arthur Ogawa, upgraded to \latex\ 2e
%which uses the document class in lieu of a style file.  Other improvements
%to version 5 included hypertext support, landscape deluxetables and
%improved figure support to facilitate electronic submission.  
%\aastex\ v5.2 was released in 2005 and introduced additional graphics
%support plus new mark up to identifier astronomical objects, datasets and
%facilities.
%
%In 1996 Maxim Markevitch modified the AAS preprint style file, aaspp4.sty,
%to closely emulate the very tight, two column style of a typeset
%Astrophysical Journal article.  The result was emulateapj.sty.  A year
%later Alexey Vikhlinin took over development and maintenance.  In 2001 he
%converted emulateapj into a class file in \latex\ 2e and in 2003 Vikhlinin
%completely rewrote emulateapj based on the APS Journal's RevTEX class.
%
%During this time emulateapj gained growing acceptance in the astronomical
%community as it filled an author need to obtain an approximate number of
%manuscript pages prior to submission for cost and length estimates. The
%tighter typeset also had the added advantage of saving paper when printing
%out hard copies.
%
%Even though author publication charges are no longer based on print pages
%\footnote{see Section \ref{sec:pubcharge} in the Appendix for more details
%about how current article costs are calculated.} the emulateapj class file
%has proven to be extremely popular with AAS Journal authors.  An informal
%analysis of submitted \latex\ manuscripts in 2015 revealed that $\sim$65\%
%either called emulateapj or have a commented emulateapj classfile call
%indicating it was used at some stage of the manuscript construction.
%Clearly authors want to have access to a tightly typeset version of the
%article when corresponding with co-authors and for preprint submissions.
%
%When planning the next \aastex\ release the popularity of emulateapj played
%an important roll in the decision to drop the old base code and adopt and
%modify emulateapj for \aastex\ v6.+ instead.  The change brings \aastex\
%inline with what the majority of authors are already using while still
%delivering new and improved features.  \aastex\ v6.0 and v6.1 were written
%by Amy Hendrickson and released in January and October 2016, respectively.
%Some of the new features in v6.0 included:
%
%\begin{enumerate}
%\item improved citations for third party data repositories and software,
%\item easier construction of matrix figures consisting of multiple 
%encapsulated postscript (EPS) or portable document format (PDF) files,
%\item figure set mark up for large collections of similar figures,
%\item color mark up to easily enable/disable revised text highlighting,
%\item improved url support, and
%\item numerous table options such as the ability to hide columns, column 
%decimal alignment, automatic column math mode and numbering, plus splitting of
%wide tables.
%\end{enumerate}
%
%The new features in v6.1 are:
%
%\begin{enumerate}
%\item ORCID support for preprints,
%\item improved author, affiliation and collaboration mark up,
%\item reintroduced the old AASTeX v5.2 {\tt\string\received}, 
%      {\tt\string\revised}, {\tt\string\accepted}, and
%      {\tt\string\published} commands plus
%      added the new {\tt\string\submitjournal} command to document
%      which AAS Journal the manuscript was submitted to, plus
%\item new typeset style options.
%\end{enumerate}
%
%The rest of this article provides information and examples on how to create
%your own AAS Journal manuscript with v6.1.  Special emphasis is placed on
%how to use the full potential of \aastex\ v6+.  The next section describes
%the different manuscript styles available and how they differ from past
%releases.  Section \ref{sec:floats} describes how tables and figures are
%placed in a \latex\ document. Specific examples of tables, Section
%\ref{subsec:tables}, and figures, Section \ref{subsec:figures}, are also
%provided.  Section \ref{sec:displaymath} discusses how to display math and
%incorporate equations in a manuscript while Section \ref{sec:highlight}
%discuss how to use the new revision mark up.  The last section,
%\ref{sec:cite}, shows how recognize software and external data as first
%class references in the manuscript bibliography.  An appendix is included
%to show how to construct one and provide some information on how article
%charges are calculated.  Additional information is available both embedded
%in the comments of this \latex\ file and in the online documentation at
%\url{http://journals.aas.org/authors/aastex.html}.

\section{Observations and modelling} 
\label{sec:data}

\subsection{Burst observations}
\label{subsec:data}

The fundamental observables that we measure include the recurrence time $\Delta t$; the persistent flux $F_{\rm per}$, a measure of the accretion rate; and the burst bolometric flux lightcurve, $(t_i,F_{i})$. From the burst lightcurve we may integrate to estimate the burst fluence, $E_b$, which may be used to estimate the $\alpha$-value, the ratio of the burst to persistent flux: 
\begin{equation}
\alpha = \frac{\Delta t F_{\rm per} c_{\rm bol}}{E_B}
\end{equation}
where $c_{\rm bol}$ is the bolometric correction giving the inverse fraction of total persistent flux emitted in the instrumental band.

\subsection{Burst models}
\label{subsec:models}

Thermonuclear burst simulations \cite[e.g. with {\sc kepler};][]{woos04} are usually carried out in a plane-parallel grid with constant gravity, $g$. The principal input parameters include the accretion rate (per unit area), $\dot{m}$; the composition of the accreted fuel, usually quantified as $(X_0, Z_{\rm CNO})$ where $X_0$ is the mass fraction of hydrogen, and $Z_{\rm CNO}$ the mass fraction of CNO nuclei, which drive the hot-CNO cycle burning between bursts; and a parameter describing the degree of heating from below the model domain, usually labeled as ``base flux'' $Q_b$.

The model predictions include a recurrence time $\Delta t_{\rm pred}$, burst energy $E_{\rm pred}$ and lightcurve $(t_i,L_{{\rm pred},i})$ covering the extent of the burst.
%
The accretion rate $\dot{m}$ may also be converted to a persistent flux level for comparison with observations.

%The default style in \aastex\ v6.1 is a tight single column style, e.g.  10
%point font, single spaced.  The single column style is very useful for
%article with wide equations. It is also the easiest to style to work with
%since figures and tables, see Section \ref{sec:floats}, will span the
%entire page, reducing the need for address float sizing.
%
%To invoke a two column style similar to the what is produced in
%the published PDF copy use \\
%
%\noindent {\tt\string\documentclass[twocolumn]\{aastex61\}}. \\
%
%\noindent Note that in the two column style figures and tables will only
%span one column unless specifically ordered across both with the ``*'' flag,
%e.g. \\
%
%\noindent{\tt\string\begin\{figure*\}} ... {\tt\string\end\{figure*\}}, \\
%\noindent{\tt\string\begin\{table*\}} ... {\tt\string\end\{table*\}}, and \\
%\noindent{\tt\string\begin\{deluxetable*\}} ... {\tt\string\end\{deluxetable*\}}. \\
%
%\noindent This option is ignored in the onecolumn style.
%
%Some other style options are outlined in the commented sections of this 
%article.  Any combination of style options can be used.
%
%Two style options that are needed to fully use the new revision tracking
%feature, see Section \ref{sec:highlight}, are {\tt\string linenumbers} which 
%uses the lineno style file to number each article line in the left margin and 
%{\tt\string trackchanges} which controls the revision and commenting highlight
%output.
%
%There is also a new {\tt\string modern} option that uses a Daniel
%Foreman-Mackey and David Hogg design to produce stylish, single column
%output that has wider left and right margins. It is designed to have fewer
%words per line to improve reader retention. It also looks better on devices
%with smaller displays such as smart phones.

\section{Analysis} 
\label{sec:analysis}

Here we describe a few different approaches that have been used to deduce system parameters from observations of thermonuclear bursts.

\subsection{Inferring fuel composition} 
\label{subsec:fuelcomp}

\subsection{Observation-model comparisons} 
\label{subsec:lccompare}

The model-predicted quantities must then be converted to what a distant observer would see, by taking into account the effects of general relativistic time dilation and redshift.

Following \cite{lampe16}, we write the corrections as follows \cite[see also][]{keek11}. For self-consistency, we need to identify a mass and radius for the neutron star for which the Newtonian potential equals the GR potentials, i.e.
\begin{equation}
\frac{GM}{r^2} = \frac{GM_{\rm GR}}{R^2_{\rm GR}\sqrt{1-2GM_{\rm GR}/(c^2R_{\rm GR})}} = \frac{GM_{\rm GR}}{R^2_{\rm GR}}(1+z)
\end{equation}
where $1+z$ is the gravitational redshift:
\begin{equation}
1+z = \frac{1}{\sqrt{1-2GM_{\rm GR}/(c^2R_{\rm GR})}}
\end{equation}
This equality is generally achieved by assuming that $M=M_{\rm GR}$, and solving for $R_{\rm GR}$. 
We define $\xi$ such that $R_{\rm GR}=\xi R$. 
One advantage of this choice is that the mass accretion rate is identical in the Newtonian and observer frames, and also that $\xi = \sqrt{1 + z}$.

An additional correction that must be applied is the effect of the anisotropic emission of the burst, based on the system inclination $i$. This effect has been simulated for a range of disk geometries \cite[e.g.][]{he16}. Because the emission region on the surface of the star is thought to be different for the burst and persistent emission, we define separate anisotropy parameters $\xi_b$ and $\xi_p$, with the sense that
\begin{equation}
L_{b,p} = 4\pi d^2\xi_{b,p}F_{b,p}
\end{equation}
where $L_b$, $L_p$ are the burst and persistent luminosity, $F_b$, $F_p$ are the corresponding (observed) fluxes, and $d$ is the source distance. With this sense it can be seen that values of $\xi_{b,p}<1$ correspond to emission preferentially beamed towards us, so that --- without the anisotropy corrections --- the isotropic luminosity inferred from the observed flux would overestimate the actual luminosity.

Of the model input parameters, we can also infer the persistent flux level expected given the model-assumed accretion rate $\dot{m}$, as follows: % gal03d, equation 2
\begin{equation}
F_{p,{\rm inf}} = \frac{L_p}{4\pi d^2} = 
% \frac{\dot{M} Q_{\rm grav}}{1+z} \left(\frac{R_{\rm NS}}{d}\right)^2 \xi_p^{-1}
% expression used in code
\frac{\dot{m} Q_{\rm grav}}{4\pi d^2 (1+z)\xi_p}% c_{\rm bol}}
\end{equation}
where $Q_{\rm grav} = c^2z/(1+z) \approx GM_{\rm NS}/R_{\rm NS}$ is the gravitational energy release per gram. % gal03d

The predicted quantities can then be converted to values suitable for comparison to observations, as follows
\begin{eqnarray}
\Delta t_{{\rm pred},\infty} & = & (1+z)\Delta t_{\rm pred} \\
% E_{{\rm pred},\infty} & = & \xi^2E_{\rm pred}/(1+z)^2 \\ % CHECK
% add in the F_per (derived from the mdot) here; see Lampe et al. (2016), eq 8
t_{i,\infty} & = & (1+z)t_i \\
F_{{\rm pred},i,\infty} & = & \xi^2\frac{L_{{\rm pred},i}}{4\pi d^2\xi_b(1+z)^2}
\end{eqnarray}

For a given pair of observed and model lightcurves, the only parameters that affect the comparison are the source distance $d$, the anistotropy parameters $\xi_b$, $\xi_p$ (each a function of the inclination $i$), and the gravitational redshift $1+z$ (which also determines the parameter $\xi$).
%
We also introduce a ``nuisance'' parameter, $t_{\rm off}$, which is required to align the observed and predicted model lightcurve so as to minimise any residual differences. This parameter and the (time dilated) model timestamps $t_{i,\infty}$ are used to overlay the model predicted lightcurve onto the observed one.

Our approach is then to explore the parameter space of $(d, i, 1+z, t_{\rm off})$ to find the best set of parameters for which the comparison likelihood is maximised:
\begin{eqnarray}
\mathcal{L} & = & -f_F\left[\left( \frac{F_p-F_{p,\rm{inf}}}{\sigma_{p}}\right)^2 
    + \log\left(\frac{2\pi}{\sigma_p}\right)\right] \\
 & & -f_t\left[ \left( \frac{\Delta t-\Delta t_{{\rm pred},\infty}}{\sigma_t}\right)^2
    + \log\left(\frac{2\pi}{\sigma_t}\right)\right] + ...\\
\end{eqnarray}
where $f_F$...

Varying the redshift $1+z$ will allow us to obtain the best match between the model and observed lightcurve.
%this parameter cannot be varied arbitrarily, 
As each model run has been performed with a particular value of the surface gravity $g$, a particular value of $1+z$ implies in turn specific values of $M_{\rm NS}$ and $R_{\rm NS}$. 
%We take the approach of keeping $R_{\rm NS}$ fixed and allowing $M_{\rm NS}$ to vary to give the required change in $1+z$, since many equations of state have roughly constant radii over a range of masses \cite[e.g.][]{lp07}.

%\startlongtable
%\begin{deluxetable}{c|cc}
%\tablecaption{ApJ costs from 1991 to 2013\tablenotemark{a} \label{tab:table}}
%\tablehead{
%\colhead{Year} & \colhead{Subscription} & \colhead{Publication} \\
%\colhead{} & \colhead{cost} & \colhead{charges\tablenotemark{b}}\\
%\colhead{} & \colhead{(\$)} & \colhead{(\$/page)}
%}
%\colnumbers
%\startdata
%1991 & 600 & 100 \\
%1992 & 650 & 105 \\
%1993 & 550 & 103 \\
%1994 & 450 & 110 \\
%1995 & 410 & 112 \\
%1996 & 400 & 114 \\
%1997 & 525 & 115 \\
%1998 & 590 & 116 \\
%1999 & 575 & 115 \\
%2000 & 450 & 103 \\
%2001 & 490 &  90 \\
%2002 & 500 &  88 \\
%2003 & 450 &  90 \\
%2004 & 460 &  88 \\
%2005 & 440 &  79 \\
%2006 & 350 &  77 \\
%2007 & 325 &  70 \\
%2008 & 320 &  65 \\
%2009 & 190 &  68 \\
%2010 & 280 &  70 \\
%2011 & 275 &  68 \\
%2012 & 150 &  56 \\
%2013 & 140 &  55 \\
%\enddata
%\tablenotetext{a}{Adjusted for inflation}
%\tablenotetext{b}{Accounts for the change from page charges to digital quanta in April, 2011}
%\tablecomments{Note that {\tt \string \colnumbers} does not work with the 
%vertical line alignment token. If you want vertical lines in the headers you
%can not use this command at this time.}
%\end{deluxetable}

%\begin{deluxetable*}{ccCrlc}[b!]
%\tablecaption{Column math mode in an observation log \label{tab:mathmode}}
%\tablecolumns{6}
%\tablenum{2}
%\tablewidth{0pt}
%\tablehead{
%\colhead{UT start time\tablenotemark{a}} &
%\colhead{MJD start time\tablenotemark{a}} &
%\colhead{Seeing} & \colhead{Filter} & \colhead{Inst.} \\
%\colhead{(YYYY-mm-dd)} & \colhead{(d)} &
%\colhead{(arcsec)} & \colhead{} & \colhead{}
%}
%\startdata
%2012-03-26 & 56012.997 & \sim 0.\arcsec5 & H$\alpha$ & NOT \\
%2012-03-27 & 56013.944 & 1.\arcsec5 & grism & SMARTS \\
%2012-03-28 & 56014.984 & \nodata & F814M & HST \\
%2012-03-30 & 56016.978 & 1.\arcsec5\pm0.25 & B\&C & Bok \\
%\enddata
%\tablenotetext{a}{At exposure start.}
%\tablecomments{The ``C'' command column identifier in the 3 column turns on
%math mode for that specific column. One could do the same for the next
%column so that dollar signs would not be needed for H$\alpha$
%but then all the other text would also be in math mode and thus typeset
%in Latin Modern math and you will need to put it back to Roman by hand.
%Note that if you do change this column to math mode the dollar signs already
%present will not cause a problem. Table \ref{tab:mathmode} is published 
%in its entirety in the machine readable format.  A portion is
%shown here for guidance regarding its form and content.}
%\end{deluxetable*}

%%% Note that the \setcounter and \renewcommand are needed here because
%%% this example is using a mix of deluxetable and tabular.  Here the
%%% deluxetable counters are set with \tablenum but the situation is a bit
%%% more complex for tabular.  Use the first command to set the Table number
%%% to ONE LESS than it should be.  The next command will auto increment it
%%% to the desired number.
%\setcounter{table}{2}
%\begin{table}[h!]
%\renewcommand{\thetable}{\arabic{table}}
%\centering
%\caption{Decimal alignment made easy} \label{tab:decimal}
%\begin{tabular}{cD@{$\pm$}D}
%\tablewidth{0pt}
%\hline
%\hline
%Column & \multicolumn2c{Value} & \multicolumn2c{Uncertainty}\\
%\hline
%\decimals
%A & 1234     & 100.0     \\
%B &  123.4   &  10.1     \\
%C &  12.34   &   1.01    \\
%D &   1.234  &   0.101   \\
%E &    .1234 &   0.01001 \\
%F &   1.0    &    .      \\
%\hline
%\multicolumn{5}{c}{NOTE. - Two decimal aligned columns}
%\end{tabular}
%\end{table}

%%% The "ht!" tells LaTeX to put the figure "here" first, at the "top" next
%%% and to override the normal way of calculating a float position
%\begin{figure}[ht!]
%\plotone{cost.eps}
%\caption{The subscription and author publication costs from 1991 to 2013.
%The data comes from Table \ref{tab:table}.\label{fig:general}}
%\end{figure}

%\begin{equation}
%\bar v(p_2,\sigma_2)P_{-\tau}\hat a_1\hat a_2\cdots
%\hat a_nu(p_1,\sigma_1) ,
%\end{equation}

%\begin{eqnarray}
%\gamma^\mu  & = &
% \left(
%\begin{array}{cc}
%0 & \sigma^\mu_+ \\
%\sigma^\mu_- & 0
%\end{array}     \right) ,
% \gamma^5= \left(
%\begin{array}{cc}
%-1 &   0\\
%0 &   1
%\end{array}     \right)  , \\
%\sigma^\mu_{\pm}  & = &   ({\bf 1} ,\pm \sigma) , 
%\end{eqnarray}

%\begin{eqnarray}
%\hat a & = & \left(
%\begin{array}{cc}
%0 & (\hat a)_+\\
%(\hat a)_- & 0
%\end{array}\right), \nonumber \\
%(\hat a)_\pm & = & a_\mu\sigma^\mu_\pm 
%\end{eqnarray}

%% If you wish to include an acknowledgments section in your paper,
%% separate it off from the body of the text using the \acknowledgments
%% command.
\acknowledgments

To be added

%% To help institutions obtain information on the effectiveness of their 
%% telescopes the AAS Journals has created a group of keywords for telescope 
%% facilities.
%
%% Following the acknowledgments section, use the following syntax and the
%% \facility{} or \facilities{} macros to list the keywords of facilities used 
%% in the research for the paper.  Each keyword is check against the master 
%% list during copy editing.  Individual instruments can be provided in 
%% parentheses, after the keyword, but they are not verified.

\vspace{5mm}
\facilities{HST(STIS), Swift(XRT and UVOT), AAVSO, CTIO:1.3m,
CTIO:1.5m,CXO}

%% Similar to \facility{}, there is the optional \software command to allow 
%% authors a place to specify which programs were used during the creation of 
%% the manusscript. Authors should list each code and include either a
%% citation or url to the code inside ()s when available.

\software{astropy \citep{astropy13}  
%          Cloudy \citep{2013RMxAA..49..137F}, 
%          SExtractor \citep{1996A&AS..117..393B}
          }

%% Appendix material should be preceded with a single \appendix command.
%% There should be a \section command for each appendix. Mark appendix
%% subsections with the same markup you use in the main body of the paper.

%% Each Appendix (indicated with \section) will be lettered A, B, C, etc.
%% The equation counter will reset when it encounters the \appendix
%% command and will number appendix equations (A1), (A2), etc. The
%% Figure and Table counter will not reset.

%\appendix
%
%\section{Appendix information}

%% The reference list follows the main body and any appendices.
%% Use LaTeX's thebibliography environment to mark up your reference list.
%% Note \begin{thebibliography} is followed by an empty set of
%% curly braces.  If you forget this, LaTeX will generate the error
%% "Perhaps a missing \item?".
%%
%% thebibliography produces citations in the text using \bibitem-\cite
%% cross-referencing. Each reference is preceded by a
%% \bibitem command that defines in curly braces the KEY that corresponds
%% to the KEY in the \cite commands (see the first section above).
%% Make sure that you provide a unique KEY for every \bibitem or else the
%% paper will not LaTeX. The square brackets should contain
%% the citation text that LaTeX will insert in
%% place of the \cite commands.

%% We have used macros to produce journal name abbreviations.
%% \aastex provides a number of these for the more frequently-cited journals.
%% See the Author Guide for a list of them.

%% Note that the style of the \bibitem labels (in []) is slightly
%% different from previous examples.  The natbib system solves a host
%% of citation expression problems, but it is necessary to clearly
%% delimit the year from the author name used in the citation.
%% See the natbib documentation for more details and options.

\bibliography{all}
\bibliographystyle{apj}

%% This command is needed to show the entire author+affilation list when
%% the collaboration and author truncation commands are used.  It has to
%% go at the end of the manuscript.
%\allauthors

%% Include this line if you are using the \added, \replaced, \deleted
%% commands to see a summary list of all changes at the end of the article.
%\listofchanges

\end{document}

% End of file `sample61.tex'.

### 2018-03-15 anisotropy.py

@@ -5,7 +5,7 @@ import astropy.io.ascii as ascii
 from scipy.interpolate import interp1d
 import astropy.units as u
 
-def anisotropy(inclination,model='he16',test=False):
+def anisotropy(inclination,model='he16',test=False,path=None):
     '''This function returns the burst and persistent anisotropy factors
     
     Factors are defined as for Fujimoto et al. 1988, i.e. the xi_b,p such that
@@ -18,6 +18,10 @@ def anisotropy(inclination,model='he16',test=False):
     xi_b, xi_p = anisotropy(45.,test=True)'''
     
     global anisotropy_he16
+    if path:
+        data_path = path+'/'
+    else:
+        data_path = ''
 
     if test == True:
         
@@ -69,7 +73,7 @@ def anisotropy(inclination,model='he16',test=False):
     elif model == 'he16':
 
         if 'anisotropy_he16' not in globals():
-            a=ascii.read('anisotropy_he16.txt')
+            a=ascii.read(data_path+'anisotropy_he16.txt')
             v=np.stack((a['col2'],a['col3'],a['col4']),axis=1).T
             anisotropy_he16 = interp1d(a['col1'],v)
 
@@ -82,7 +86,7 @@ def anisotropy(inclination,model='he16',test=False):
         print ("** ERROR ** model ",model," not yet implemented!")
         return None, None
 
-def inclination(xi,model='he16',burst=True):
+def inclination(xi,model='he16',burst=True,path=None):
     '''This function returns the inclination given a burst or persistent 
        anisotropy factor. It is the inverse of anisotropy():
        
@@ -91,13 +95,19 @@ def inclination(xi,model='he16',burst=True):
     
     results are returned in units of degrees'''
 
+    global anisotropy_he16
+    if path:
+        data_path = path+'/'
+    else:
+        data_path = ''
+
     if model == 'fuji88':
         if burst:
             return np.arccos(1.0/xi - 0.5) * 180./np.pi * u.degree / u.rad
         else:
             return np.arccos(0.5/xi) * 180./np.pi * u.degree / u.rad
     elif model == 'he16':
-        a=ascii.read('anisotropy_he16.txt')
+        a=ascii.read(data_path+'anisotropy_he16.txt')
         if burst:
             q = 1./(a['col2']+a['col3'])
             incl_he16 = interp1d(q,a['col1'])

### 2018-03-15 burstclass.py

@@ -3,6 +3,11 @@
 # This code is a preliminary attempt to develop a burst class for MINBAR,
 # including model-observation comparisons. 
 # 
+# See also the repository on BitBucket:
+#   https://bitbucket.org/minbar/concord
+#
+# Duncan.Galloway@monash.edu, 2018
+#
 # Here are the classes and functions making up this module:
 #
 # class Lightcurve(object):
@@ -31,6 +36,31 @@ from astroquery.vizier import Vizier
 from chainconsumer import ChainConsumer
 
 from anisotropy import *
+
+# ------- --------- --------- --------- --------- --------- --------- ---------
+
+def g(M,R,Newt=False,units='cm/s^2'):
+    '''
+    This function calculates the surface gravity given a mass M and radius R,
+    using the Newtonian expression if the flag Newt is set to True
+
+    The result is returned in units of cm/s^2 by default
+    '''
+
+    if Newt:
+        return (const.G*M/R**2).to(units)
+    else:
+#        return const.G*M/(R**2*sqrt(1.-2.*const.G*M/(const.c**2*R))).to(units)
+        return opz(M,R)*g(M,R,Newt=True).to(units)
+
+# ------- --------- --------- --------- --------- --------- --------- ---------
+
+def opz(M,R):
+    '''
+    This function calculates the gravitational redshift 1+z
+    '''
+
+    return 1./sqrt(1.-2.*const.G*M/(const.c**2*R))
 
 # ------- --------- --------- --------- --------- --------- --------- ---------
 
@@ -116,7 +146,7 @@ class Lightcurve(object):
 
 # Initialise the various attributes, where present. We expect to have at
 # least one of flux or lumin (and the related uncertainty)
-# Units are handled by the parent classes ObservedBurst and ModelBurst
+# Units are handled by the parent classes ObservedBurst and KeplerBurst
 
         self.time = kwargs.get('time',None)
         self.timepixr = kwargs.get('timepixr',0.0)
@@ -154,18 +184,22 @@ class ObservedBurst(Lightcurve):
     Observed burst class. Apart from the lightcurve (which is
     defined with time, flux, and flux_err columns), the additional
     attributes set are:
+
     filename - source file name
     tdel, tdel_err - recurrence time and error (hr)
     comments - ASCII file header text
[…]

### 2018-03-15 anisotropy.py
Reverted to correct version anisotropy.py

@@ -4,27 +4,39 @@ import matplotlib.pyplot as plt
 import astropy.io.ascii as ascii
 from scipy.interpolate import interp1d
 import astropy.units as u
+import os
 
-def anisotropy(inclination,model='he16',test=False,path=None):
+CONCORD_PATH = os.environ['CONCORD_PATH']
+he16_models = ['he16_a', 'he16_b', 'he16_c', 'he16_c_short', 'he16_d']
+
+
+def load_he16(model):
+    """Reads in and returns the He & Keek (2016) model specified."""
+    model_str = model.split('he16_')[1]
+    he16_filename = 'anisotropy_{}.txt'.format(model_str)
+    he16_filepath = os.path.join(CONCORD_PATH, he16_filename)
+    a=ascii.read(he16_filepath)
+
+    return a
+
+
+
+def anisotropy(inclination, model='he16_a', test=False):
     '''This function returns the burst and persistent anisotropy factors
-    
+
     Factors are defined as for Fujimoto et al. 1988, i.e. the xi_b,p such that
     L_b,p = 4pi d^2 xi_b,p F_b,p
     This can be understood as xi_b,p<1 indicating flux that is beamed
     preferentially towards us (so that the luminosity would otherwise be
     exaggerated), and xi_b,p>1 indicating flux beamed preferentially away
-    
-    Generate a test of the model using 
+
+    Generate a test of the model using
     xi_b, xi_p = anisotropy(45.,test=True)'''
-    
+
     global anisotropy_he16
-    if path:
-        data_path = path+'/'
-    else:
-        data_path = ''
 
     if test == True:
-        
+
 # Optionally plot a figure showing the behaviour
 # want to replicate Figure 2 from fuji88
 
@@ -43,7 +55,7 @@ def anisotropy(inclination,model='he16',test=False,path=None):
             xi_b = np.zeros(len(_theta))
             for i, _th in enumerate(_theta):
                 xi_b[i], xi_p[i] = anisotropy(_th,model=m)
-  
+
             plt.plot(ct,1./xi_b,'b'+s,label=r"$\xi_b^{-1}$ ("+m+")")
             plt.plot(ct,1./xi_p,'r'+s,label=r"$\xi_p^{-1}$ ("+m+")")
             plt.plot(ct,xi_p/xi_b,'g'+s,label=r"$\xi_p/\xi_b$ ("+m+")")
@@ -59,61 +71,97 @@ def anisotropy(inclination,model='he16',test=False,path=None):
     theta = inclination
[…]

### 2018-03-16 burstclass.py
Added observe method to Lightcurve

@@ -40,7 +40,10 @@ from chainconsumer import ChainConsumer
 
 from anisotropy import *
 
+# Get the path for the concord files from the environment variable
+
 CONCORD_PATH = os.environ['CONCORD_PATH']
+
 # ------- --------- --------- --------- --------- --------- --------- ---------
 
 def g(M,R,Newt=False,units='cm/s^2'):
@@ -133,20 +136,42 @@ def modelFunc(p,obs,model, disc_model):
     return ( (model.xi/opz)**2 * fInterp(obs.time+(0.5-obs.timepixr)*obs.dt)*model.lumin.unit
             / (4.*pi*dist.to('cm')**2) / xi_b )
 
-# ------- --------- --------- --------- --------- --------- --------- ---------
-
-# The basic object, which has attributes like time (array),
-# flux/luminosity etc. and methods including plot
+# ======= ========= ========= ========= ========= ========= ========= =========
 
 class Lightcurve(object):
-    """Test class for a burst lightcurve"""
+    """
+    The fundamental burst object, which has attributes like time (array),
+    flux/luminosity etc. and methods including plot. Minimal attributes
+    are:
+
+    time, dt - array of times and time bin duration
+    timepixr - set to 0.0 (default) if time is the start of the bin, 0.5 for
+               midpoint
+    flux, flux_err - flux and error (no units assumed)
+    lumin, lumin_err - luminosity and error
+
+    Normally only one of flux or luminosity would be supplied
+
+    Example: simulated burst (giving time and luminosity)
+
+    Lightcurve.__init__(self, time=d['time']*u.s,
+        lumin=d['luminosity']*u.erg/u.s, lumin_err=d['u_luminosity']*u.erg/u.s)
+
+    Example: observed burst giving flux
+
+    Lightcurve.__init__(self, time=d['col1']*u.s, dt=d['col2']*u.s,
+        flux=d['col3']*1e-9*u.erg/u.cm**2/u.s,
+        flux_err=d['col4']*1e-9*u.erg/u.cm**2/u.s)
+    """
 
     def __init__(self, *args, **kwargs):
         """initialise the object by assigning named entities from the kwargs"""
 
 # Initialise the various attributes, where present. We expect to have at
 # least one of flux or lumin (and the related uncertainty)
-# Units are handled by the parent classes ObservedBurst and KeplerBurst
+# Units are handled by the parent classes (e.g. ObservedBurst and KeplerBurst)
+
+        self.filename = kwargs.get('filename',None)
 
[…]

### 2018-03-16 burstclass.py
Fixed value of xi to use for modelFunc

@@ -112,18 +112,24 @@ def modelFunc(p,obs,model, disc_model):
     ( distance, inclination, redshift, time offset )
     '''
 
-    dist, inclination, opz, t_off = p
+    dist, inclination, _opz, t_off = p
 
 # Use the anisotropy function to calculate the anisotropy factors given
 # the inclination
 
     xi_b, xi_p = anisotropy(inclination)
 
+# Here we calculate the value of xi (ratio of GR to Newtonian radii),
+# appropriate for the adopted value of (1+z). This is used instead of the
+# value attached to the model, because that's for a different redshift
+
+    xi = sqrt(_opz)
+
 # First interpolate the model values onto the observed grid
 # This step defines the interpolation function, using a shifted and rescaled
 # model time to account for time dilation in the NS surface frame
 
-    fInterp = interp1d(opz*(model.time-t_off),model.lumin,bounds_error=False,
+    fInterp = interp1d(_opz*(model.time-t_off),model.lumin,bounds_error=False,
                       fill_value = min(model.lumin))
 
 # Then we return the predicted model flux, interpolated onto the observed
@@ -133,7 +139,9 @@ def modelFunc(p,obs,model, disc_model):
 # convention of Fujimoto et al. 1988, i.e.
 #   L_b = 4\pi d^2 \xi_b F_b
 
-    return ( (model.xi/opz)**2 * fInterp(obs.time+(0.5-obs.timepixr)*obs.dt)*model.lumin.unit
+#    return ( (model.xi/opz)**2 
+    return ( (xi/_opz)**2 
+            * fInterp(obs.time+(0.5-obs.timepixr)*obs.dt)*model.lumin.unit
             / (4.*pi*dist.to('cm')**2) / xi_b )
 
 # ======= ========= ========= ========= ========= ========= ========= =========
@@ -217,7 +225,7 @@ class Lightcurve(object):
             print ("** ERROR ** need luminosity to simulate observation")
             return None
 
-        dist, inclination, opz, t_off = param
+        dist, inclination, _opz, t_off = param
         xi_b, xi_p = anisotropy(inclination, model=disc_model)
 
         if (obs == None):
@@ -226,11 +234,13 @@ class Lightcurve(object):
 # with some dummy values for the time bins
 
             dt = 0.25*u.s
[…]

### 2018-03-16 burstclass.py
Replaced fluxwt in likelihood with passed version from dict

@@ -453,15 +453,6 @@ class ObservedBurst(Lightcurve):
 #        Q_grav = const.G*M_NS/mburst.R_NS # approximate
         Q_grav = const.c**2*(_opz-1)/_opz
 
-# These parameters give the relative weight to the tdel and persistent
-# flux for the likelihood. Since you have many more points in the
-# lightcurve, you may want to weight these greater than one so that the 
-# MCMC code will try to match those preferentially
-
-#        tdelwt=2.5e3
-        tdelwt=1.0
-        fluxwt=1.0
-        
 # Calculate the rescaled model flux with the passed parameters
 # This is equivlent to simulating an observation of the burst with the
 # chosen parameters, so we should use instead the observe method of the
@@ -490,7 +481,7 @@ class ObservedBurst(Lightcurve):
         fper_pred = fper_pred.to(u.erg/u.cm**2/u.s)
 
         fper_sig2 = 1.0/(self.fper_err.value**2)
-        lhood_cpt = np.append(lhood_cpt, -fluxwt*( 
+        lhood_cpt = np.append(lhood_cpt, -weights['fluxwt']*( 
                (self.fper.value-fper_pred.value)**2*fper_sig2 
                +np.log(2.*pi/fper_sig2) ) )
 

### 2018-03-19 doc/ms.tex
Created function calc_mr for mass, radius from g, 1+z

@@
 \section{Introduction} \label{sec:intro}
+
+Here we explore the ways that host system  parameters affect the observed properties of thermonuclear bursts, and assess to what degree these effects can be corrected for.
+%
+In \S\ref{sec:data} we describe the typical characteristics of burst observations along with those properties that can be predicted by numerical models.
+% 
+In \S\ref{sec:analysis} we explore a number of methods to infer properties of the bursting source, and assess their accuracy.
+%
+In \S\ref{sec:simulations} we apply these methods to a range of simulated data to assess what degree of systematic errors these measurements might be subject to.
 
@@
 
+\section{Application to simulated data} 
+\label{sec:simulations}
+
+Here we assemble and analyse a number of datasets to assess the precision and accuracy of the approaches described in the previous section.
+
 %% If you wish to include an acknowledgments section in your paper,

### 2018-03-19 burstclass.py

@@ -20,7 +20,7 @@
 # def modelFunc(p,obs,model):
 # def lnprior(theta):
 # def apply_units(params,units = (u.kpc, u.degree, None, u.s)):
-# def lhoodClass(params,obs,model):
+# def lhoodClass(params, obs, model, weights, disc_model):
 # def plot_comparison(obs,models,param=None,sampler=None,ibest=None):
 
 import numpy as np
@@ -68,6 +68,31 @@ def opz(M,R):
     '''
 
     return 1./sqrt(1.-2.*const.G*M/(const.c**2*R))
+
+# ------- --------- --------- --------- --------- --------- --------- ---------
+
+def calc_mr(g,opz):
+    ''''
+    this function calculates neutron star mass and radius given a surface
+    gravity and redshift
+    '''
+
+# First some checks
+
+    if hasattr(opz,'unit'):
+        assert (opz.unit == '')
+    try:
+        test = g.to('cm / s2')
+    except ValueError:
+        print ("Incorrect units for surface gravity")
+        return -1, -1
+
+# Now calculate the mass and radius and convert to cgs units
+
+    R_NS = (const.c**2*(opz**2-1)/(2.*g*opz)).to(u.cm)
+    M_NS = (g*R_NS**2/(const.G*opz)).to(u.g)
+
+    return M_NS, R_NS
 
 # ------- --------- --------- --------- --------- --------- --------- ---------
 
@@ -431,8 +456,7 @@ class ObservedBurst(Lightcurve):
 #        M_NS = (mburst.g*mburst.R_NS**2/const.G * (-_t + sqrt(_t+1)))
 #        M_NS = (mburst.g*mburst.R_NS**2/(const.G*opz)).to(u.g)
 
-        R_NS = (const.c**2*(_opz**2-1)/(2.*mburst.g*_opz)).to(u.cm)
-        M_NS = (mburst.g*R_NS**2/(const.G*_opz)).to(u.g)
+        M_NS, R_NS = calc_mr(mburst.g,_opz)
 
 # Check here:
 
@@ -932,13 +956,13 @@ def plot_comparison(obs,models,param=None,sampler=None,ibest=None):
 # to compare
 
     n = len(obs)
-    # assert (n == 3)
+    assert (n == 3)
     #
-    # b1, b2, b3 = obs
-    # m1, m2, m3 = models
+    b1, b2, b3 = obs
+    m1, m2, m3 = models
 
-    b1, b2 = obs
-    m1, m2 = models
+#    b1, b2 = obs
[…]

### 2018-03-19 burstclass.py
Now reads kepler_grids files specified by batch,run

@@ -316,7 +316,7 @@ class ObservedBurst(Lightcurve):
         self.path = path
         self.filename = filename
 
-        d=ascii.read(path+'/'+self.filename)
+        d=ascii.read(self.path+'/'+self.filename)
 
 # Now we define a Lightcurve instance, using the columns from the file
 
@@ -626,22 +626,23 @@ class KeplerBurst(Lightcurve):
                   tdel=4.06/opz,tdel_err=0.17/opz,
                   g = 1.858e+14*u.cm/u.s**2, R_NS=11.2*u.km)
 
-    If you provide a run_id value, like so:
+    If you provide a batch and run value, like so:
+
+    c = KeplerBurst(batch=2,run=9)
+
+    or (for the older results) a run_id value, like so:
 
     c = KeplerBurst(run_id='a028',path='../../burst/reference')
 
-    most of these parameters will be populated from the table
+    the appropriate model lightcurve will be read in, and the model parameters
+    will be populated from the table
     '''
 
-    def __init__(self, filename=None, run_id=None, path=None, **kwargs):
+    def __init__(self, filename=None, run_id=None, path=None, 
+                 source='gs1826', batch=None, run=None, 
+                 **kwargs):
 
         eta = 1e-6	# tolerance level for derived parameters
-
-        if run_id != None:
-            # For a KEPLER run, we use the convention for filename as follows:
-            self.filename = 'kepler_'+run_id+'_mean.txt'
-        elif filename!= None:
-            self.filename = filename
 
 # Don't add the path to the filename, because we want to use the latter as
 # a plot label (for example)
@@ -650,9 +651,23 @@ class KeplerBurst(Lightcurve):
             path = '.'
         self.path = path
 
+# Different file specifications here
+
+        if run_id != None:
+            # For a KEPLER run, we use the convention for filename as follows:
+            self.filename = 'kepler_'+run_id+'_mean.txt'
+        elif ((batch != None) & (run != None)):
+            self.filename = source+"_{}_xrb{}_mean.data".format(batch,run)
+            self.path = self.path+"/kepler_grids/sources/{}/mean_lightcurves/{}_{}".format(source,source,batch)
+        elif filename!= None:
+            self.filename = filename
+        else:
+            print ("** ERROR ** no valid model specification")
+            return
+
 # Read in the file, and initialise the lightcurve
 
[…]

### 2018-03-19 burstclass.py
Now sets mass correctly for kepler_grids models

@@ -96,6 +96,22 @@ def calc_mr(g,opz):
 
 # ------- --------- --------- --------- --------- --------- --------- ---------
 
+def solve_radius(M,R_Newt,eta=1e-6):
+    '''
+    This routine determines the GR radius given the NS mass and Newtonian
+    radius, assuming the GR and Newtonian masses are identical
+
+    Solving is tricky so we just use an iterative approach
+    '''
+
+    R_NS = R_Newt	# trial
+    while (abs(g(M,R_NS)-g(M,R_Newt,Newt=True))/g(M,R_NS) > eta):
+        R_NS = R_Newt*sqrt(opz(M,R_NS))
+
+    return R_NS
+
+# ------- --------- --------- --------- --------- --------- --------- ---------
+
 def decode_LaTeX(string):
     '''
     This function converts a LaTeX numerical value (with error) to one
@@ -689,14 +705,24 @@ class KeplerBurst(Lightcurve):
 
             if ((batch != None) & (run != None)):
 
-# For the newer models, read data from the summary table
+# For the newer models, read data from the summary table (and the
+# companion parameter table)
 
                 self.data = ascii.read(self.path+"/../../summ_{}.txt".format(source))
+                self.param = ascii.read(self.path+"/../../params_{}.txt".format(source))
 
                 self.row = np.where(np.logical_and(self.data['batch'] == batch,self.data['run'] == run))[0]
+                self.row_p = np.where(np.logical_and(self.param['batch'] == batch,self.param['run'] == run))[0]
+
+# Set some special parameters here (others are set with the kwargs later
+# on)
 
                 self.tdel = self.data['tDel'][self.row]/3600.*u.hr
                 self.tdel_err = self.data['uTDel'][self.row]/3600.*u.hr
+
+# Set the gravity and NS mass
+
+                self.M_NS = self.param['mass'][self.row_p]*u.Msun
 
             else:
 
@@ -739,23 +765,30 @@ class KeplerBurst(Lightcurve):
                     self.tdel = self.data['tDel'][self.row]/3600.*u.hr
                     self.tdel_err = self.data['uTDel'][self.row]/3600.*u.hr
 
+# Set the parameters for the old Kepler models
+
 # Additional parameter here gives the radius conversion between the assumed
 # (Newtonian) value of 10 km and the GR equivalent for a neutron star of
[…]

### 2018-04-07 mn_burst_compare.py
Added MultiNest implementation example

@@ -1,0 +1,99 @@
+# Script to perform burst model comparisons, and run multinest
+# Duncan.Galloway@monash.edu, 2018 Mar
+#
+# Initialise the environment
+
+from burstclass import *
+import pymultinest
+
+# Read in an observed burst, from the reference sample of Galloway et al.
+# (2017)
+
+b = ObservedBurst('gs1826-24_4.177h.dat',path='../../burst/reference/')
+
+# Now also define a model burst. Lampe et al. models are deprecated;
+
+## c = ModelBurst('a005',path='../../reference')
+# c = KeplerBurst(run_id='a028',path='../../burst/reference')
+
+# Low-metallicity run supplied by Z. Johnston, 2017 April
+# accretion rate, fuel composition and recurrence time were all provided by Zac;
+# xi, g and R_NS are guesses
+
+c=KeplerBurst(filename='mean_run930.dat',lAcc=0.108,Z=0.005,H=0.7,
+              tdel=4.5/1.259,tdel_err=0.2/1.259,
+              g = 1.86552e+14*u.cm/u.s**2, R_NS=11.2*u.km)
+
+# ------- --------- --------- --------- --------- --------- --------- ---------
+
+def mnPrior(cube, ndim, nparams):
+    '''
+    Prior function for use with pymultinest. Should transform the unit 
+    cube into the parameter cube.
+    Parameter order (equivalent to dimensions of cube) is: 
+    distance (kpc), inclination (degrees), redshift, and offset time 
+    (seconds)'''
+
+# Distance is in the range 1-16 kpc
+    cube[0] = 1. + cube[0]*15.
+    
+# inclination in the range 0-90; could be more restrictive for a non-dipper
+    cube[1] = np.arccos(cube[1])*180./pi
+
+# redshift is in the range 1-2
+    cube[2] = 1. + cube[2]
+    
+# one or more time offsets are in the range -20,+20
+    for i in range(ndim-3):
+        cube[i+3] = (cube[i+3]-0.5)*40.
+
+# ------- --------- --------- --------- --------- --------- --------- ---------
+
+def Loglike(cube, ndim, nparams):
+    '''
+    Function to return the likelihood, takes as input the parameter cube
+    and should return the logarithm of the likelihood.
+    Don't know how to pass additional parameters to this function
+    (ans: seems we don't need to, we can just access them as global 
+    variables)
+    Although the cube parameter name implies multiple parameter points to
+    evaluate, practically it seems to only have one
[…]

### 2018-04-08 burstclass.py
Added write method for Lightcurve, fper function
Various features added to improve the utility for burst simulations; in
the observe method, now copies the flux_err from the observed burst
(from which we get the time bins) to the simulated observation.

Can now also write the lightcurves with some limited header information;
these can be subsequently read in with the same code, if required.

Also split off the calculation of fper (used in the compare method) into a
separate function, so we can use it in other contexts

@@ -30,6 +30,7 @@ import os
 import astropy.units as u
 import astropy.constants as const
 import astropy.io.ascii as ascii
+import csv
 from astropy.table import Table
 import matplotlib.pyplot as plt
 from matplotlib import gridspec
@@ -37,6 +38,7 @@ import re
 from scipy.interpolate import interp1d
 from astroquery.vizier import Vizier
 from chainconsumer import ChainConsumer
+from datetime import datetime
 
 from anisotropy import *
 
@@ -109,6 +111,28 @@ def solve_radius(M,R_Newt,eta=1e-6):
         R_NS = R_Newt*sqrt(opz(M,R_NS))
 
     return R_NS
+
+# ------- --------- --------- --------- --------- --------- --------- ---------
+
+def fper(mdot,opz,dist,xi_p,c_bol=1.0):
+    '''
+    Calculates the persistent flux, based on the inferred mdot, redshift
+    etc.
+    '''
+
+# First we calculate Q_grav, to calculate the inferred persistent flux that
+# we should see (based on the accretion rate); that is not given in Lampe
+# et al.  2016, but is given in gal03d
+# Note that (as for the M_NS) we're using the passed parameter opz here,
+# not the value that is associated with the model burst
+
+#        Q_grav = const.G*M_NS/mburst.R_NS # approximate
+    Q_grav = const.c**2*(opz-1)/opz
+
+# persistent flux (see lampe16, eq. 8, noting that our mdot is averaged
+# over the neutron star surface):
+
+    return ( mdot*Q_grav/ (4.*pi*opz*dist**2*xi_p*c_bol) ).to(u.erg/u.cm**2/u.s)
 
 # ------- --------- --------- --------- --------- --------- --------- ---------
 
@@ -213,6 +237,8 @@ class Lightcurve(object):
         flux_err=d['col4']*1e-9*u.erg/u.cm**2/u.s)
     """
 
+# ------- --------- --------- --------- --------- --------- --------- ---------
+
     def __init__(self, *args, **kwargs):
         """initialise the object by assigning named entities from the kwargs"""
 
@@ -231,6 +257,48 @@ class Lightcurve(object):
 
         self.lumin = kwargs.get('lumin',None)
         self.lumin_err = kwargs.get('lumin_err',None)
+
+# ------- --------- --------- --------- --------- --------- --------- ---------
+
+    def write(self,filename='lightcurve.csv',addhdr=None):
+        '''
[…]

### 2018-04-08 burstclass.py
Added noise to simulated observations

@@ -251,7 +251,8 @@ class Lightcurve(object):
         self.time = kwargs.get('time',None)
         self.timepixr = kwargs.get('timepixr',0.0)
         self.dt = kwargs.get('dt',None)
-        if kwargs.get('flux',None):
+#        if kwargs.get('flux',None):
+        if 'flux' in kwargs:
             self.flux = kwargs.get('flux',None)
             self.flux_err = kwargs.get('flux_err',None)
 
@@ -318,10 +319,17 @@ class Lightcurve(object):
             ylabel = "Luminosity ({0.unit:latex_inline})".format(self.lumin)
         
         plt.step(self.time,y,where='post',label=self.filename)
-        if (yerror & (self.dt != None) & (yerr != None)):
-            plt.errorbar(self.time.value+(0.5-self.timepixr)*self.dt.value,
-                y.value, yerr=yerr.value,fmt='b.')
+#        print (type(self.dt), type(yerr))
+#        print (yerror & (self.dt != None) & (yerr != None))
+#        if (yerror & (self.dt != None) & (yerr != None)):
+        if yerror:
+            try:
+                plt.errorbar(self.time.value+(0.5-self.timepixr)*self.dt.value,
+                    y.value, yerr=yerr.value,fmt='b.')
 #            plt.plot(self.time,y,label=self.filename)
+            except:
+                pass
+#            print ("** WARNING ** errors not present, can't plot errorbars")
 
         plt.xlabel("Time ({0.unit:latex_inline})".format(self.time))
         plt.ylabel(ylabel)
@@ -330,7 +338,10 @@ class Lightcurve(object):
 
     def observe(self, param = [6.1*u.kpc,60.*u.degree,1.26,-10.*u.s], obs=None,
         disc_model='he16_a',c_bol=1.0):
-        """Convert a luminosity profile to a simulated observation"""
+        """
+        Convert a luminosity profile to a simulated observation, with
+        plausible errors
+        """
 
         if not hasattr(self,'lumin'):
             print ("** ERROR ** need luminosity to simulate observation")
@@ -349,8 +360,19 @@ class Lightcurve(object):
             obs = Lightcurve(time=np.arange(npts)*dt+t_off, 
                              dt=np.full(npts,dt)*u.s,
                              flux=np.zeros(npts),flux_err=np.zeros(npts))
+#            print (obs.flux_err)
+
+        else:
+            npts = len(obs.time)
 
         model = modelFunc(param, obs, self, disc_model)
+
[…]

### 2018-04-09 doc/ms.tex
Expanded paper discussion; minor edits to code
Significantly expanded the analysis section of the paper, and added the
full derivation of the fuel composition calculation along with a priori
distributions of xi_b, etc.

Updated the model used for the "test" mode of anisotropy.py, this was
still at the legacy value of he16 (changed to he16_a)

Slight modification of the flux calculation in burstclass.py; original
expression had \xi and 1+z, but because the former is just sqrt(1+z) you
can simplify this a little

@@
 
+\subsection{Simulated observations}
+\label{subsec:simobs}
+
+In order to test the analysis approaches described in this paper, we generated sets of simulated data based on the {\sc kepler} models described in \S\ref{subsec:models}. 
+%
+This procedure is identical to that used to convert model predictions to data for direct comparison to observations, as described in \S\ref{subsec:lccompare}.
+%
+The process can be summarised as follows.
+\begin{enumerate}
+\item Multiply the time bins for the predicted burst lightcurve by the gravitational redshift $1+z$, thereby ``stretching'' the profile to account for the general relativistic time dilation at the neutron-star surface
+% \item Shift the predicted burst lightcurve by $t_{\rm off}$
+\item Apply the same correction to the model-predicted recurrence time $\Delta t_{\rm pred}$.
+\item Interpolate the model-predicted lightcurve onto a set of observational time bins, corresponding (for example) to the typical resolution for time-resolved spectroscopy (0.25~s)
+\item Translate the model-predicted luminosity to the corresponding quantity measured by a distant observer, by 
+% multiplying by $[\xi/(1+z)]^2$
+dividing by $1+z$
+\item Convert the luminosity to (isotropic) flux by dividing by the distance factor, $4\pi d^2$ \label{distfac}
+\item Take into account the expected anisotropy effects due to the system inclination, by dividing the luminosity by the anisotropy factor $\xi_b$ \label{anisofac}
+\item Calculate the persistent flux expected for the model-assumed accretion rate $\dot{m}$, and apply the same corrections as for the burst flux in steps \ref{distfac} and \ref{anisofac} (adopting a separate anisotropy factor $\xi_p$ appropriate for the persistent flux).
+We also divide by a bolometric correction factor $c_{\rm bol}$ accounting for the limited instrumental passband.
+\end{enumerate}
+% Below we describe each of these steps in more detail
+
+We determine the general relativistic (GR) corrections within the constraints of the numerical models, which are typically calculated assuming a Newtonian potential with gravity $g=GM/R^2$, where $M$ and $R$ are the equivalent Newtonian mass and radius.
+%
+However, we are free to vary the surface redshift $1+z$, for example to achieve improved agreement with an observed profile.
+
+The model time bins  $t_i$ and the predicted burst recurrence time $\Delta t_{\rm pred}$ are converted to values as would be measured by a distant observer as follows:
+\begin{eqnarray}
+t_{i,\infty} & = & (1+z)t_i \\
+\Delta t_{{\rm pred},\infty} & = & (1+z)\Delta t_{\rm pred}
+\end{eqnarray}
+
+Care must be taken to ensure the Newtonian model predictions can be correctly translated to include the GR corrections expected for quantities at the neutron star surface.
+%
+Following \cite{lampe16}, 
+% we write the corrections as follows \cite[see also][]{keek11}. For self-consistency, 
+we identify a mass and radius for the neutron star for which the Newtonian potential equals the GR potentials, i.e.
+\begin{equation}
+\frac{GM}{R^2} = \frac{GM_{\rm GR}}{R^2_{\rm GR}\sqrt{1-2GM_{\rm GR}/(c^2R_{\rm GR})}} = \frac{GM_{\rm GR}}{R^2_{\rm GR}}(1+z)
+\end{equation}
+where $1+z$ is the gravitational redshift:
+\begin{equation}
+1+z = \frac{1}{\sqrt{1-2GM_{\rm GR}/(c^2R_{\rm GR})}}
+\end{equation}
+This equality is generally achieved by assuming that $M=M_{\rm GR}$, and solving for $R_{\rm GR}$. 
+We define $\xi$ such that $R_{\rm GR}=\xi R$. 
+One advantage of this choice is that the mass accretion rate is identical in the Newtonian and observer frames, and also that $\xi = \sqrt{1 + z}$.
+In that case, the model-predicted luminosity is related to the luminosity measured by a distant observer, by
+\begin{eqnarray}
+L_\infty & = & \xi^2L/(1+z)^2 \nonumber \\
+& = & L/(1+z)
+\end{eqnarray}
+
+We note that the combination of surface gravity $g$ (assumed for the model) and adopted $1+z$  uniquely specifies the neutron star mass $M_{\rm NS}$ and radius $R_{\rm NS}$:
+\begin{eqnarray}
+% as used in function calc_mr in burstclass.py
+R_{\rm NS} &=& c^2\frac{(1+z)^2-1}{2g(1+z)}\\
+M_{\rm NS} &=& \frac{gR_{\rm NS}^2}{G(1+z)}
+\end{eqnarray}
+Thus, by identifying the optimal value of $1+z$ for comparison to a particular observation, we can constrain the mass and radius, at a fixed $g$.
+
+An additional correction that must be applied is the effect of the anisotropic emission of the burst, based on the system inclination $i$.
+This effect has been simulated for a range of disk geometries \cite[e.g.][]{he16}.
+Because the emission region on the surface of the star is thought to be different for the burst and persistent emission, we define separate anisotropy parameters $\xi_b$ and $\xi_p$, with the sense that
+\begin{equation}
+L_{b,p} = 4\pi d^2\xi_{b,p}F_{b,p}
+\end{equation}
+where $L_b$, $L_p$ are the burst and persistent luminosity, $F_b$, $F_p$ are the corresponding (observed) fluxes, and $d$ is the source distance.
+With this sense it can be seen that values of $\xi_{b,p}<1$ correspond to emission preferentially beamed towards us, so that --- without the anisotropy corrections --- the isotropic luminosity inferred from the observed flux would overestimate the actual luminosity.
+
+Of the model input parameters, we can also infer the persistent flux level expected given the model-assumed accretion rate $\dot{m}$, as follows: % gal03d, equation 2
+\begin{equation}
+% function fper in burstclass.py
+F_{p,\infty} = \frac{L_p}{4\pi d^2} = 
+% \frac{\dot{M} Q_{\rm grav}}{1+z} \left(\frac{R_{\rm NS}}{d}\right)^2 \xi_p^{-1}
+% expression used in code
+\frac{\dot{m} Q_{\rm grav}}{4\pi d^2 (1+z)\xi_p c_{\rm bol}}
+\end{equation}
+% Q_grav already defined in the previous section
+% where $Q_{\rm grav} = c^2z/(1+z) \approx GM_{\rm NS}/R_{\rm NS}$ is the gravitational energy release per gram. % gal03d 
+%
+The bolometric correction $c_{\rm bol}$ accounts for the expermental limitation that the persistent flux can only be measured over a limited instrumental passband.
+The bolometric correction is the ratio of the estimated bolometric flux to the band-limited value.
+
+We performed simulations based on {\sc kepler} simulations drawn from our range of in-hand simulations. The range of input values are listed in Table \ref{tab:simparams}.
+
+\begin{deluxetable}{lccc}
+\tablecaption{Parameter ranges for simulated data \label{tab:simparams}}
+\tablehead{
+\colhead{Parameter} & \colhead{Symbol} & \colhead{Units} & \colhead{Range} }
+\startdata
+Surface gravity & $g$ & $10^{14}\ {\rm g\,cm^{-2}}$ & \\
+Accreted hydrogen mass fraction & $X_0$ & \nodata & \\
+Metallicity & $Z_{\rm CNO}$ & \nodata & \\
+Base flux & $Q_b$ & MeV/nucleon & \\
+%
+% samples are drawn randomly over the ranges below, which might be tightened for the 
+% final production versions. In particular I don't know if such high redshifts are 
+% physical
+Distance & $d$ & kpc & 4--12\\
+System inclination\tablenotemark{a} & $i$ & degrees & 0--75\\
+Surface redshift & $1+z$ & \nodata & 1--2\\
+\enddata
+\tablenotetext{a}{Drawn from a random distribution uniform in $\cos i$}
+% \tablecomments{Note that {\tt \string \colnumbers} does not work with the 
+% vertical line alignment token. If you want vertical lines in the headers you
+% can not use this command at this time.}
+\end{deluxetable}
+
+%The predicted quantities can then be converted to values suitable for comparison to observations, as follows
+%\begin{equation}
+%F_{{\rm pred},i,\infty}  =  \xi^2\frac{L_{{\rm pred},i}}{4\pi d^2\xi_b(1+z)^2}
+%\end{equation}
+
 %The default style in \aastex\ v6.1 is a tight single column style, e.g.  10
@@
 
+Given a set of burst observations as described in \S\ref{subsec:data}, a common approach \cite[e.g][]{falanga11} is to estimate the fuel composition at ignition, and hence the accreted fuel composition, based on simple analytic estimates of the burst energy production $Q_{\rm nuc}$.
+
+One issue with this approach is that the effects of anisotropy are frequently neglected.
+%
+A second issue which has emerged recently is that the commonly used estimates for the nuclear energy generation rate $Q_{\rm nuc}$ are not supported by recent {\sc kepler} models. 
+%
+We describe the general approach below including the effects of these two issues.
+
+Given the measured $\alpha$-value, the nuclear energy generation rate $Q_{\rm nuc}$ can be estimated, as follows:
+\begin{eqnarray}
+\alpha & = & \frac{\Delta t F_{\rm per} c_{\rm bol}}{E_B} \\
+& = & \frac{Q_{\rm grav}}{Q_{\rm nuc}}\frac{\xi_b}{\xi_p}(1+z)
+\end{eqnarray}
+where $Q_{\rm grav} = c^2z/(1+z) \approx GM_{\rm NS}/R_{\rm NS}$ is the gravitational energy release per gram.
+Note the dependence of the observational $\alpha$-value on both the surface redshift and the anisotropy parameters.
+The nuclear energy generation rate is thus
+\begin{equation}
+Q_{\rm nuc} = \frac{c^2 z}{\alpha}\frac{\xi_b}{\xi_p} \label{qnuc}
+\end{equation}
+Thus, given an estimate of $z$ and the inclination, the burst measurements can be used to infer $Q_{\rm nuc}$.
+
+% describe plausible assumptions for ration of $\xi_b/\xi_p$.
+The system inclination for LMXBs is notoriously difficult to measure.
+For non-dipping sources, the likely range is up to $72^\circ$, while dipping sources are likely $i\gtrsim79.3^\circ$ \cite[]{gal16a}.
+% 
+% 5, 16, 50, 68, 95 th percentiles:
+% [ 0.58815149  0.72522467  1.12987208  1.49149624  1.59287602]
+% min, max
+% 0.525409532417 1.63715782279
+For an isotropically-distributed population of non-dipping sources, the ratio $\xi_b/\xi_p$ is roughly uniformly-distributed between 0.53--1.64, rising a little towards the high end.
+The median value is 1.13, while the 95\% confidence interval is 0.73--1.49.
+
+\begin{itemize}
+\item describe plausible assumptions for $z$, e.g. based on recent analysis of bursts
+\end{itemize}
+
+Given an estimate of $Q_{\rm nuc}$, the average hydrogen fraction $\langle X\rangle$ in the fuel layer at ignition has been estimated to date using the relation $Q_{\rm nuc}=1.6 + 4\langle X\rangle$~MeV/nucleon \cite[e.g.][and references therein]{gal03d}.
+This expression includes $\approx35$\% losses
+attributed to neutrino emission \cite[]{fuji87}.
+However, the 35\% value applies only to $\beta$-decays in the much more extensive {\it rp}-process burning chain, and recent {\sc kepler} simulations indicate instead an approximation of $Q{\rm nuc}=0.96+6.6\langle X\rangle$ (Goodwin et al.
+2018, in preparation).
+[{\it comment on improved quadratic expression}]
+%
+Thus, by substituting in equation \ref{qnuc} we can estimate 
+\begin{equation}
+\langle X\rangle 
+% In [42]: c.c**2/(1.*u.MeV/c.m_p).to("m**2/s**2")/6.6
+% Out[42]: <Quantity 142.16243658783114>
+% = \frac{c^2z/\alpha}{6.6\ {\rm Mev/nucleon}}\frac{\xi_b}{\xi_p} - 0.145
+= z\frac{142}{\alpha}\frac{\xi_b}{\xi_p} - 0.145 \label{xbar}
+\end{equation}
+[{\it comment on the degree of error introduced by ignoring anisotropy etc.}] 
+%
+Clearly this expression is only applicable for $\alpha$-values up to some limit, which we calculate as
+\begin{equation}
+%In [43]: c.c**2/(1.*u.MeV/c.m_p).to("m**2/s**2")/6.6/0.145
+%Out[43]: <Quantity 980.4305971574562>
+\alpha \leq 980\, z \frac{\xi_b}{\xi_p}
+\end{equation}
+Interestingly, this limit is $\approx70$\% larger than the equivalent value derived for the old expression for $Q_{\rm nuc}$.
+This result suggests that larger observed values of $\alpha$ may be accommodated without resorting to explanations including incomplete burning of burst fuel \cite[e.g.][]{bcatalog}.
+
+The hydrogen fraction at ignition $\langle X\rangle$ and the burst recurrence time may then be used to estimate the hydrogen fraction in the fuel, $X_0$.
+The H-fraction at the base of the fuel layer is reduced steadily by $\beta$-limited hot CNO burning, and will completely exhaust the accreted hydrogen in a time \cite[]{lampe16}:
+\begin{equation}
+t_{\rm CNO} =9.8\left(\frac{X_0}{0.7}\right) \left(\frac{Z_{\rm CNO}}{0.02}\right)^{-1}
+\end{equation}
+The hydrogen fraction at the base will thus be $X_0(1-\Delta t/t_{\rm CNO})$, [{\it should this expression include time dilation?}] so that the average H-fraction in the layer will be $\langle X \rangle = X_0(1-0.5\Delta t/t_{\rm CNO})$.
+We can combine this expression with equation \ref{xbar} to give
+\begin{eqnarray}
+X_0 & = & \langle X\rangle +0.35\left(\frac{\Delta t}{9.8\ {\rm hr}}\right)
+                            \left(\frac{Z_{\rm CNO}}{0.02}\right) \nonumber \\
+    & = & z\frac{142}{\alpha}\frac{\xi_b}{\xi_p} - 0.145 
+                           +\left(\frac{\Delta t}{28\ {\rm hr}}\right)
+                            \left(\frac{Z_{\rm CNO}}{0.02}\right)
+\end{eqnarray}
+Clearly this estimate of $X_0$ will incorporate systematic errors arising from the assumed values of $z$, the inclination, and $Z_{\rm CNO}$. We explore the typical range of these errors in 
+% the next section.
+\S\ref{sec:simulations}
+
+% comment on the inability to measure distance/mass/radius? Or can we actually do that?
+% I think you can get distance by comparing the ignition column with the fluence
+
+\subsection{Inferring distance} 
+\label{subsec:distance}
+
+The observed burst energy permits estimation of the ignition depth, assuming complete fuel consumption.
+For a burst fluence $E_b$, the total burst energy release (taking into account the anisotropy effect) is
+\begin{equation}
+E_{\rm burst} = 4\pi d^2 \xi_b E_b 
+\end{equation}
+The ignition depth is given by 
+\begin{eqnarray}
+y_{\rm ign} & = & E_{\rm burst}\frac{1 + z}{4\pi R_{\rm NS}^2 Q_{\rm nuc}} \\
+  & = & E_b \frac{\xi_b (1 + z)}{ Q_{\rm nuc}} \left(\frac{d}{R_{\rm NS}}\right)^2 
+\end{eqnarray}
+Independently, we can estimate the accreted mass as simply $\Delta M = ...$ and thus estimate the distance...?
+
 \subsection{Observation-model comparisons} 
@@
 
-The model-predicted quantities must then be converted to what a distant observer would see, by taking into account the effects of general relativistic time dilation and redshift.
-
-Following \cite{lampe16}, we write the corrections as follows \cite[see also][]{keek11}.
-For self-consistency, we need to identify a mass and radius for the neutron star for which the Newtonian potential equals the GR potentials, i.e.
-\begin{equation}
-\frac{GM}{r^2} = \frac{GM_{\rm GR}}{R^2_{\rm GR}\sqrt{1-2GM_{\rm GR}/(c^2R_{\rm GR})}} = \frac{GM_{\rm GR}}{R^2_{\rm GR}}(1+z)
-\end{equation}
-where $1+z$ is the gravitational redshift:
-\begin{equation}
-1+z = \frac{1}{\sqrt{1-2GM_{\rm GR}/(c^2R_{\rm GR})}}
-\end{equation}
-This equality is generally achieved by assuming that $M=M_{\rm GR}$, and solving for $R_{\rm GR}$. 
-We define $\xi$ such that $R_{\rm GR}=\xi R$. 
-One advantage of this choice is that the mass accretion rate is identical in the Newtonian and observer frames, and also that $\xi = \sqrt{1 + z}$.
-
-An additional correction that must be applied is the effect of the anisotropic emission of the burst, based on the system inclination $i$.
-This effect has been simulated for a range of disk geometries \cite[e.g.][]{he16}.
-Because the emission region on the surface of the star is thought to be different for the burst and persistent emission, we define separate anisotropy parameters $\xi_b$ and $\xi_p$, with the sense that
-\begin{equation}
-L_{b,p} = 4\pi d^2\xi_{b,p}F_{b,p}
-\end{equation}
-where $L_b$, $L_p$ are the burst and persistent luminosity, $F_b$, $F_p$ are the corresponding (observed) fluxes, and $d$ is the source distance.
-With this sense it can be seen that values of $\xi_{b,p}<1$ correspond to emission preferentially beamed towards us, so that --- without the anisotropy corrections --- the isotropic luminosity inferred from the observed flux would overestimate the actual luminosity.
-
-Of the model input parameters, we can also infer the persistent flux level expected given the model-assumed accretion rate $\dot{m}$, as follows: % gal03d, equation 2
-\begin{equation}
-F_{p,{\rm inf}} = \frac{L_p}{4\pi d^2} = 
-% \frac{\dot{M} Q_{\rm grav}}{1+z} \left(\frac{R_{\rm NS}}{d}\right)^2 \xi_p^{-1}
-% expression used in code
-\frac{\dot{m} Q_{\rm grav}}{4\pi d^2 (1+z)\xi_p}% c_{\rm bol}}
-\end{equation}
-where $Q_{\rm grav} = c^2z/(1+z) \approx GM_{\rm NS}/R_{\rm NS}$ is the gravitational energy release per gram. % gal03d
-
-The predicted quantities can then be converted to values suitable for comparison to observations, as follows
-\begin{eqnarray}
-\Delta t_{{\rm pred},\infty} & = & (1+z)\Delta t_{\rm pred} \\
-% E_{{\rm pred},\infty} & = & \xi^2E_{\rm pred}/(1+z)^2 \\ % CHECK
-% add in the F_per (derived from the mdot) here; see Lampe et al. (2016), eq 8
-t_{i,\infty} & = & (1+z)t_i \\
-F_{{\rm pred},i,\infty} & = & \xi^2\frac{L_{{\rm pred},i}}{4\pi d^2\xi_b(1+z)^2}
-\end{eqnarray}
+The model-predicted quantities must then be converted to what a distant observer would see, by taking into account the effects of source distance, emission anisotropy, general relativistic time dilation and redshift.
+For the purposes of comparison, we convert the model lightcurves to what a distant observer would see, following the approach described in \S\ref{subsec:simobs}.
 

### 2018-04-09 anisotropy.py

@@ -50,7 +50,8 @@ def anisotropy(inclination, model='he16_a', test=False):
         plt.ylabel(r"$\xi_p/\xi_b$")
 
         s='-'
-        for m in ['fuji88' ,'he16']:
+#        for m in ['fuji88' ,'he16']:
+        for m in ['fuji88' ,'he16_a']:
             xi_p = np.zeros(len(_theta))
             xi_b = np.zeros(len(_theta))
             for i, _th in enumerate(_theta):

### 2018-04-09 burstclass.py

@@ -187,6 +187,8 @@ def modelFunc(p,obs,model, disc_model):
 # Here we calculate the value of xi (ratio of GR to Newtonian radii),
 # appropriate for the adopted value of (1+z). This is used instead of the
 # value attached to the model, because that's for a different redshift
+#
+# Note that this isn't actually used below
 
     xi = sqrt(_opz)
 
@@ -205,7 +207,9 @@ def modelFunc(p,obs,model, disc_model):
 #   L_b = 4\pi d^2 \xi_b F_b
 
 #    return ( (model.xi/opz)**2 
-    return ( (xi/_opz)**2 
+# since xi = sqrt(_opz) (see above), this first term just becomes 1/_opz
+#    return ( (xi/_opz)**2 
+    return ( (1/_opz)
             * fInterp(obs.time+(0.5-obs.timepixr)*obs.dt)*model.lumin.unit
             / (4.*pi*dist.to('cm')**2) / xi_b )
 

### 2018-09-23 doc/ms.tex
Fixed bug in fluence method; force inclination to degrees for prior

@@
 
-\title{Thermonuclear burst model-observation comparisons with {\sc concord}}
+% \title{Thermonuclear burst model-observation comparisons with {\sc concord}}
+\title{Robust inference of neutron-star parameters from thermonuclear burst lightcurves}
 
@@
 \author[0000-0002-6558-5121]{Duncan K. Galloway}
+
+\author{Zac Johnston}
+\author{Adelle Goodwin}
+
 \affil{School of Physics \& Astronomy \\
@@
 
-Here we describe a few different approaches that have been used to deduce system parameters from observations of thermonuclear bursts.
+Here we describe three different approaches that have been used to deduce system parameters from observations of thermonuclear bursts.
 
@@
 \langle X\rangle 
-% In [42]: c.c**2/(1.*u.MeV/c.m_p).to("m**2/s**2")/6.6
+% In [42]: c.c**2/(6.6*u.MeV/c.m_p).to("m**2/s**2")
 % Out[42]: <Quantity 142.16243658783114>
@@
 
+% upper limit on xbar
+Conversely, for low values of $\alpha$, we may apply the constraint that $\langle X\rangle \lesssim 0.77$, corresponding to the expected maximum possible for accreted fuel with primordial abundances.
+
 The hydrogen fraction at ignition $\langle X\rangle$ and the burst recurrence time may then be used to estimate the hydrogen fraction in the fuel, $X_0$.
@@
 \end{equation}
-The hydrogen fraction at the base will thus be $X_0(1-\Delta t/t_{\rm CNO})$, [{\it should this expression include time dilation?}] so that the average H-fraction in the layer will be $\langle X \rangle = X_0(1-0.5\Delta t/t_{\rm CNO})$.
-We can combine this expression with equation \ref{xbar} to give
+measured in the neutron star frame.
+The hydrogen fraction at the base will thus be $X_0[1-\Delta t/(1+z)t_{\rm CNO}]$, 
+% [{\it should this expression include time dilation?}] -- yes
+so that the average H-fraction in the layer will be $\langle X \rangle = X_0[1-0.5\Delta t/(1+z)t_{\rm CNO}]$. We can combine this expression with equation \ref{xbar} to give
 \begin{eqnarray}
-X_0 & = & \langle X\rangle +0.35\left(\frac{\Delta t}{9.8\ {\rm hr}}\right)
+X_0 & = & \langle X\rangle +0.35\left[\frac{\Delta t}{(1+z)9.8\ {\rm hr}}\right]
                             \left(\frac{Z_{\rm CNO}}{0.02}\right) \nonumber \\
     & = & z\frac{142}{\alpha}\frac{\xi_b}{\xi_p} - 0.145 
-                           +\left(\frac{\Delta t}{28\ {\rm hr}}\right)
+                           +\left[\frac{\Delta t}{(1+z)28\ {\rm hr}}\right]
                             \left(\frac{Z_{\rm CNO}}{0.02}\right)
@@
 \S\ref{sec:simulations}
+
+\begin{itemize}
+\item Apply the method to the reference bursts
+\end{itemize}
 
@@
 %We take the approach of keeping $R_{\rm NS}$ fixed and allowing $M_{\rm NS}$ to vary to give the required change in $1+z$, since many equations of state have roughly constant radii over a range of masses \cite[e.g.][]{lp07}.
+
+\begin{itemize}
+\item Describe Zac's method
+\item Describe my method (via multinest)
+\end{itemize}
 

### 2018-09-23 burstclass.py

@@ -512,7 +512,7 @@ class Lightcurve(object):
             print ('** WARNING ** extrapolated fluence > stat_error, replacing')
             return fluen, f_int
 
-        return fluen, fluen_stat
+        return fluen, fluene_stat
 
 # ======= ========= ========= ========= ========= ========= ========= =========
 
@@ -1128,6 +1128,11 @@ class KeplerBurst(Lightcurve):
 
 def lnprior(theta):
     dist, inclination, _opz, t_off = theta
+
+# Have to enforce this here or the prior will come out wrong
+
+    assert hasattr(inclination,'unit')
+    assert inclination.unit == 'deg'
 
 # We have currently flat priors for everything but the inclination, which
 # has a probability distribution proportional to sin(i)

### 2018-09-23 concord/data/table2.tex
Converted concord to package and updated scripts
A major rearrangement to allow concord to be structured as a package, and
installed using pip. This modification means that the $CONCORD_PATH
environemnt variable is no longer required; the code can find the data
files required via other methods

The anisotropy function is now part of the diskmodel subpackage.

Also updated the tutorial notebook to make sure the code still works after
all the changes. The data files required are now part of the repo, in the
example_data subdirectory

\begin{table*}[h]
\caption{Properties of thermonuclear bursts observed from target sources by
the {\it Rossi X-ray Timing Explorer}}\label{tab:bursts}
\begin{tabular}{lccccccccc}
Epoch & $N_{\rm burst}$ ($N_{\rm av}$) & burst IDs$^a$ & $\Delta t$ (hr) & $F_{\rm per}^b$ ($10^{-9}\ \epcs$) & $c_{\rm bol}$  & $\dot{m}$ ($\dot{m}_{\rm Edd}$) & $E_b$ ($10^{-6}\ \epc$) & $F_{\rm pk}$ ($10^{-9}\ \epcs$) & $\alpha$ \\
1998 Jun & 6(1) & 7 & $5.14\pm0.07$ & $1.167\pm0.006$ & $1.806\pm0.009$ & 0.0513 & $1.102\pm0.011$ & $30.9\pm1.0$ & $34.2\pm0.5$ \\ 
2000 Sep & 11(7) & 14--20 & $4.177\pm0.010$ & $1.593\pm0.017$ & $1.787\pm0.003$ & 0.0692 & $1.126\pm0.016$ & $29.1\pm0.5$ & $38.6\pm0.3$ \\ 
2007 Mar & 10(7) & 58--65 & $3.530\pm0.004$ & $1.87\pm0.02$ & $1.751\pm0.003$ & 0.0796 & $1.18\pm0.04$ & $28.4\pm0.4$ & $35.3\pm1.0$ \\ 
2002 Oct & 1(1) & 2 & $16.55\pm0.06^c$ & 2.541--2.298 & $2.085\pm 0.019$ & 0.0472$^d$ & $2.649\pm0.018$ & $229\pm4$ & $114.4\pm1.9$ \\ 
         & 1(1) & 3 & 21.10 & 2.298--1.946 & $2.13\pm0.04$ & 0.0432$^d$ & $2.990\pm0.017$ & $232\pm4$ & $118.2\pm1.9$ \\ 
         & 1(1) & 4 & 29.82 & 1.946--1.826 & $2.157\pm0.002$ & 0.0384$^d$ & $3.46\pm0.02$ & $232\pm4$ & $128.2\pm2.1$ \\ 
1997 May 4 & 3(1) & 1 & $2.681\pm0.007$ & $3.72\pm0.18$ & $1.45\pm0.09$ & 0.144 & $0.381\pm0.003$ & $61\pm2$ & $138.5\pm1.4$ \\
2009 Jun 12 & 2(2) & & $1.892^e$ & $5.70\pm0.04$ & 1.4981 & 0.226 & $0.371\pm0.010$ & $56.6\pm1.4$ & $160.1\pm1.8$\\ 
2001 Feb & 1(1) & & $\leq 1.53\times10^{4}$~$^f$ & $4.73\pm0.03$ & 1.5346 & 0.167 & $110\pm9$~$^g$ & $21.9\pm0.6$~$^g$ & \\
\hline
\end{tabular}
\medskip\\
$^a$ burst index number in the catalog of \cite{bcatalog}.\\
$^b$ persistent flux in the energy range 3--25 keV \\
$^c$ inferred from ignition model comparisons \\
$^d$ calculated based on linear interpolation between observations falling within the burst interval \\
$^e$ only two bursts detected, so may not have been part of a regular train \\
$^f$ recurrence time upper limit of 1.75~yr based on the next earliest event, detected by the \xte/ASM \cite[]{kuul04} \\
$^g$ values taken from the ``optimal'' fits of \cite{keek14a}

\end{table*}

### 2018-09-23 concord/__init__.py

@@ -1,0 +1,8 @@
+import pkg_resources
+
+CONCORD_PATH = pkg_resources.resource_filename('concord','data')
+
+# Ensure objects are available at the package level
+
+# from concord import burstclass
+from .burstclass import *

### 2018-09-23 concord/burstclass.py

@@ -52,11 +52,14 @@ from astroquery.vizier import Vizier
 from chainconsumer import ChainConsumer
 from datetime import datetime
 
-from anisotropy import *
+from concord import diskmodel as dm
 
 # Get the path for the concord files from the environment variable
 
-CONCORD_PATH = os.environ['CONCORD_PATH']
+import pkg_resources
+
+CONCORD_PATH = pkg_resources.resource_filename('concord','data')
+# CONCORD_PATH = os.environ['CONCORD_PATH']
 
 # ------- --------- --------- --------- --------- --------- --------- ---------
 
@@ -194,7 +197,7 @@ def modelFunc(p,obs,model, disc_model):
 # Use the anisotropy function to calculate the anisotropy factors given
 # the inclination
 
-    xi_b, xi_p = anisotropy(inclination)
+    xi_b, xi_p = dm.anisotropy(inclination)
 
 # Here we calculate the value of xi (ratio of GR to Newtonian radii),
 # appropriate for the adopted value of (1+z). This is used instead of the
@@ -376,7 +379,7 @@ class Lightcurve(object):
             return None
 
         dist, inclination, _opz, t_off = param
-        xi_b, xi_p = anisotropy(inclination, model=disc_model)
+        xi_b, xi_p = dm.anisotropy(inclination, model=disc_model)
 
         if (obs == None):
 
@@ -700,7 +703,7 @@ class ObservedBurst(Lightcurve):
         if not 0. < inclination.value < 90.:
             return -np.inf
 
-        xi_b, xi_p = anisotropy(inclination, model=disc_model)
+        xi_b, xi_p = dm.anisotropy(inclination, model=disc_model)
 
 # Here we calculate the equivalent mass and radius given the redshift and
 # surface gravity. Since we allow the redshift to vary, but the model is
@@ -1179,13 +1182,16 @@ def apply_units(params,units = (u.kpc, u.degree, None, u.s)):
 
 # ------- --------- --------- --------- --------- --------- --------- ---------
 
-def lhoodClass(params, obs, model, weights, disc_model):
+def lhoodClass(params, obs, model, **kwargs):
     '''
     Calculate the likelihood related to one or more model-observation
     comparisons The corresponding call to emcee will (necessarily) look
     something like this:
 
         sampler = emcee.EnsembleSampler(nwalkers, ndim, lhoodClass, args=[obs, models, weights])
+
+    Use the kwargs construction to pass the additional parameters weights,
[…]

### 2018-09-23 concord/diskmodel.py

@@ -6,7 +6,10 @@ from scipy.interpolate import interp1d
 import astropy.units as u
 import os
 
-CONCORD_PATH = os.environ['CONCORD_PATH']
+import pkg_resources
+
+# CONCORD_PATH = os.environ['CONCORD_PATH']
+CONCORD_PATH = pkg_resources.resource_filename('concord','data')
 he16_models = ['he16_a', 'he16_b', 'he16_c', 'he16_c_short', 'he16_d']
 
 
@@ -64,6 +67,8 @@ def anisotropy(inclination, model='he16_a', test=False):
             s=':'
 
         plt.legend()
+
+        plt.show()
 
 # Calculate the values for the passed quantity, and return
 # numpy cos will correctly treat the units, so no need to do a conversion

### 2018-09-23 setup.py

@@ -1,0 +1,30 @@
+import setuptools
+import glob
+
+with open("README.md", "r") as fh:
+    long_description = fh.read()
+
+setuptools.setup(
+    name="concord",
+    version="0.0.1",
+    author="Duncan Galloway",
+    author_email="Duncan.Galloway@monash.edu",
+    description="Package for multi-epoch comparisons of model thermonuclear bursts to observations",
+    long_description=long_description,
+    long_description_content_type="text/markdown",
+    url="http://bitbucket.org/minbar/concord",
+    packages=setuptools.find_packages(),
+#    packages=['concord'],
+#    package_dir={'concord': "concord"},
+# The following line was intended to force importing of the diskmodel
+# module, but it doesn't seem to work
+    py_modules = ["diskmodel"],
+#
+    package_data={'': ['data/*']},
+    scripts=glob.glob('scripts/*'),
+    classifiers=[
+        "Programming Language :: Python :: 3",
+        "License :: OSI Approved :: MIT License",
+        "Operating System :: OS Independent",
+    ],
+)

### 2019-02-25 concord/burstclass.py
Updated the observe method for the Lightcurve class (applicable to any object with a lumin attribute, including KeplerBursts). Verified to work whether or not you supply an observed lightcurve

@@ -1,6 +1,5 @@
 # Burst class for MINBAR
-
-
+#
 # This code is a preliminary attempt to develop a burst class for MINBAR,
 # including model-observation comparisons. 
 # 
@@ -140,12 +139,12 @@ class Lightcurve(object):
 
     Example: simulated burst (giving time and luminosity)
 
-    Lightcurve.__init__(self, time=d['time']*u.s,
+    Lightcurve(self, time=d['time']*u.s,
         lumin=d['luminosity']*u.erg/u.s, lumin_err=d['u_luminosity']*u.erg/u.s)
 
     Example: observed burst giving flux
 
-    Lightcurve.__init__(self, time=d['col1']*u.s, dt=d['col2']*u.s,
+    Lightcurve(self, time=d['col1']*u.s, dt=d['col2']*u.s,
         flux=d['col3']*1e-9*u.erg/u.cm**2/u.s,
         flux_err=d['col4']*1e-9*u.erg/u.cm**2/u.s)
     """
@@ -184,10 +183,13 @@ class Lightcurve(object):
         if 'flux' in kwargs:
             self.flux = kwargs.get('flux',None)
             self.flux_err = kwargs.get('flux_err',None)
+            # print ('flux',self.flux)
+            # print ('flux_err',self.flux_err)
 
             self.good = np.where(self.flux_err < self.flux)[0]
-            for i in range(len(self.good)-1):
-                self.dt_nogap[self.good[i]] = max(
+            if len(self.good) > 0:
+                for i in range(len(self.good)-1):
+                    self.dt_nogap[self.good[i]] = max(
     [self.dt[self.good[i]],self.time[self.good[i+1]]-self.time[self.good[i]]] )
         else:
             self.good = np.arange(len(self.lumin))
@@ -297,9 +299,11 @@ class Lightcurve(object):
         disc_model='he16_a',c_bol=1.0):
         """
         Convert a luminosity profile to a simulated observation, with
-        plausible errors
+        plausible errors.
         This method might make more sense as part of the KeplerBurst class
-        (or a parent SimulatedBurst class)
+        (or a parent SimulatedBurst class), but for now we include this method
+        as part of the lightcurve class, so that future classes of model
+        bursts can access it
         """
 
         if not hasattr(self,'lumin'):
@@ -311,14 +315,16 @@ class Lightcurve(object):
 
         if (obs == None):
 
-# No observation is provided, so we have to make up a burst lightcurve
-# with some dummy values for the time bins and fluxes
[…]

### 2019-02-25 doc/ms.tex
Minor changes to the info routines, also started updating the paper text

@@
 %
+We seek to calculate the burst lightcurve that would be observed given a model lightcurve and neutron star parameters including distance $d$; inclination $i$; and surface redshift $(1+z)_{\rm fit}$ (so labeled to distinguish from the corresponding value implict to the model).
 The process can be summarised as follows.
+[{\bf note doesn't include any steps to deal with the redshift inconsistency, see below}]
 \begin{enumerate}
-\item Multiply the time bins for the predicted burst lightcurve by the gravitational redshift $1+z$, thereby ``stretching'' the profile to account for the general relativistic time dilation at the neutron-star surface
+\item Multiply the time bins for the predicted burst lightcurve by the trial gravitational redshift $(1+z)_{\rm fit}$, thereby ``stretching'' the profile to account for the general relativistic time dilation at the neutron-star surface
 % \item Shift the predicted burst lightcurve by $t_{\rm off}$
@@
 % multiplying by $[\xi/(1+z)]^2$
-dividing by $1+z$
+dividing by $(1+z)_{\rm fit}$
 \item Convert the luminosity to (isotropic) flux by dividing by the distance factor, $4\pi d^2$ \label{distfac}
@@
 %
-However, we are free to vary the surface redshift $1+z$, for example to achieve improved agreement with an observed profile.
+We are free (in principle) to vary the surface redshift $1+z$, for example to achieve improved agreement with an observed profile (although see below).
 
@@
 
-We note that the combination of surface gravity $g$ (assumed for the model) and adopted $1+z$  uniquely specifies the neutron star mass $M_{\rm NS}$ and radius $R_{\rm NS}$:
+% The corrections described below
+However, since the model is to be transformed by a different value of $(1+z)_{\rm fit}$, we must take care to consistently transform the model parameters in accordance with this choice.
+%
+We can achieve this objective as follows.
+The surface gravity $g$ adopted for the model, and the trial redshift $(1+z)_{\rm fit}$ imply a different neutron star mass and radius than adopted for the model:
+%
+% We note that the combination of surface gravity $g$ (assumed for the model) and adopted $1+z$  uniquely specifies the neutron star mass $M_{\rm NS}$ and radius $R_{\rm NS}$:
 \begin{eqnarray}
 % as used in function calc_mr in burstclass.py
-R_{\rm NS} &=& c^2\frac{(1+z)^2-1}{2g(1+z)}\\
-M_{\rm NS} &=& \frac{gR_{\rm NS}^2}{G(1+z)}
+R_{\rm NS,fit} &=& c^2\frac{(1+z)_{\rm fit}^2-1}{2g(1+z)_{\rm fit}}\\
+M_{\rm NS,fit} &=& \frac{gR_{\rm NS}^2}{G(1+z)_{\rm fit}}
 \end{eqnarray}
-Thus, by identifying the optimal value of $1+z$ for comparison to a particular observation, we can constrain the mass and radius, at a fixed $g$.
+Thus, by identifying the optimal value of $(1+z)_{\rm fit}$ for comparison to a particular observation, we can constrain the mass and radius, at a fixed $g$.
+[{\bf so which one should I use?
+and should I make any other changes to the recipe?}]
 

### 2019-02-25 concord/burstclass.py

@@ -50,6 +50,7 @@ import pkg_resources
 
 CONCORD_PATH = pkg_resources.resource_filename('concord','data')
 # CONCORD_PATH = os.environ['CONCORD_PATH']
+ETA = 1.e-6
 
 # ------- --------- --------- --------- --------- --------- --------- ---------
 
@@ -343,6 +344,11 @@ class Lightcurve(object):
         # Have a problem here with the doubling up of units, as the ObservedBurst.__init__
         # method will apply it's own
 
+        if hasattr(self,'opz'):
+            if abs(self.opz-_opz) > ETA:
+                print ('observe: ** WARNING ** inconsistent simulation redshift for model value')
+                print ('         model value: {:.4f}, simulation value: {:.4f}\n'.format(self.opz,_opz))
+
         # print ('obs.time:',obs.time)
         # print ('obs.dt:',obs.dt)
         # print ('flux:',model)
@@ -351,7 +357,12 @@ class Lightcurve(object):
                           flux=model.value,flux_err=obs.flux_err.value,
                           tdel = self.tdel*_opz,tdel_err=self.tdel_err*_opz,
                           fper = fper(self.mdot,_opz,dist,xi_p,c_bol=c_bol),
-                          filename="{} @ {}".format(self.filename,dist))
+                          filename="{} @ {}".format(self.filename,dist),
+        # Include the simulation parameters in the burst description
+                          sim_dist=dist, sim_inclination=inclination, sim_opz=_opz,
+                          sim_t_off=t_off, sim_xi_b=xi_b, sim_xi_p=xi_p, sim_disc_model=disc_model
+        # Should potentially also include the model burst parameters, where available
+                            )
 
         return sim
 
@@ -667,13 +678,28 @@ class ObservedBurst(Lightcurve):
 
         print("\nObservedBurst parameters:")
         if hasattr(self,'tdel'):
-            print ("  tdel = {}".format(self.tdel))
+            print ("  tdel = {:.4f}".format(self.tdel))
         fluen, fluen_err = self.fluence(warnings=False)
         print ("  Fluence = {:.3e} +/- {:.3e}".format(fluen.value, fluen_err))
-        if hasattr(self,'fper'):
-            print ("  F_per = {} +/- {}".format(self.fper,self.fper_err))
+        if hasattr(self,'fper') & hasattr(self,'fper_err'):
[…]

### 2019-02-26 concord/burstclass.py
Updated observe method to use correct scalings for GR. Modified the compare method to use the observe method (instead of it's own "custom" version). Updated the plot routine to allow arguments to be passed to the plot commands, and also set up a convention for colors for observed and simulated bursts

@@ -16,7 +16,7 @@
 # class ObservedBurst(Lightcurve):
 # class KeplerBurst(Lightcurve):
 #
-# def fper(mdot,opz,dist,xi_p,c_bol=1.0):
+# def fper(mburst, param, c_bol=1.0):
 # def modelFunc(p,obs,model, disc_model):
 # def lnprior(theta):
 # def apply_units(params,units = (u.kpc, u.degree, None, u.s)):
@@ -51,28 +51,53 @@ import pkg_resources
 CONCORD_PATH = pkg_resources.resource_filename('concord','data')
 # CONCORD_PATH = os.environ['CONCORD_PATH']
 ETA = 1.e-6
+# default colours for plotting
+OBS_COLOUR = 'b'
+MODEL_COLOUR = 'b'
 
 # ------- --------- --------- --------- --------- --------- --------- ---------
 
-def fper(mdot,opz,dist,xi_p,c_bol=1.0):
+def fper(mburst, param, c_bol=1.0):
     '''
     Calculates the persistent flux, based on the inferred mdot, redshift
-    etc.
+    etc. Earlier we passed the individual parameters, but easiest to just provide
+    the model burst
     '''
 
-# First we calculate Q_grav, to calculate the inferred persistent flux that
-# we should see (based on the accretion rate); that is not given in Lampe
-# et al.  2016, but is given in gal03d
-# Note that (as for the M_NS) we're using the passed parameter opz here,
-# not the value that is associated with the model burst
+    dist, inclination, _opz, t_off = param
+    xi_b, xi_p = dm.anisotropy(inclination)
+
+    # The combination of an input redshift an model gravity uniquely define
+    # a mass-radius combination (as for modelFunc)
+
+    _M, _R = calc_mr(mburst.g, _opz)
+
+    # Here we calculate the value of xi (ratio of GR to Newtonian radii),
+    # appropriate for the adopted value of (1+z). This is used instead of the
+    # value attached to the model, because that's for a different redshift
+
+    # xi = sqrt(_opz)
+    xi = (_R / mburst.R_Newt).decompose()
+    # print (xi)
+
+    # The mdot measured by the distant observer now depends upon the assumed
+    # redshift; see Keek & Heger (2011) eq. B16
+
+    mdot_infty = xi**2 * mburst.mdot / _opz
+
+    # Next we calculate Q_grav, to calculate the inferred persistent flux that
+    # we should see (based on the accretion rate); that is not given in Lampe
[…]

### 2019-08-26 doc/ms.tex
Updated observe method to use correct scalings for GR. Modified the compare method to use the observe method (instead of it's own "custom" version). Updated the plot routine to allow arguments to be passed to the plot commands, and also set up a convention for colors for observed and simulated bursts
Added extra routines designed for inferring neutron star properties from bursts, copied from Inferring\ composition.ipynb

@@
 
+\usepackage{amsmath}	% to enable cases environment
+
 %% AASTeX v6.* now includes \hyperref support. While we have built in specific
@@
 \newcommand\latex{La\TeX}
+
+% Custom definitions
+
+\newcommand{\eps}{{\rm erg\,s^{-1}}}
+\newcommand{\epcs}{{\rm erg\,cm^{-2}\,s^{-1}}}
+\newcommand{\epc}{{\rm erg\,cm^{-2}}}
 
@@
 % \title{Thermonuclear burst model-observation comparisons with {\sc concord}}
-\title{Robust inference of neutron-star parameters from thermonuclear burst lightcurves}
+\title{Robust inference of neutron-star parameters from thermonuclear burst 
+% lightcurves}
+observations}
 
@@
 
-The fundamental observables that we measure include the recurrence time $\Delta t$; the persistent flux $F_{\rm per}$, a measure of the accretion rate; and the burst bolometric flux lightcurve, $(t_i,F_{i})$.
-From the burst lightcurve we may integrate to estimate the burst fluence, $E_b$, which may be used to estimate the $\alpha$-value, the ratio of the burst to persistent flux:
+The fundamental observable that we measure is the time-history of the burst flux, the lightcurve $(t_i,F_{i})$.
+Here the $F_{i}$ may be in instrumental units of count~s$^{-1}$, or may be estimates of the bolometric flux at time $t_i$, derived from model fits of the spectrum at time $t_i$.
+%
+Provided the $F_i$ are in units of flux, or the flux can be estimated from the values,  we may measure the peak flux, $F_{\rm pk}$, and integrate to estimate the burst fluence, $E_b$.
+
+The peak flux $F_{\rm pk}$ provides constraints on the source distance, since the maximum burst luminosity is limited to (roughly) the Eddington luminosity \cite[e.g.][]{lew93}
+% copied from bcatalog
+\begin{eqnarray}
+  L_{\rm Edd,\infty} & = & \frac{8\pi G m_p M_{\rm NS} c
+  [1+(\alpha_{\rm T}T_{\rm e})^{0.86}]} {\sigma_T(1+X)[1+z(R)]} % \xi_b
+       \nonumber \\
+  & = & 2.7\times10^{38} \left(\frac{M_{\rm NS}}{1.4M_\odot}\right)
+ \frac{1+(\alpha_{\rm T}T_{\rm e})^{0.86}}{(1+X)}
+\nonumber \\ & & \times\  % line break
+    \left[\frac{1+z(R)}{1.31}\right]^{-1}\ % \xi_b\
+              \eps
+  \label{ledd}
+\end{eqnarray}
+Here
+$T_{\rm e}$ is the
+effective temperature of the atmosphere, $\alpha_{\rm T}$ parametrizes the temperature dependence of the electron scattering
+opacity \cite[$\simeq 2.2\times10^{-9}$~K$^{-1}$;][]{lew93},
+$m_p$ is the proton mass, $\sigma_T$ the Thompson 
+cross-section, and $X$ the
+mass fraction of hydrogen in the atmosphere.
+% ($\approx0.7$ for cosmic abundances).
+The final factor in
+square brackets represents the gravitational redshift 
+at the photosphere $1+z(R)=(1-2GM_{\rm NS}/R
+c^2)^{-1/2}$, which may be lower than the value at the NS surface, while the photosphere is expanded during the radius expansion episode (i.e. $R\ge R_{\rm NS}$).
+%
+Mention also the empirical value here...
+
+A significant fraction of observed bursts exhibit characteristic variations in their spectral hardness (or the blackbody temperature and normalisation) around their maximum that indicate the presence of photospheric radius-expansion (PRE).
+By equating the maximum flux of these events with the Eddington luminosity, the distance to the source can be estimated as
+\begin{eqnarray}
+ d & = & \left(\frac{L_{\rm Edd,\infty}}{4\pi \xi_b F_{\rm pk, RE}}\right)^{1/2} \nonumber \\
+   & = & 8.6
+	\left( \frac{\xi_b F_{\rm pk, RE}}{3\times10^{-8}\ \epcs} \right)^{-1/2}
+       	\left(\frac{M_{\rm NS}}{1.4M_\odot}\right)^{1/2}
+\nonumber \\ & & \times\  % line break
+	\left[\frac{1+z(R)}{1.31}\right]^{-1/2}
+	(1+X)^{-1/2}\ {\rm kpc}
+ \label{disteq}
+\end{eqnarray}
+% 
+Here the $\xi_b$ factor accounts for the possible anisotropy of the burst emission, and is defined in the same sense as \citep{fuji88}, such that $L_b=4\pi d^2\xi_b F_{\rm pk}$.
+Since the distribution of the burst flux over the neutron star is likely different from that of the persistent emission, we define a different anisotropy factor $\xi_p$ relating the observed persistent flux $F_p$ to the total luminosity.
+Modelling \cite[]{he16}...
+
+As is obvious from equation \ref{disteq}, several factors will contribute to statistical (and possibly systematic) uncertainties in distance estimates derived in this manner.
+First, the neutron star mass $M_{\rm NS}$ is unknown...
+%
+Second, the radius $R$ at which the redshift should be calculated may not be clear from the observations.
+In principle one could take the inferred blackbody nromalisation at the time of maximum flux, but for most PRE bursts the maximum flux is achieved close to the ``touchdown'' point, where the photosphere has returned (more or less) to the neutron star surface \cite[e.g.][]{gal06a}.
+%
+Third, the possible range/likely value for $X$...
+%
+Fourth, even if the inclination $i$ of the system is known, converting to the anisotropy factor requires modelling of the effect of the disk on the radiation field arising from the burst \cite[e.g.][]{he16}.
+
+These issues may be avoided by adopting instead the empirical Eddington luminosity measured by \cite[]{kuul03a}...
+
+The fluence (and the shape of the burst profile) is determined by the amount of accumulated fuel, and its composition.
+The nuclear energy generation rate $Q_{\rm nuc}$ depends on the composition; studies with the 1-D numerical code {\sc kepler} suggest that
+\begin{eqnarray}
+Q_{\rm nuc} &= & 1.31+6.95\bar{X}-1.92\bar{X}^2 \nonumber \\
+& \approx & 1.35 + 6.05 \bar{X}
+\end{eqnarray}
+\cite[]{goodwin19a} where $\bar{X}$ is the average H-fraction of the fuel layer at ignition.
+
+Perhaps put in here some estimates of the fuel composition as a function of $X_0$ and recurrence time; could perhaps develop a ``prior'' for $\bar{X}$
+
+Provided an estimate of $\bar{X}$ is available, we can estimate the burst column that has ignited as
+\begin{eqnarray}
+y & = & \frac{L_b d^2 (1+z)}{R_{\rm NS}^2 Q_{\rm nuc}}\ 
+                                                            \nonumber \\
+  & = & 3.0\times10^8 \left(\frac{\xi_b E_{\rm b}}{10^{-6}\ \epc}\right)
+                      \left(\frac{d}{10\ {\rm kpc}}\right)^2\
+             \left(\frac{Q_{\rm nuc}}{4.4\ {\rm MeV/nucleon}}\right)^{-1}
+\nonumber \\ & & \times\  % line break
+                      \left(\frac{1+z}{1.31}\right)
+                      \left(\frac{R_{\rm NS}}{10\ {\rm km}}\right)^{-2}
+                      {\rm g\,cm^{-2}}
+\label{column}
+\end{eqnarray}
+{\it need to update prefactor, denominator in $Q_{\rm nuc}$ term here }
+
+From data taken prior to the burst ignition we can measure the persistent flux $F_{\rm per}$, a measure of the accretion rate; and provided more than one burst is observed, we may also measure (or estimate) the recurrence time, $\Delta t$.
+
+These parameters may be combined to estimate the $\alpha$-value, the ratio of the burst to persistent flux: 
 \begin{equation}
@@
 R_{\rm NS,fit} &=& c^2\frac{(1+z)_{\rm fit}^2-1}{2g(1+z)_{\rm fit}}\\
-M_{\rm NS,fit} &=& \frac{gR_{\rm NS}^2}{G(1+z)_{\rm fit}}
+M_{\rm NS,fit} &=& \frac{gR_{\rm NS,fit}^2}{G(1+z)_{\rm fit}}
 \end{eqnarray}
@@
 attributed to neutrino emission \cite[]{fuji87}.
-However, the 35\% value applies only to $\beta$-decays in the much more extensive {\it rp}-process burning chain, and recent {\sc kepler} simulations indicate instead an approximation of $Q{\rm nuc}=0.96+6.6\langle X\rangle$ (Goodwin et al.
-2018, in preparation).
+However, the 35\% value applies only to $\beta$-decays in the much more extensive {\it rp}-process burning chain, and recent {\sc kepler} simulations indicate instead an approximation of
+% $Q{\rm nuc}=0.96+6.6\langle X\rangle$ 
+$Q{\rm nuc}=1.35+6.05\langle X\rangle$ 
+% (Goodwin et al. 2018, in preparation). 
+\cite[]{goodwin19a}.
 [{\it comment on improved quadratic expression}]
@@
 \langle X\rangle 
-% In [42]: c.c**2/(6.6*u.MeV/c.m_p).to("m**2/s**2")
-% Out[42]: <Quantity 142.16243658783114>
-% = \frac{c^2z/\alpha}{6.6\ {\rm Mev/nucleon}}\frac{\xi_b}{\xi_p} - 0.145
-= z\frac{142}{\alpha}\frac{\xi_b}{\xi_p} - 0.145 \label{xbar}
+% In [42]: c.c**2/(6.0455*u.MeV/c.m_p).to("m**2/s**2")                             
+% Out[41]: <Quantity 155.20173377>
+%% = \frac{c^2z/\alpha}{6.6\ {\rm Mev/nucleon}}\frac{\xi_b}{\xi_p} - 0.145
+% = z\frac{142}{\alpha}\frac{\xi_b}{\xi_p} - 0.145 \label{xbar}
+ = z\frac{155}{\alpha}\frac{\xi_b}{\xi_p} - 0.223 \label{xbar}
 \end{equation}
@@
 \begin{equation}
-%In [43]: c.c**2/(1.*u.MeV/c.m_p).to("m**2/s**2")/6.6/0.145
-%Out[43]: <Quantity 980.4305971574562>
-\alpha \leq 980\, z \frac{\xi_b}{\xi_p}
-\end{equation}
-Interestingly, this limit is $\approx70$\% larger than the equivalent value derived for the old expression for $Q_{\rm nuc}$.
+%In [43]: c.c**2/(1.*u.MeV/c.m_p).to("m**2/s**2")/1.3455        
+% Out[43]: <Quantity 697.34082607>
+% \alpha \leq 980\, z \frac{\xi_b}{\xi_p}
+\alpha \leq 697\, z \frac{\xi_b}{\xi_p}
+\end{equation}
+Interestingly, this limit is 
+% $\approx70$\% 
+%  c.c**2/(1.*u.MeV/c.m_p).to("m**2/s**2")/1.6                 
+% Out[44]: <Quantity 586.42005092>
+% In [45]: 697./586.                                                               
+% Out[45]: 1.189
+$\approx20$\%
+%
+larger than the equivalent value derived for the old expression for $Q_{\rm nuc}$.
 This result suggests that larger observed values of $\alpha$ may be accommodated without resorting to explanations including incomplete burning of burst fuel \cite[e.g.][]{bcatalog}.
@@
 measured in the neutron star frame.
+% updated 2019 Aug 26 with exhaustion case
 The hydrogen fraction at the base will thus be $X_0[1-\Delta t/(1+z)t_{\rm CNO}]$, 
 % [{\it should this expression include time dilation?}] -- yes
-so that the average H-fraction in the layer will be $\langle X \rangle = X_0[1-0.5\Delta t/(1+z)t_{\rm CNO}]$. We can combine this expression with equation \ref{xbar} to give
-\begin{eqnarray}
-X_0 & = & \langle X\rangle +0.35\left[\frac{\Delta t}{(1+z)9.8\ {\rm hr}}\right]
-                            \left(\frac{Z_{\rm CNO}}{0.02}\right) \nonumber \\
-    & = & z\frac{142}{\alpha}\frac{\xi_b}{\xi_p} - 0.145 
-                           +\left[\frac{\Delta t}{(1+z)28\ {\rm hr}}\right]
-                            \left(\frac{Z_{\rm CNO}}{0.02}\right)
-\end{eqnarray}
+%
+provided that $\Delta t < (1+z)t_{\rm CNO}$.
+Once $\Delta t$ exceeds the time to burn all the hydrogen at the base, the abundance there will be zero, and a growing layer of pure He fuel will develop.
+The average H-fraction in the layer for these two cases will be
+%so that the average H-fraction in the layer will be 
+\begin{equation}
+\langle X \rangle = \begin{cases}
+%  X_0\left[1-0.5\frac{\Delta t}{(1+z)t_{\rm CNO}}\right] & \Delta t \leq (1+z)t_{\rm CNO} \\
+%  0.5X_0\frac{(1+z)t_{\rm CNO}}{\Delta t} & \Delta t > (1+z)t_{\rm CNO}
+  X_0(1-0.5f_{\rm burn}) & \Delta t \leq (1+z)t_{\rm CNO} \\
+  0.5X_0/f_{\rm burn} & \Delta t > (1+z)t_{\rm CNO}
+  \end{cases}
+\end{equation}
+where $f_{\rm burn} = \frac{\Delta t}{(1+z)t_{\rm CNO}}$ is the ratio of the recurrence time to the time to burn all the H.
+%
+We can combine this expression with equation \ref{xbar} to give
+\begin{equation}
+X_0 = \begin{cases}
+%        & \langle X\rangle +0.35\left[\frac{\Delta t}{(1+z)9.8\ {\rm hr}}\right]
+%                            \left(\frac{Z_{\rm CNO}}{0.02}\right) \nonumber \\
+         z\frac{142}{\alpha}\frac{\xi_b}{\xi_p} - 0.145 
+                           +\left[\frac{\Delta t}{(1+z)28\,{\rm hr}}\right]
+                       \left(\frac{Z_{\rm CNO}}{0.02}\right) & f_{\rm burn} \leq 1\\
+         \sqrt{ \frac{\Delta t}{(1+z)7\,{\rm hr}}\frac{Z_{\rm CNO}}{0.02}
+                        \left(z\frac{142}{\alpha}\frac{\xi_b}{\xi_p} - 0.145\right) }
+                        & f_{\rm burn} > 1
+        \end{cases}
+\end{equation}
+Practically, the issue with these expressions is that calculating $f_{\rm burn}$ requires knowledge of $X_0$, which is the unknown we are trying to constrain.
+One approach is to adopt a trial value of $X_0$, calculate $f_{\rm burn}$ and hence an updated estimate of $X_0$ via equation \ref{x0}, and iterate until no further change in the estimate arises.
+
 Clearly this estimate of $X_0$ will incorporate systematic errors arising from the assumed values of $z$, the inclination, and $Z_{\rm CNO}$. We explore the typical range of these errors in 

### 2019-08-26 concord/utils.py

@@ -1,13 +1,22 @@
 # Various utilities moved from burstclass.py
+# Augmented 2019 Aug with routines from Inferring\ composition.ipynb
 #
 # def g(M,R,Newt=False,units='cm/s^2'):
 # def opz(M,R):
 # def calc_mr(g,opz):
 # def solve_radius(M,R_Newt,eta=1e-6):
 # def decode_LaTeX(string):
+# def Q_nuc(Xbar, quadratic=False, old_relation=False, coeff=False)
+# def hfrac(alpha, tdel, opz=1.259, zcno=0.02, old_relation=False, ...)
+# def iso_dist(nsamp=1000, imin=0., imax=75.)
+# def dist(F_pk, F_pk_err, nsamp=10000, X=0.0, empirical=False, ...)
+# def mdot(F_per, F_per_err, dist, nsamp=10000, M_NS=1.4 * u.M_sun, R_NS=11.2 * u.km, ...)
+# def yign(E_b, E_b_err, dist, R_NS=10., Xbar=0.7, opz=1.259, ...)
+# def L_Edd(F_pk, F_pk_err, dist, dist_err=0.0, nsamp=10000, dip=False, ...)
 
 import numpy as np
 from math import *
+from .diskmodel import *
 import astropy.constants as const
 import astropy.units as u
 import re
@@ -109,3 +118,444 @@ def decode_LaTeX(string):
     else:
         return val, float(err_match.group(1))
 
+# ------- --------- --------- --------- --------- --------- --------- ---------
+
+def Q_nuc(Xbar, quadratic=False, old_relation=False, coeff=False):
+    '''
+    This function implements the approximation to the nuclear energy generation rate
+    Q_nuc given the mean hydrogen fraction in the fuel layer, Xbar, determined by
+    Goodwin et al. 2019 (ApJ 870, #64). The value returned is in units of MeV/nucleon:
+    q = Q_nuc(0.73)
+    5.35517578
+    There are three versions of the approximation, which can be selected by using the
+    quadratic and old_relation flags; by default the most precise (and recent) version
+    is used:
+    q = Q_nuc(0.73, quadratic=True)
+    5.758715
+    q = Q_nuc(0.73, old_relation=True)
+    4.52
+    If you want the coefficients, rather than the actual Q_nuc value, you can use the
+    coeff flag (in which case the first argument is ignored):
[…]

### 2019-08-28 doc/ms.tex
Updated paper

@@
 \newcommand{\epc}{{\rm erg\,cm^{-2}}}
+\newcommand{\Xb}{\ensuremath{\overline{X}}}
 
@@
 
-Thermonuclear bursts arise from unstable ignition of accumulated fuel on the surface of neutron stars in low-mass X-ray binaries.
-Comparison of observed bursts with numerical models is a key factor in determining the properties of the host neutron star.
+Thermonuclear (type-I) bursts arise from unstable ignition of accumulated fuel on the surface of neutron stars in low-mass X-ray binaries.
+Measurements of burst properties enable observers to infer the properties of the host neutron stars, but a number of confounding astrophysical effects contribute to systematic uncertainties.
+%
+In particular, comparison of observed burst properties with the predictions of numerical models is a 
+% key factor in determining the properties of the host neutron star.
+powerful method of constraining host properties.
 %
@@
 \section{Introduction} \label{sec:intro}
+
+Thermonuclear (type-I) bursts are observed from neutron stars accreting from low-mass ($\lesssim 1\ M_\odot$) binary companions \cite[e.g.][]{gal17b}. ...
 
@@
 %
-Provided the $F_i$ are in units of flux, or the flux can be estimated from the values,  we may measure the peak flux, $F_{\rm pk}$, and integrate to estimate the burst fluence, $E_b$.
+Provided the $F_i$ are in units of flux, or the flux can be estimated from the intensity,  we may measure the peak flux, $F_{\rm pk}$, and integrate to estimate the burst fluence, $E_b$.
 
 The peak flux $F_{\rm pk}$ provides constraints on the source distance, since the maximum burst luminosity is limited to (roughly) the Eddington luminosity \cite[e.g.][]{lew93}
-% copied from bcatalog
+% copied from bcatalog, and updated prefac
+% (8.*pi*const.G*const.m_p*1.4*const.M_sun*const.c/(const.sigma_T*1.259)).to('erg s-1')                                                         
+% Out[16]: <Quantity 2.79569695e+38 erg / s>
 \begin{eqnarray}
@@
        \nonumber \\
-  & = & 2.7\times10^{38} \left(\frac{M_{\rm NS}}{1.4M_\odot}\right)
+%  & = & 2.7\times10^{38} \left(\frac{M_{\rm NS}}{1.4M_\odot}\right)
+  & = & 2.80\times10^{38} \left(\frac{M_{\rm NS}}{1.4M_\odot}\right)
  \frac{1+(\alpha_{\rm T}T_{\rm e})^{0.86}}{(1+X)}
 \nonumber \\ & & \times\  % line break
-    \left[\frac{1+z(R)}{1.31}\right]^{-1}\ % \xi_b\
+%    \left[\frac{1+z(R)}{1.31}\right]^{-1}\ % \xi_b\
+    \left[\frac{1+z(R)}{1.259}\right]^{-1}\ % \xi_b\
               \eps
@@
 c^2)^{-1/2}$, which may be lower than the value at the NS surface, while the photosphere is expanded during the radius expansion episode (i.e. $R\ge R_{\rm NS}$).
-%
-Mention also the empirical value here...
+
+% {\it Mention also the empirical value here... } actually later
 
@@
 By equating the maximum flux of these events with the Eddington luminosity, the distance to the source can be estimated as
+% p is the prefactor from the L_Edd expression above
+% np.sqrt(p/(4.*pi*3e-8*u.erg/u.cm**2/u.s)).to('kpc')                     
+% Out[26]: <Quantity 8.82528911 kpc>
 \begin{eqnarray}
  d & = & \left(\frac{L_{\rm Edd,\infty}}{4\pi \xi_b F_{\rm pk, RE}}\right)^{1/2} \nonumber \\
-   & = & 8.6
+%   & = & 8.6
+   & = & 8.83
 	\left( \frac{\xi_b F_{\rm pk, RE}}{3\times10^{-8}\ \epcs} \right)^{-1/2}
@@
 \nonumber \\ & & \times\  % line break
-	\left[\frac{1+z(R)}{1.31}\right]^{-1/2}
+%	\left[\frac{1+z(R)}{1.31}\right]^{-1/2}
+	\left[\frac{1+z(R)}{1.259}\right]^{-1/2}
 	(1+X)^{-1/2}\ {\rm kpc}
@@
 % 
-Here the $\xi_b$ factor accounts for the possible anisotropy of the burst emission, and is defined in the same sense as \citep{fuji88}, such that $L_b=4\pi d^2\xi_b F_{\rm pk}$.
+Here the $\xi_b$ factor accounts for the possible anisotropy of the burst emission, 
+which is discussed in the next section.
+
+As is obvious from equation \ref{disteq}, several factors will contribute to statistical (and possibly systematic) uncertainties in distance estimates derived in this manner.
+First, the neutron star mass $M_{\rm NS}$ is unknown...
+%
+Second, the radius $R$ at which the redshift should be calculated may not be clear from the observations.
+In principle one could take the inferred blackbody nromalisation at the time of maximum flux, but for most PRE bursts the maximum flux is achieved close to the ``touchdown'' point, where the photosphere has returned (more or less) to the neutron star surface \cite[e.g.][]{gal06a}.
+%
+Third, the possible range/likely value for $X$...
+%
+Fourth, even if the inclination $i$ of the system is known, converting to the anisotropy factor requires modelling of the effect of the disk on the radiation field arising from the burst \cite[e.g.][]{he16}.
+
+These issues may be avoided by adopting instead the empirical Eddington luminosity measured by \cite[]{kuul03a}...
+
+\subsection{Emission anisotropy}
+\label{subsec:anisotropy}
+
+The anisotropy factor $\xi_b$ introduced into the expression for the distance (equation  \ref{disteq}) is defined in the same sense as \citep{fuji88}, such that the total luminosity of the burst $L_b=4\pi d^2\xi_b F_{\rm pk}$.
+The range of values of $\xi_b$ can thus be understood that if $\xi_b<1$ the burst flux is enhanced (i.e.
+preferentially beamed) toward our line of sight (so that $L_b$ would be {\it overestimated} were it not included in the calculation), while if $\xi_b>1$ the burst flux is suppressed.
+%
 Since the distribution of the burst flux over the neutron star is likely different from that of the persistent emission, we define a different anisotropy factor $\xi_p$ relating the observed persistent flux $F_p$ to the total luminosity.
@@
 
-As is obvious from equation \ref{disteq}, several factors will contribute to statistical (and possibly systematic) uncertainties in distance estimates derived in this manner.
-First, the neutron star mass $M_{\rm NS}$ is unknown...
-%
-Second, the radius $R$ at which the redshift should be calculated may not be clear from the observations.
-In principle one could take the inferred blackbody nromalisation at the time of maximum flux, but for most PRE bursts the maximum flux is achieved close to the ``touchdown'' point, where the photosphere has returned (more or less) to the neutron star surface \cite[e.g.][]{gal06a}.
-%
-Third, the possible range/likely value for $X$...
-%
-Fourth, even if the inclination $i$ of the system is known, converting to the anisotropy factor requires modelling of the effect of the disk on the radiation field arising from the burst \cite[e.g.][]{he16}.
-
-These issues may be avoided by adopting instead the empirical Eddington luminosity measured by \cite[]{kuul03a}...
+{\it describe how we treat this }
+
+\subsection{Burst energetics}
+\label{subsec:energetics}
 
@@
 \begin{eqnarray}
-Q_{\rm nuc} &= & 1.31+6.95\bar{X}-1.92\bar{X}^2 \nonumber \\
-& \approx & 1.35 + 6.05 \bar{X}
+Q_{\rm nuc} &= & 1.31+6.95\Xb-1.92\Xb^2 \nonumber \\
+& \approx & 1.35 + 6.05 \Xb
 \end{eqnarray}
-\cite[]{goodwin19a} where $\bar{X}$ is the average H-fraction of the fuel layer at ignition.
-
-Perhaps put in here some estimates of the fuel composition as a function of $X_0$ and recurrence time; could perhaps develop a ``prior'' for $\bar{X}$
-
-Provided an estimate of $\bar{X}$ is available, we can estimate the burst column that has ignited as
+\cite[]{goodwin19a} where $\Xb$ is the average H-fraction of the fuel layer at ignition. For fuel with solar composition (i.e. $X=0.7$), $Q_{\rm nuc}=5.22$~MeV/nucleon.
+
+{\it Perhaps put in here some estimates of the fuel composition as a function of $X_0$ and recurrence time; could perhaps develop a ``prior'' for $\Xb$ }
+
+Provided an estimate of $\Xb$ is available, we can estimate the burst column that has ignited as
+% (1e-6*u.erg/u.cm**2*(10.*u.kpc).to('cm')**2/(5.22*u.MeV/const.m_p)*1.259
+%     ...: /(11.2*u.km)**2).to('g cm-2')
+% Out[36]: <Quantity 1.9112067e+08 g / cm2>
 \begin{eqnarray}
@@
                                                             \nonumber \\
-  & = & 3.0\times10^8 \left(\frac{\xi_b E_{\rm b}}{10^{-6}\ \epc}\right)
+%  & = & 3.0\times10^8 \left(\frac{\xi_b E_{\rm b}}{10^{-6}\ \epc}\right)
+  & = & 1.91\times10^8 \left(\frac{\xi_b E_{\rm b}}{10^{-6}\ \epc}\right)
                       \left(\frac{d}{10\ {\rm kpc}}\right)^2\
-             \left(\frac{Q_{\rm nuc}}{4.4\ {\rm MeV/nucleon}}\right)^{-1}
+%             \left(\frac{Q_{\rm nuc}}{4.4\ {\rm MeV/nucleon}}\right)^{-1}
+             \left(\frac{Q_{\rm nuc}}{5.22\ {\rm MeV/nucleon}}\right)^{-1}
 \nonumber \\ & & \times\  % line break
-                      \left(\frac{1+z}{1.31}\right)
-                      \left(\frac{R_{\rm NS}}{10\ {\rm km}}\right)^{-2}
+%                      \left(\frac{1+z}{1.31}\right)
+%                      \left(\frac{R_{\rm NS}}{10\ {\rm km}}\right)^{-2}
+                      \left(\frac{1+z}{1.259}\right)
+                      \left(\frac{R_{\rm NS}}{11.2\ {\rm km}}\right)^{-2}
                       {\rm g\,cm^{-2}}
@@
 \end{eqnarray}
-{\it need to update prefactor, denominator in $Q_{\rm nuc}$ term here }
+% {\it need to update prefactor, denominator in $Q_{\rm nuc}$ term here } - done
 
@@
 
+{\it This likely duplicates earlier discussion} \\
+%
 An additional correction that must be applied is the effect of the anisotropic emission of the burst, based on the system inclination $i$.
@@
 
-Here we describe three different approaches that have been used to deduce system parameters from observations of thermonuclear bursts.
-
-\subsection{Inferring fuel composition} 
+Here we describe different approaches that have been used to deduce system parameters from observations of thermonuclear bursts.
+%
+We focus on adopting a suitable approach given the availability of data, which will differ for different sources
+
+\subsection{Zero bursts -- constraining the distance}
+\label{subsec:zerobursts}
+
+While it may seem strange to base an analysis on the {\it non-}detection of bursts, practically even for the best-studied sources the X-ray observation duty cycle is of order a few percent.
+There is a high probability that bursts will be missed, particularly if the accretion rate is low (and hence the bursts are infrequent).
+In extreme cases this can mean that no bursts whatsoever are observed, which may instead be explained by the compact object being a black hole rather than a neutron star.
+
+In cases where we can be confident that the compact object is a neutron star (e.g.
+where persistent pulsations are detected) but no bursts are observed, we can constrain the distance by adopting a composition for the fuel and comparing the predictions of burst models to the good-time intervals of our X-ray data.
+
+For example, based on {\it RXTE}/PCA observations of the accretion-powered millisecond pulsar IGR~J00291+5934, \cite{gal06b} derived joint constraints over the distance $d$ and fuel H-fraction $X_0$, adopting  the predictions of a simple numerical model ({\it ref to {\sc settle}}).
+These constraints suggest the distance is $\gtrsim5$~kpc (at 3-sigma significance), provided $X_0\approx0.7$, a reasonable choice given the expectation of a H-rich donor in this $2.46$~hr binary orbital period system \cite[]{gal05a}.
+
+In a subsequent outburst in 2015, a single burst was detected by {\it Swift}/XRT \cite[]{kuin15}, offering the opportunity to verify the previously-determined distance limit. The estimated peak 
+% 0.3--10~keV flux in the range (3.0--$5.6)\times10^{-11}\ \epcs$. 
+0.1--35~keV flux was $(18\pm4)\times10^{-8}\ \epcs$ \cite[]{defalco17}, with the burst exhibiting spectral variations indicative of PRE.
+By comparing the peak flux to the empirical Eddington luminosity, the corresponding distance
+\cite[including the inclination range of 22--32$^\circ$ suggested by][]{torres08},
+% possible inclinations as described in sec \S\ref{subsec:data}) is $4.4_{-0.7}^{+0.8}$~kpc, which is mostly inconsistent with the earlier limit. Clearly, such limits must be viewed with some caution.
+is $5.0_{-0.5}^{+0.7}$~kpc, which is fully consistent with the previously established limit. The effect of the inclination constraints is illustrated in Fig. \ref{fig:dist_igr00291}).
+
+\begin{figure}[ht!]
+\plotone{fig1.pdf}
+\caption{Estimated distance distributions for the accretion-powered X-ray pulsar IGR~J00291+5934, based on the peak flux of the sole thermonuclear burst observed by \cite{defalco17}.
+%
+The distribution for an isotropic distribution of system inclinations (up to a maximum value of $75^\circ$, motivated by the lack of dips in the X-ray intensity) is shown as the blue shaded histogram.
+Imposing the inclination constraint of 22--32$^\circ$ suggested by \cite{torres08}, for which the predicted burst anisotropy factor is $\xi_b=0.704$ on average, gives instead the orange-shaded histogram.
+The resulting $1\sigma$ confidence interval is $5.0_{-0.5}^{+0.7}$~kpc.
+\label{fig:dist_igr00291}}
+\end{figure}
+
+\subsection{One burst -- estimating the burst rate} 
+\label{subsec:burstrate}
+
+{\it copy the analysis for the IGR source here }
+
+\subsection{Two bursts -- inferring fuel composition} 
 \label{subsec:fuelcomp}
 
-Given a set of burst observations as described in \S\ref{subsec:data}, a common approach \cite[e.g][]{falanga11} is to estimate the fuel composition at ignition, and hence the accreted fuel composition, based on simple analytic estimates of the burst energy production $Q_{\rm nuc}$.
-
-One issue with this approach is that the effects of anisotropy are frequently neglected.
+Given a minimally complete set of burst observations as described in \S\ref{subsec:data}, a common approach \cite[e.g][]{falanga11} is to estimate the fuel composition at ignition, and hence the accreted fuel composition, based on simple analytic estimates of the burst energy production $Q_{\rm nuc}$.
+
+One issue with this approach (as for analyses based on one burst, as described in \S\ref{subsec:burstrate}) is that the effects of anisotropy are frequently neglected.
 %

### 2020-05-18 doc/ms.tex
Split off X_0 function from hfrac, for use with the new notebook

@@
 %% Beginning of file 'ms.tex'
-%%
 %% Copied from sample61.tex by dkg 15.3.2018
@@
 %% See https://bitbucket.org/minbar/concord
-%%
+%% and the companion jupyter notebook, Inferring burster properties.ipynb
+
 %% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
@@
 %%
-%% using aastex version 6.1
-\documentclass{aastex61}
+%% using aastex version 6.3 from 2020 May; originally 6.1
+\documentclass{aastex63}
 
@@
 \newcommand{\Xb}{\ensuremath{\overline{X}}}
+\newcommand{\qnuc}{Q_{\rm nuc}}
+\newcommand{\yign}{y_{\rm ign}}
+\newcommand{\zcno}{Z_{\rm CNO}}
+\newcommand{\xte}{{\it RXTE}}
+\newcommand{\sax}{{\it BeppoSAX}}
+\newcommand{\igr}{{\it INTEGRAL}}
+
+\newcommand{\todo}[1]{\textbf{\textcolor{red}{TODO: #1}}} % Duncan
+
+% standard redshift to use
+% 1./sqrt(1.-2.*c.G*1.4*c.M_sun/((11.2*u.km).to('m')*c.c**2))
+
+\newcommand{\opz}{1.259}
 
@@
 
+% ------------------------------------------------------------------------------
 %% From the front matter, we move on to the body of the paper.
@@
 %
-In \S\ref{sec:data} we describe the typical characteristics of burst observations along with those properties that can be predicted by numerical models.
-% 
-In \S\ref{sec:analysis} we explore a number of methods to infer properties of the bursting source, and assess their accuracy.
+In \S\ref{sec:data} we describe the typical characteristics of burst observations along with those properties that can be 
+% predicted by numerical models.
+inferred from simple calculations or by comparison with numerical models.
+% 
+In \S\ref{sec:analysis} we explore a number of methods to infer properties of the bursting source, based on the extent of available observational data, and assess their accuracy.
 %
@@
 
+% ------------------------------------------------------------------------------
 \section{Observations and modelling} 
@@
 
-\subsection{Burst observations}
-\label{subsec:data}
+%\subsection{Burst observations}
+%\label{subsec:data}
 
 The fundamental observable that we measure is the time-history of the burst flux, the lightcurve $(t_i,F_{i})$.
-Here the $F_{i}$ may be in instrumental units of count~s$^{-1}$, or may be estimates of the bolometric flux at time $t_i$, derived from model fits of the spectrum at time $t_i$.
+Here the $F_{i}$ may be in instrumental units of count~s$^{-1}$ (or perhaps count~s$^{-1}$~cm$^{-2}$), or may be estimates of the bolometric flux at time $t_i$, derived from model fits of the spectrum at time $t_i$.
 %
@@
 
-The peak flux $F_{\rm pk}$ provides constraints on the source distance, since the maximum burst luminosity is limited to (roughly) the Eddington luminosity \cite[e.g.][]{lew93}
+Given one or more bursts, we can derive constraints on the recurrence time $\Delta t$.
+A pair of bursts separated by an interval with complete coverage by X-ray instruments offer an unambiguous measurement of the recurrence time, but such measurements are rare due to the typically low duty cycle for most instruments \cite[e.g.][]{minbar}.
+%
+Such cases also tend to be biased towards lower recurrence times, which may be dominated by so-called ``short waiting time'' bursts, which likely arise from ignition of unburnt fuel from the previous event \cite[]{keek10}.
+The recurrence time for such events likely does not represent the actual time required to meet the ignition conditions at the base of the fuel layer, and so are generally excluded for the purposes of inferring properties of the burst sources.
+
+Where one or more bursts $N$ are observed in low-duty cycle observations, the rate can be estimated based on the total exposure, i.e.
+\begin{equation}
+\langle R\rangle = \frac{N}{\sum^n T_i}
+\end{equation}
+where each of $n$ observartions has exposure $T_i$.
+The probable range of the average rate can be estimated assuming Poisson counting statistics for the uncertainty on the burst number $N$.
+Because of the approximately periodic behavior common for bursts, and the similarity between low-Earth orbital periods and the burst recurrence times, such uncertainty estimates may be wildly incorrect.
+
+From data taken prior to the burst ignition we can measure the persistent flux $F_{\rm per}$, a measure of the accretion rate; 
+% and provided more than one burst is observed, we may also measure (or estimate) the recurrence time, $\Delta t$. 
+%
+these parameters may be combined to estimate the $\alpha$-value, the (observed) ratio of the burst to persistent flux: 
+\begin{equation}
+\alpha = \frac{\Delta t F_{\rm per} c_{\rm bol}}{E_B}
+\end{equation}
+where $c_{\rm bol}$ is the bolometric correction giving the inverse fraction of total persistent flux emitted in the instrumental band.
+
+additional data analysis comments
+\begin{itemize}
+\item correcting lightcurve for variations e.g. from dips, or achromatic variations
+\item using different spectral models than blackbody
+\item allowing the persistent flux to vary
+\item maybe a table with the parameters?
+\end{itemize}
+
+\subsection{Burster distances}
+\label{subsec:distance}
+
+A significant fraction of observed bursts exhibit characteristic variations in their spectral hardness (or the blackbody temperature and normalisation) around their maximum that indicate the presence of photospheric radius-expansion (PRE).
+%
+Even if such variations are not present, 
+%
+the peak flux $F_{\rm pk}$ provides constraints on the source distance, since the maximum burst luminosity is limited to (roughly) the Eddington luminosity \cite[e.g.][]{lew93}:
 % copied from bcatalog, and updated prefac
@@
 %    \left[\frac{1+z(R)}{1.31}\right]^{-1}\ % \xi_b\
-    \left[\frac{1+z(R)}{1.259}\right]^{-1}\ % \xi_b\
+    \left[\frac{1+z(R)}{\opz}\right]^{-1}\ % \xi_b\
               \eps
@@
 at the photosphere $1+z(R)=(1-2GM_{\rm NS}/R
-c^2)^{-1/2}$, which may be lower than the value at the NS surface, while the photosphere is expanded during the radius expansion episode (i.e. $R\ge R_{\rm NS}$).
+c^2)^{-1/2}=1.259$ for $M_{\rm NS}=1.4\ M_\odot$ and $R=R_{\rm NS}=11.2$~km.
+%
+The effective redshift measured at the peak of a PRE burst may be lower than the value at the NS surface, while the photosphere is expanded during the radius expansion episode (i.e.
+$R\ge R_{\rm NS}$).
 
@@
 
-A significant fraction of observed bursts exhibit characteristic variations in their spectral hardness (or the blackbody temperature and normalisation) around their maximum that indicate the presence of photospheric radius-expansion (PRE).
 By equating the maximum flux of these events with the Eddington luminosity, the distance to the source can be estimated as
@@
 %	\left[\frac{1+z(R)}{1.31}\right]^{-1/2}
-	\left[\frac{1+z(R)}{1.259}\right]^{-1/2}
+	\left[\frac{1+z(R)}{\opz}\right]^{-1/2}
 	(1+X)^{-1/2}\ {\rm kpc}
@@
 Second, the radius $R$ at which the redshift should be calculated may not be clear from the observations.
-In principle one could take the inferred blackbody nromalisation at the time of maximum flux, but for most PRE bursts the maximum flux is achieved close to the ``touchdown'' point, where the photosphere has returned (more or less) to the neutron star surface \cite[e.g.][]{gal06a}.
+In principle one could take the inferred blackbody normalisation at the time of maximum flux, but for most PRE bursts the maximum flux is achieved close to the ``touchdown'' point, where the photosphere has returned (more or less) to the neutron star surface \cite[e.g.][]{gal06a}.
 %
@@
 The fluence (and the shape of the burst profile) is determined by the amount of accumulated fuel, and its composition.
-The nuclear energy generation rate $Q_{\rm nuc}$ depends on the composition; studies with the 1-D numerical code {\sc kepler} suggest that
+For accretion of mixed H/He with hydrogen mass fraction $X_0$, the composition at ignition may be modified substantially by $\beta$-limited CNO burning.
+In extreme cases the recurrence time is sufficiently long that the accreted hydrogen at the base of the fuel layer is exhausted, and ignition of intense, fast He-rich bursts occur.
+
+For a fuel layer consisting of mixed H/He, $\qnuc$ depends on the mean hydrogen fraction at ignition, \Xb:
+% The nuclear energy generation rate $\qnuc$ depends on the composition; 
+studies with the 1-D numerical code {\sc kepler} suggest that 
 \begin{eqnarray}
@@
 & \approx & 1.35 + 6.05 \Xb
+\label{eq:qnuc}
 \end{eqnarray}
-\cite[]{goodwin19a} where $\Xb$ is the average H-fraction of the fuel layer at ignition. For fuel with solar composition (i.e. $X=0.7$), $Q_{\rm nuc}=5.22$~MeV/nucleon.
+\cite[]{goodwin19a}.
+% where $\Xb$ is the average H-fraction of the fuel layer at ignition. 
+For fuel with solar composition (i.e. $X=0.7$), $Q_{\rm nuc}=5.22$~MeV/nucleon.
 
@@
 %                      \left(\frac{R_{\rm NS}}{10\ {\rm km}}\right)^{-2}
-                      \left(\frac{1+z}{1.259}\right)
+                      \left(\frac{1+z}{\opz}\right)
                       \left(\frac{R_{\rm NS}}{11.2\ {\rm km}}\right)^{-2}
                       {\rm g\,cm^{-2}}
-\label{column}
+\label{eq:column}
 \end{eqnarray}
 % {\it need to update prefactor, denominator in $Q_{\rm nuc}$ term here } - done
-
-From data taken prior to the burst ignition we can measure the persistent flux $F_{\rm per}$, a measure of the accretion rate; and provided more than one burst is observed, we may also measure (or estimate) the recurrence time, $\Delta t$.
-
-These parameters may be combined to estimate the $\alpha$-value, the ratio of the burst to persistent flux: 
-\begin{equation}
-\alpha = \frac{\Delta t F_{\rm per} c_{\rm bol}}{E_B}
-\end{equation}
-where $c_{\rm bol}$ is the bolometric correction giving the inverse fraction of total persistent flux emitted in the instrumental band.
 
@@
 These constraints suggest the distance is $\gtrsim5$~kpc (at 3-sigma significance), provided $X_0\approx0.7$, a reasonable choice given the expectation of a H-rich donor in this $2.46$~hr binary orbital period system \cite[]{gal05a}.
+\todo{give more detail about how this was done}
 
@@
 0.1--35~keV flux was $(18\pm4)\times10^{-8}\ \epcs$ \cite[]{defalco17}, with the burst exhibiting spectral variations indicative of PRE.
-By comparing the peak flux to the empirical Eddington luminosity, the corresponding distance
+By comparing the peak flux to the empirical Eddington luminosity (see \S\ref{subsec:1burst}, below), the corresponding distance
 \cite[including the inclination range of 22--32$^\circ$ suggested by][]{torres08},
@@
 \subsection{One burst -- estimating the burst rate} 
-\label{subsec:burstrate}
-
-{\it copy the analysis for the IGR source here }
-
+\label{subsec:1burst}
+
+% {\it copy the analysis for the IGR source here }
+
+For a single burst detected from a source, we can expect to measure the burst peak flux, fluence, and timescale, and infer the distance and hence the expected burst rate.
+The detailed shape of the lightcurve can provide constraints from comparisons with numerical models (see \S\ref{subsec:lccompare}), but here we focus on the more straightforward calculation which may be done from the simple burst measurements alone.
+
+We use for example the sole burst detected with \igr/JEM-X from IGR~J17591$-$2342 \cite[]{kuiper20}.
+This burst was observed  when the estimated persistent bolometric flux was $(1.2\pm0.2)\times10^{-9}\ \epcs$, and exhibited a fluence of $E_b =(1.1\pm0.1)\times10^{-6}\ {\rm erg\,cm^{-2}}$.
+The peak flux was $(7.6\pm1.4)\times10^{-6}\ \epcs$.
+
+\cite{kuiper20} first estimated the distance to the source based on the empirical value of the Eddington luminosity, and incorporating the expected anisotropy for an inclination range of 
+% kuiper20 uses slightly more restrictive 28-30
+24--30$^\circ$ \cite[]{sanna18} as $7.7_{-0.6}^{+0.8}$~kpc.
+Here the uncertainty is dominated by the peak flux, since the anisotropy factor is effectively fixed for such a narrow range of inclination, at $\xi_b=0.70$.
+
+\todo{potentially also demonstrate here the effect of inclination}
+
+We can use this distance constraint to estimate in turn the accretion and burst rate, as follows.
+%
+The persistent flux can then be converted to an accretion rate, given the estimated anisotropy factor predicted (separately) for the persistent emission, as
+\begin{eqnarray}
+\dot{m} & = L_{\rm pers} (1+z) (4\pi R^2(GM/R_{\rm NS}))^{-1}\\
+& = \ldots
+% TODO: give expanded expression for this and move to earlier section
+\end{eqnarray}
+via the {\tt mdot} method, and incorporating the estimated distance distribution already determined.
+The estimated value is within the range $(2.0_{-0.4}^{+0.6})\times10^3\ \rm{g\,cm^{-2}\,s^{-1}}$ with the usual assumed range for neutron star mass and radius (and hence redshift).
+
+In parallel, we can estimate the ignition column using equation \ref{eq:column}, given an estimate of the nuclear burning yield, $\qnuc$.
+; as before, we could potentially obtain an estimate of this quantity by comparison of the observed lightcurve to simulations.
+However, in the absence of an estimate of this quantity, all is not lost.
+
+Instead we can work backwards, 
+% 
+and assume a set of randomly generated values of \Xb\
+to estimate $\yign$ from equation \ref{column}, and $\Delta t_{\rm rec} = \yign/\dot{m}$.
+%
+We choose a uniform distribution for \Xb\ in the range 0--0.7; this distribution is probably not realistic for a wide sample of bursts, but it is difficult to do better.
+% text copied from Kuiper et al. 2020 down to ********
+% not yet edited (need to modify sufficiently to avoid plagiarism issues
+Given each pair of (\Xb,  $\Delta t_{\rm rec}$) we can infer the accreted H-fraction $X_0$, based on 
+% the expectation that hydrogen burns via the hot-CNO cycle and will be exhausted in a time $t_{\rm CNO} =9.8(X_0/0.7)(Z_{\rm CNO}/0.02)^{-1}$ \citep[e.g.][]{lampe16}.
+equation \ref{eq:X_0}.
+%
+Since $Z_{\rm CNO}$ is also unknown, we similarly draw random values from a uniform distribution in the range 0.0--0.02.
+We then select only those $X_0$ values within a physically realistic range $X_0<0.75$, and compute the confidence limits on the remaining parameters.
+
+For the single burst observed from IGR~J17591$-$2342 the inferred ignition column is $2.2_{-0.8}^{+1.4}\times10^8\,{\rm g\,cm^{-2}}$, and the average expected recurrence time is $1.8\pm0.7$~d. 
+% Such long recurrence times likely guarantee that the hydrogen in the accreted fuel is exhausted at the base prior to ignition, and explaining the fast rise and short duration of the burst. 
+We can also infer lower (upper) limits on $X_0$ ($Z_{\rm CNO}$), although these limits are not strongly constraining; we find $X_0>0.17$ and $Z_{\rm  CNO}<0.017$ at 95\% confidence.
+% *******
+
+% Figure produced from the notebook; used 100000 samples for a smoother set of 
+% distributions
+\begin{figure}[ht!]
+\plotone{igrJ17591-2342.pdf}
+\caption{Inferred distributions of burst parameters for the single event observed from IGR~J17591$-$2342 with \igr/JEM-X, as reported by \cite{kuiper20}.
+An initially uniform distribution of \Xb, the mean H fraction at ignition, is used to estimate the burst ignition column $\yign$\ from the measured fluence, $E_b$.
+The column $\yign$\ and accretion rate are then used to estimate the recurrence time $\Delta t$ and the accreted H-fraction $X_0$, and we select only the physically realistic values $X_0<0.75$.
+The resulting constraints on the fuel composition parameters $X_0$ and $\zcno$ are rather weak, but the recurrence time is somewhat better constrained.
+\label{fig:param_igr17591}}
+\end{figure}
+
+% ------------------------------------------------------------------------------
 \subsection{Two bursts -- inferring fuel composition} 
@@
 
-One issue with this approach (as for analyses based on one burst, as described in \S\ref{subsec:burstrate}) is that the effects of anisotropy are frequently neglected.
+One issue with this approach (as for analyses based on one burst, as described in \S\ref{subsec:1burst}) is that the effects of anisotropy are frequently neglected.
 %
@@
         \end{cases}
+    \label{eq:X_0}
 \end{equation}

### 2020-05-18 concord/utils.py

@@ -155,6 +155,57 @@ def Q_nuc(Xbar, quadratic=False, old_relation=False, coeff=False):
 
 # ------- --------- --------- --------- --------- --------- --------- ---------
 
+def X_0(xbar, zcno, tdel, opz=1.259, debug=False, old_relation=False):
+    '''
+    Routine (extracted from hfrac) to determine the accreted H fraction X_0
+    given the average H-fraction at ignition, xbar, the CNO metallicity zcno,
+    and the burst recurrence time, tdel
+    '''
+
+    # This parameter sets the prefactor for the time to burn all H via hot-CNO; see
+    # Lampe et al. (2016, ApJ 819, 46)
+
+    tpref = 9.8 * u.hr
+
+    # Here we allow an option to use the old version if need be; this is already passed to the
+    # Q_nuc function, so we'll get the old parameters from that if need be
+
+    if old_relation:
+        tpref = 11. * u.hr
+
+    # Also here now we distinguish between the two possible cases, the first where there
+    # remains hydrogen at the base, and the second where it is exhausted
+    # This division is quantified via the f_burn fraction, which initially at least we
+    # must estimate
+
+    t_CNO = tpref * (xbar / 0.7) / (zcno / 0.02)
+    f_burn = tdel / (opz * t_CNO)
+    f_burn_prev = 1. / f_burn  # dummy value first off to ensure the loop gets called
+
+    # Loop here to make sure we have a consistent choice
+
+    while (1. - f_burn) * (1. - f_burn_prev) < 0.:
+
+        if (f_burn <= 1.):
+            # still hydrogen at the base
+            x_0 = xbar + 0.35 * (tdel / (opz * tpref)) * (zcno / 0.02)
+        else:
+            # hydrogen exhausted at the base
+            x_0 = np.sqrt(1.4 * xbar * tdel / (opz * tpref) * (zcno / 0.02))
+
+        f_burn_prev = f_burn
+        t_CNO = tpref * (x_0 / 0.7) / (zcno / 0.02)
+        f_burn = tdel / (opz * t_CNO)
+
+        if debug:
+            print(xbar, x_0, t_CNO, xi_b, xi_p)
+            print('flipping f_burn {} -> {}'.format(f_burn_prev, f_burn))
+
+    #        print (i,xbar,t_CNO,f_burn,x_0,xi_b,xi_p)
+    return x_0
+
[…]

### 2021-09-14 doc/ms.tex
Added homogenize_params to help ensure compatible parameter types
homogenize_params checks the size of all the passed parameters, and if any of them are a distribution, will convert the others to distributions of the same size (by calling value_to_dist). So far the alpha and mdot methods have been converted to use this routine, and the others should also be converted.

Also added logging (in utils.py) via the logging package, to make the informational and error messages look nicer

@@
 Thermonuclear (type-I) bursts arise from unstable ignition of accumulated fuel on the surface of neutron stars in low-mass X-ray binaries.
-Measurements of burst properties enable observers to infer the properties of the host neutron stars, but a number of confounding astrophysical effects contribute to systematic uncertainties.
-%
-In particular, comparison of observed burst properties with the predictions of numerical models is a 
-% key factor in determining the properties of the host neutron star.
-powerful method of constraining host properties.
-%
-Here we describe some commonly-used approaches for determining system parameters, including composition of the burst fuel, and introduce a new suite of software tools, {\sc concord}, intended to make comprehensive model-observation comparisons practical.
-
+Measurements of burst properties enable observers to infer the properties of the host neutron star and mass donors, but a number of confounding astrophysical effects contribute to systematic uncertainties.
+Here we describe some commonly-used approaches for determining system parameters, including composition of the burst fuel, and introduce a new suite of software tools, {\sc concord}, intended to fully account for astrophysical uncertainties.
+Furthermore, comparison of observed burst properties with the predictions of numerical models is a powerful method of constraining host properties, and the tools presented here are intended to make comprehensive model-observation comparisons straightforward.
+When combined with the extensive samples of burst observations accumulated by X-ray observatories, these software tools will provide a step-change in the amount of information that can be inferred about typical burst sources.
 \end{abstract}
@@
 % predicted by numerical models.
-inferred from simple calculations or by comparison with numerical models.
+inferred from simple calculations. % or by comparison with numerical models.
 % 
@@
 % ------------------------------------------------------------------------------
-\section{Observations and modelling} 
+\section{Observations and burster properties} 
 \label{sec:data}
 
+% describe the typical characteristics of burst observations along with those 
+% properties that can be inferred from simple calculations
+
 %\subsection{Burst observations}
-%\label{subsec:data}
+\label{subsec:data}
 
 The fundamental observable that we measure is the time-history of the burst flux, the lightcurve $(t_i,F_{i})$.
-Here the $F_{i}$ may be in instrumental units of count~s$^{-1}$ (or perhaps count~s$^{-1}$~cm$^{-2}$), or may be estimates of the bolometric flux at time $t_i$, derived from model fits of the spectrum at time $t_i$.
-%
+Here the $F_{i}$ may be in instrumental units of count~s$^{-1}$ (or perhaps count~s$^{-1}$~cm$^{-2}$), or may be estimates of the bolometric flux at time $t_i$, derived from spectral model fits of the spectrum at time $t_i$.
+%
+The net (excluding the persistent emission) burst spectra generally can be well fit with a blackbody model, and it is thought such models can provide a reasonable estimate of the bolometric flux [{\bf ref}]
+
+[{\bf insert table with key burst parameters here? }]
+
 Provided the $F_i$ are in units of flux, or the flux can be estimated from the intensity,  we may measure the peak flux, $F_{\rm pk}$, and integrate to estimate the burst fluence, $E_b$. 
+%
+Where the bursts are very short, or observed with low signal-to-noise, a spectrum extracted over an interval covering the entire event may substitute for a fluence measurement.
 
 Given one or more bursts, we can derive constraints on the recurrence time $\Delta t$.
-A pair of bursts separated by an interval with complete coverage by X-ray instruments offer an unambiguous measurement of the recurrence time, but such measurements are rare due to the typically low duty cycle for most instruments \cite[e.g.][]{minbar}.
+A pair of bursts separated by an interval with uninterrupted coverage by X-ray instruments offer an unambiguous measurement of the recurrence time, but such measurements are rare due to the typically low duty cycle for most instruments \cite[e.g.][]{minbar}.
 %
@@
 
-Where one or more bursts $N$ are observed in low-duty cycle observations, the rate can be estimated based on the total exposure, i.e.
+Where one or more bursts $N$ are observed in $n$ low-duty cycle observations with exposure $T_i$, the rate can be estimated based on the total exposure, i.e.
 \begin{equation}
@@
 \end{equation}
-where each of $n$ observartions has exposure $T_i$.
+% where each of $n$ observartions has exposure $T_i$. 
 The probable range of the average rate can be estimated assuming Poisson counting statistics for the uncertainty on the burst number $N$.
@@
 
-From data taken prior to the burst ignition we can measure the persistent flux $F_{\rm per}$, a measure of the accretion rate; 
-% and provided more than one burst is observed, we may also measure (or estimate) the recurrence time, $\Delta t$. 
-%
-these parameters may be combined to estimate the $\alpha$-value, the (observed) ratio of the burst to persistent flux: 
-\begin{equation}
-\alpha = \frac{\Delta t F_{\rm per} c_{\rm bol}}{E_B}
-\end{equation}
-where $c_{\rm bol}$ is the bolometric correction giving the inverse fraction of total persistent flux emitted in the instrumental band.
-
-additional data analysis comments
-\begin{itemize}
-\item correcting lightcurve for variations e.g. from dips, or achromatic variations
-\item using different spectral models than blackbody
-\item allowing the persistent flux to vary
-\item maybe a table with the parameters?
-\end{itemize}
+One of the most significant astrophysical uncertainties that has largely been neglected to date is the anisotropy of the burst (and persistent) emission.
+% [{\bf move discussion from later section to here?}]
+% \subsection{Emission anisotropy}
+% \label{subsec:anisotropy}
+[{\bf bit more description here }]
+
+The anisotropy factor $\xi_b$ introduced into the expression for the distance (equation  \ref{disteq}) is defined in the same sense as \citep{fuji88}, such that the total luminosity of the burst $L_b=4\pi d^2\xi_b F_{\rm pk}$.
+The range of values of $\xi_b$ can thus be understood that if $\xi_b<1$ the burst flux is enhanced (i.e.
+preferentially beamed) toward our line of sight (so that $L_b$ would be {\it overestimated} were it not included in the calculation), while if $\xi_b>1$ the burst flux is suppressed.
+%
+Since the distribution of the burst flux over the neutron star is likely different from that of the persistent emission, we define a different anisotropy factor $\xi_p$ relating the observed persistent flux $F_p$ to the total luminosity.
+Modelling \cite[]{he16}...
+
+{\it describe how we treat this }
 
@@
 
-A significant fraction of observed bursts exhibit characteristic variations in their spectral hardness (or the blackbody temperature and normalisation) around their maximum that indicate the presence of photospheric radius-expansion (PRE).
+A significant fraction of observed bursts exhibit characteristic variations in their spectral hardness (or the blackbody temperature and normalisation) around their maximum that indicate the presence of photospheric radius-expansion (PRE) [{\bf ref}].
 %
@@
 
-\subsection{Emission anisotropy}
-\label{subsec:anisotropy}
-
-The anisotropy factor $\xi_b$ introduced into the expression for the distance (equation  \ref{disteq}) is defined in the same sense as \citep{fuji88}, such that the total luminosity of the burst $L_b=4\pi d^2\xi_b F_{\rm pk}$.
-The range of values of $\xi_b$ can thus be understood that if $\xi_b<1$ the burst flux is enhanced (i.e.
-preferentially beamed) toward our line of sight (so that $L_b$ would be {\it overestimated} were it not included in the calculation), while if $\xi_b>1$ the burst flux is suppressed.
-%
-Since the distribution of the burst flux over the neutron star is likely different from that of the persistent emission, we define a different anisotropy factor $\xi_p$ relating the observed persistent flux $F_p$ to the total luminosity.
-Modelling \cite[]{he16}...
-
-{\it describe how we treat this }
-
 \subsection{Burst energetics}
 \label{subsec:energetics}
+
+% \subsection{Burst energetics}
+\label{subsec:energy}
 
@@
 In extreme cases the recurrence time is sufficiently long that the accreted hydrogen at the base of the fuel layer is exhausted, and ignition of intense, fast He-rich bursts occur.
+
+From data taken prior to the burst ignition we can measure the persistent flux $F_{\rm per}$, a measure of the accretion rate; [{\bf add expression here}]
+% and provided more than one burst is observed, we may also measure (or estimate) the recurrence time, $\Delta t$. 
+the persistent spectra of burst sources are  more complex than the bursts themselves, and typically exhibit thermal as well as non-thermal components [{\bf ref}].
+%
+These parameters may be combined to estimate the $\alpha$-value, the (observed) ratio of the burst to persistent flux: 
+\begin{equation}
+\alpha = \frac{\Delta t F_{\rm per} c_{\rm bol}}{E_B}
+\end{equation}
+where $c_{\rm bol}$ is the bolometric correction giving the inverse fraction of total persistent flux emitted in the instrumental band.
+%
+The $\alpha$ parameter is an important diagnostic of the burst fuel and ignition conditions.
+%
+[{\bf mention the average method, as used in the rates paper }]
+
+Given the measured $\alpha$-value, the nuclear energy generation rate $Q_{\rm nuc}$ can be estimated, as follows:
+\begin{eqnarray}
+\alpha & = & \frac{\Delta t F_{\rm per} c_{\rm bol}}{E_B} \\
+& = & \frac{Q_{\rm grav}}{Q_{\rm nuc}}\frac{\xi_b}{\xi_p}(1+z)
+\end{eqnarray}
+where $Q_{\rm grav} = c^2z/(1+z) \approx GM_{\rm NS}/R_{\rm NS}$ is the gravitational energy release per gram.
+Note the dependence of the observational $\alpha$-value on both the surface redshift and the anisotropy parameters.
+The nuclear energy generation rate is thus
+\begin{equation}
+Q_{\rm nuc} = \frac{c^2 z}{\alpha}\frac{\xi_b}{\xi_p} \label{qnuc}
+\end{equation}
+Thus, given an estimate of $z$ and the inclination, the burst measurements can be used to infer $Q_{\rm nuc}$.
+
+Given an estimate of $Q_{\rm nuc}$, the average hydrogen fraction $\langle X\rangle$ in the fuel layer at ignition has been estimated to date using the relation $Q_{\rm nuc}=1.6 + 4\langle X\rangle$~MeV/nucleon \cite[e.g.][and references therein]{gal03d}.
+This expression includes $\approx35$\% losses
+attributed to neutrino emission \cite[]{fuji87}.
+However, the 35\% value applies only to $\beta$-decays in the much more extensive {\it rp}-process burning chain, and recent {\sc kepler} simulations indicate instead an approximation of
+% $Q{\rm nuc}=0.96+6.6\langle X\rangle$ 
+$Q{\rm nuc}=1.35+6.05\langle X\rangle$ 
+% (Goodwin et al. 2018, in preparation). 
+\cite[]{goodwin19a}.
+[{\it comment on improved quadratic expression}]
+%
+The average H-fraction at ignition $\langle X\rangle$ is a lower limit for the H-fraction in the accreted fuel, $X_0$, as for most sources the H-fraction will be reduced by steady burning prior to burst ignition.
+%
+The recurrence time for the burst can further be used to estimate $X_0$, although the burning rate also depends on the CNO mass fraction, $Z_{\rm CNO}$.
+
+additional data analysis comments
+\begin{itemize}
+\item correcting lightcurve for variations e.g. from dips, or achromatic variations
+\item using different spectral models than blackbody
+\item allowing the persistent flux to vary
+\item maybe a table with the parameters?
+\end{itemize}
 
@@
 %
+\todo{if I just use Adelle's data I don't need to mention the lightcurves at all}
+
 We seek to calculate the burst lightcurve that would be observed given a model lightcurve and neutron star parameters including distance $d$; inclination $i$; and surface redshift $(1+z)_{\rm fit}$ (so labeled to distinguish from the corresponding value implict to the model).
@@
 
+\subsection{Code architecture}
+\label{subsec:code}
+
+Describe how the code is written and how it works, the astropy.distribution thingy, MCMC for error determination etc.
+
+% ------------------------------------------------------------------------------
 \section{Analysis} 
@@
 
+% ... explore a number of methods to infer properties of the bursting source,  
+% based on the extent of available observational data, and assess their accuracy.
+
+% seems to be a newer/alternative version of sec:results?
+
 Here we describe different approaches that have been used to deduce system parameters from observations of thermonuclear bursts.
 %
-We focus on adopting a suitable approach given the availability of data, which will differ for different sources
+We focus on adopting a suitable approach given the availability of data, which will differ for different sources. 
+%
+We also describe how the {\sc concord} code can be used to fully account for the astrophysical uncertainties, and as a companion to this section (and the code itself) we provide a companion {\sc jupyter} notebook [{\bf ref}] which replicates the analyses described below.
 
@@
 
-While it may seem strange to base an analysis on the {\it non-}detection of bursts, practically even for the best-studied sources the X-ray observation duty cycle is of order a few percent.
+While it may seem strange to base an analysis on the {\it non-}detection of bursts, practically even for the best-studied sources the X-ray observation duty cycle is of order a few percent \cite[e.g.][]{minbar}.
 There is a high probability that bursts will be missed, particularly if the accretion rate is low (and hence the bursts are infrequent).
-In extreme cases this can mean that no bursts whatsoever are observed, which may instead be explained by the compact object being a black hole rather than a neutron star.
+In extreme cases this can mean that no bursts whatsoever are observed,  which may also be explained by the compact object being a black hole rather than a neutron star.
 
@@
 For example, based on {\it RXTE}/PCA observations of the accretion-powered millisecond pulsar IGR~J00291+5934, \cite{gal06b} derived joint constraints over the distance $d$ and fuel H-fraction $X_0$, adopting  the predictions of a simple numerical model ({\it ref to {\sc settle}}).
-These constraints suggest the distance is $\gtrsim5$~kpc (at 3-sigma significance), provided $X_0\approx0.7$, a reasonable choice given the expectation of a H-rich donor in this $2.46$~hr binary orbital period system \cite[]{gal05a}.
-\todo{give more detail about how this was done}
-
-In a subsequent outburst in 2015, a single burst was detected by {\it Swift}/XRT \cite[]{kuin15}, offering the opportunity to verify the previously-determined distance limit. The estimated peak 
+The method is relatively simple\footnote{Although the method is straighforward, we choose not to implement it in {\sc concord}, as this is a relatively niche case.
+We include it as an example here for completeness.};
+%
+the persistent flux history is converted to an accretion rate history given an assumed distance (via eq.
+??), and a burst train is generated.
+The times of the predicted bursts are compared to the good time intervals of the observations, and the simulation is repeated over a range of distances and times for the first burst.
+The likelihood for a given distance is then estimated as the fraction of simulated burst trains for which all the bursts fall within data gaps, and hence are missed.
+
+Larger distance (or H-fraction) will imply larger burst rates, and hence a smaller likelihood of missing any.
+The constraints obtained suggest the distance is $\lesssim6$~kpc (at 3-sigma significance), provided $X_0\approx0.7$, a reasonable choice given the expectation of a H-rich donor in this $2.46$~hr binary orbital period system \cite[]{gal05a}.
+%
+A separate argument is used to give a lower limit on the distance.
+
+% moved the discussion to the results section
+% ... and later moved it back --- dkg 9/21
+
+%\subsubsection{The distance to IGR J00291+5934}
+%
+%Previously we used the absence of thermonuclear bursts in the accretion-powered millisecond pulsar IGR~J00291+5934 to constrain the distance (\S\ref{subsec:zerobursts}).
+%%
+In a subsequent outburst in 2015, a single burst was detected by {\it Swift}/XRT \cite[]{kuin15}, offering the opportunity to verify the previously-determined distance range. The estimated peak 
 % 0.3--10~keV flux in the range (3.0--$5.6)\times10^{-11}\ \epcs$. 
 0.1--35~keV flux was $(18\pm4)\times10^{-8}\ \epcs$ \cite[]{defalco17}, with the burst exhibiting spectral variations indicative of PRE.
-By comparing the peak flux to the empirical Eddington luminosity (see \S\ref{subsec:1burst}, below), the corresponding distance
+By comparing the peak flux to the empirical Eddington luminosity (see \S\ref{subsec:1burst}), the corresponding distance
 \cite[including the inclination range of 22--32$^\circ$ suggested by][]{torres08},
@@
 \plotone{fig1.pdf}
-\caption{Estimated distance distributions for the accretion-powered X-ray pulsar IGR~J00291+5934, based on the peak flux of the sole thermonuclear burst observed by \cite{defalco17}.
+\caption{Estimated distance for the accretion-powered X-ray pulsar IGR~J00291+5934, based on the peak flux of the sole thermonuclear burst observed by \citet[][solid histograms]{defalco17};
+%
+and based on the {\it non-}detection of bursts by \citet[][red dashed lines]{gal06b}.
+%
+Note the good agreement between the methods.
 %
@@
 Imposing the inclination constraint of 22--32$^\circ$ suggested by \cite{torres08}, for which the predicted burst anisotropy factor is $\xi_b=0.704$ on average, gives instead the orange-shaded histogram.
-The resulting $1\sigma$ confidence interval is $5.0_{-0.5}^{+0.7}$~kpc.
+The resulting $1\sigma$ confidence interval (green lines) is $5.0_{-0.5}^{+0.7}$~kpc.
 \label{fig:dist_igr00291}}
@@
 
-\subsection{One burst -- estimating the burst rate} 
+\subsection{One burst -- estimating the distance and burst rate} 
 \label{subsec:1burst}
@@
 
-For a single burst detected from a source, we can expect to measure the burst peak flux, fluence, and timescale, and infer the distance and hence the expected burst rate.
+For a single burst detected from a bursting source, provided the data quality and signal-to-noise are sufficient, we can expect to measure the burst peak flux, fluence, and timescale.
+If the burst exhibits PRE, we can infer the distance (\S\ref{subsec:distance}); from the persistent flux, the accretion rate; and (via burst models) the expected burst rate.
 The detailed shape of the lightcurve can provide constraints from comparisons with numerical models (see \S\ref{subsec:lccompare}), but here we focus on the more straightforward calculation which may be done from the simple burst measurements alone.
 
-We use for example the sole burst detected with \igr/JEM-X from IGR~J17591$-$2342 \cite[]{kuiper20}.
+[{\bf detailed description of approach for distance here}]
+
+The distance can be estimated from the peak flux of a PRE burst based on equation \ref{disteq}.
+The challenge is taking into account the uncertainties in the various parameters that go into this calculation.
+If we adopt the empirical Eddington luminosity of \cite[]{kuul03a}, the only remaining factor is the burst emission anisotropy $\xi_b$.
+%
+As for the case of the burst in IGR~J00291+5934, the approach adopted in {\sc concord} to take this into account is to generate a distribution of $\xi_b$ values, based on an isotropic distribution of inclinations within known limits.
+The presence (absence) of X-ray dips occurring at the binary orbital period will constrain the inclination to greater (less) than $75^\circ$ [{\bf ref}].
+However, more stringent constraints may be available.
+
+% [{\bf what if you don't want to use the empirical value? }]
+The distance method in {\sc concord} can instead estimate the distance based on the theoretical Eddington luminosity (equation \ref{ledd}).
+In that case, values for $M_{\rm NS}$, $T_e$, $X$ and $1+z$ are assumed; these values, or distributions thereof, can also be provided by the user.
+No intrinsic uncertainty is otherwise assigned to the theoretical value; in contrast, the uncertainty in the empirical value will contribute to the overall distance uncertainty.
+
+% moved from later section
+% \subsubsection{System properties of IGR~J17591$-$2342}
+
+We use as an example 
+% of analysis of a single burst (\S\ref{subsec:1burst}) 
+the event detected with \igr/JEM-X from IGR~J17591$-$2342 \cite[]{kuiper20}.
 This burst was observed  when the estimated persistent bolometric flux was $(1.2\pm0.2)\times10^{-9}\ \epcs$, and exhibited a fluence of $E_b =(1.1\pm0.1)\times10^{-6}\ {\rm erg\,cm^{-2}}$.
@@
 
-\cite{kuiper20} first estimated the distance to the source based on the empirical value of the Eddington luminosity, and incorporating the expected anisotropy for an inclination range of 
+\cite{kuiper20} estimated the distance to the source based on the empirical value of the Eddington luminosity, and incorporating the expected anisotropy for an inclination range of 
 % kuiper20 uses slightly more restrictive 28-30
@@
 Here the uncertainty is dominated by the peak flux, since the anisotropy factor is effectively fixed for such a narrow range of inclination, at $\xi_b=0.70$.
+%
+The PDF for the distance, which is distinctly non-Gaussian, is plotted in Figure \ref{fig:dist_igr17591}.
+We also show the corresponding distribution resulting from adoption of the theoretical Eddington luminosity, according to equation \ref{ledd}.
 
@@
 
-We can use this distance constraint to estimate in turn the accretion and burst rate, as follows.
+\begin{figure}[ht!]
+\plotone{fig2.pdf}
+\caption{Distance estimates for IGR J17591$-$2342, based on the single PRE burst detected with JEM-X \cite[]{kuiper20}.
+The blue histogram and green 1-$\sigma$ limits replicate the analysis in that paper, comparing the flux to the empirical Eddington limit of \cite[]{kuul03a}, and adopting the inclination limis of \cite[]{sanna18}.
+For comparison we also include the inferred distance distribution (orange histogram) adopting the theoretical Eddington luminosity, for a neutron-star mass of $1.4M_\odot$, and surface redshift of $1+z=1.259$.
+\label{fig:dist_igr17591}}
+\end{figure}
+
+The {\sc concord} code allows us to pass the derived distance distribution directly to the method for calculating the accretion rate at the time of the burst, also incorporating the dependence of the persistent flux anisotropy (which is not the same as for the burst flux).
 %
@@
 via the {\tt mdot} method, and incorporating the estimated distance distribution already determined. 
-The estimated value is within the range $(2.0_{-0.4}^{+0.6})\times10^3\ \rm{g\,cm^{-2}\,s^{-1}}$ with the usual assumed range for neutron star mass and radius (and hence redshift).
-
-In parallel, we can estimate the ignition column using equation \ref{eq:column}, given an estimate of the nuclear burning yield, $\qnuc$.
-; as before, we could potentially obtain an estimate of this quantity by comparison of the observed lightcurve to simulations.
-However, in the absence of an estimate of this quantity, all is not lost.
-
-Instead we can work backwards, 
-% 
-and assume a set of randomly generated values of \Xb\
-to estimate $\yign$ from equation \ref{column}, and $\Delta t_{\rm rec} = \yign/\dot{m}$.
-%
-We choose a uniform distribution for \Xb\ in the range 0--0.7; this distribution is probably not realistic for a wide sample of bursts, but it is difficult to do better.
-% text copied from Kuiper et al. 2020 down to ********
-% not yet edited (need to modify sufficiently to avoid plagiarism issues
-Given each pair of (\Xb,  $\Delta t_{\rm rec}$) we can infer the accreted H-fraction $X_0$, based on 
-% the expectation that hydrogen burns via the hot-CNO cycle and will be exhausted in a time $t_{\rm CNO} =9.8(X_0/0.7)(Z_{\rm CNO}/0.02)^{-1}$ \citep[e.g.][]{lampe16}.
-equation \ref{eq:X_0}.
-%
-Since $Z_{\rm CNO}$ is also unknown, we similarly draw random values from a uniform distribution in the range 0.0--0.02.
-We then select only those $X_0$ values within a physically realistic range $X_0<0.75$, and compute the confidence limits on the remaining parameters.
-
-For the single burst observed from IGR~J17591$-$2342 the inferred ignition column is $2.2_{-0.8}^{+1.4}\times10^8\,{\rm g\,cm^{-2}}$, and the average expected recurrence time is $1.8\pm0.7$~d. 
-% Such long recurrence times likely guarantee that the hydrogen in the accreted fuel is exhausted at the base prior to ignition, and explaining the fast rise and short duration of the burst. 
-We can also infer lower (upper) limits on $X_0$ ($Z_{\rm CNO}$), although these limits are not strongly constraining; we find $X_0>0.17$ and $Z_{\rm  CNO}<0.017$ at 95\% confidence.
-% *******
+%
+Based on the distance constraints, the estimated accretion rate  is within the range $(2.0_{-0.4}^{+0.6})\times10^3\ \rm{g\,cm^{-2}\,s^{-1}}$ for the usual assumed range for neutron star mass and radius (and hence redshift).
 
@@
 
+% here's the clever bit
+
+Given a numerical burst ignition model, we could now go on to estimate the recurrence time, and check if it was consistent with the detection of only a single burst over the entire observed outburst.
+However, there is an extremely wide range of burst recurrence times at fixed accretion rate, depending upon the unknown fuel composition.
+we could potentially obtain constraints on the fuel composition by comparison of the observed lightcurve to simulations, but it's also possible that multiple different combinations of accretion rate and composition could give rise to similar composition at ignition (and hence shape of the lightcurve).
+
+Instead, we can estimate the ignition column using equation \ref{eq:column}, given an estimate of the nuclear burning yield, $\qnuc$.
+Since this quantity is unkown, we assume a set of randomly generated values of \Xb\
+to estimate $\yign$ from equation \ref{eq:column}
+%
+We choose a uniform distribution for \Xb\ in the range 0--0.7; this distribution is probably not realistic for a wide sample of bursts, but it is difficult to do better.
+% text copied from Kuiper et al. 2020 down to ********
+% For the single burst observed from IGR~J17591$-$2342 
+The inferred ignition column is $2.2_{-0.8}^{+1.4}\times10^8\,{\rm g\,cm^{-2}}$, and the average expected recurrence time is $1.8\pm0.7$~d. 
+% Such long recurrence times likely guarantee that the hydrogen in the accreted fuel is exhausted at the base prior to ignition, and explaining the fast rise and short duration of the burst. 
+, and $\Delta t_{\rm rec} = \yign/\dot{m}$.
+% not yet edited (need to modify sufficiently to avoid plagiarism issues
+Given each pair of (\Xb,  $\Delta t_{\rm rec}$) we can infer the accreted H-fraction $X_0$, based on 
+% the expectation that hydrogen burns via the hot-CNO cycle and will be exhausted in a time $t_{\rm CNO} =9.8(X_0/0.7)(Z_{\rm CNO}/0.02)^{-1}$ \citep[e.g.][]{lampe16}.
+equation \ref{eq:X_0}.
+%
+Since $Z_{\rm CNO}$ is also unknown, we similarly draw random values from a uniform distribution in the range 0.0--0.02.
+We then select only those $X_0$ values within a physically realistic range $X_0<0.75$, and compute the confidence limits on the remaining parameters. 
+%
+We can also infer lower (upper) limits on $X_0$ ($Z_{\rm CNO}$), although these limits are not strongly constraining; we find $X_0>0.17$ and $Z_{\rm  CNO}<0.017$ at 95\% confidence.
+% *******
+
 % ------------------------------------------------------------------------------
@@
 We describe the general approach below including the effects of these two issues.
-
-Given the measured $\alpha$-value, the nuclear energy generation rate $Q_{\rm nuc}$ can be estimated, as follows:
-\begin{eqnarray}
-\alpha & = & \frac{\Delta t F_{\rm per} c_{\rm bol}}{E_B} \\
-& = & \frac{Q_{\rm grav}}{Q_{\rm nuc}}\frac{\xi_b}{\xi_p}(1+z)
-\end{eqnarray}
-where $Q_{\rm grav} = c^2z/(1+z) \approx GM_{\rm NS}/R_{\rm NS}$ is the gravitational energy release per gram.
-Note the dependence of the observational $\alpha$-value on both the surface redshift and the anisotropy parameters.
-The nuclear energy generation rate is thus
-\begin{equation}
-Q_{\rm nuc} = \frac{c^2 z}{\alpha}\frac{\xi_b}{\xi_p} \label{qnuc}
-\end{equation}
-Thus, given an estimate of $z$ and the inclination, the burst measurements can be used to infer $Q_{\rm nuc}$.
 
@@
 
-Given an estimate of $Q_{\rm nuc}$, the average hydrogen fraction $\langle X\rangle$ in the fuel layer at ignition has been estimated to date using the relation $Q_{\rm nuc}=1.6 + 4\langle X\rangle$~MeV/nucleon \cite[e.g.][and references therein]{gal03d}.
-This expression includes $\approx35$\% losses
-attributed to neutrino emission \cite[]{fuji87}.
-However, the 35\% value applies only to $\beta$-decays in the much more extensive {\it rp}-process burning chain, and recent {\sc kepler} simulations indicate instead an approximation of
-% $Q{\rm nuc}=0.96+6.6\langle X\rangle$ 
-$Q{\rm nuc}=1.35+6.05\langle X\rangle$ 
-% (Goodwin et al. 2018, in preparation). 
-\cite[]{goodwin19a}.
-[{\it comment on improved quadratic expression}]
-%
 Thus, by substituting in equation \ref{qnuc} we can estimate 
@@
 \end{itemize}
-
-% comment on the inability to measure distance/mass/radius? Or can we actually do that?
-% I think you can get distance by comparing the ignition column with the fluence
-
-\subsection{Inferring distance} 
-\label{subsec:distance}
-
-The observed burst energy permits estimation of the ignition depth, assuming complete fuel consumption.
-For a burst fluence $E_b$, the total burst energy release (taking into account the anisotropy effect) is
-\begin{equation}
-E_{\rm burst} = 4\pi d^2 \xi_b E_b 
-\end{equation}
-The ignition depth is given by 
-\begin{eqnarray}
-y_{\rm ign} & = & E_{\rm burst}\frac{1 + z}{4\pi R_{\rm NS}^2 Q_{\rm nuc}} \\
-  & = & E_b \frac{\xi_b (1 + z)}{ Q_{\rm nuc}} \left(\frac{d}{R_{\rm NS}}\right)^2 
-\end{eqnarray}
-Independently, we can estimate the accreted mass as simply $\Delta M = ...$ and thus estimate the distance...?
 
@@
 
-\section{Application to simulated data} 
+\section{Results and analysis}
+\label{sec:results}
+
+Here we assemble and analyse a number of datasets to assess the precision and accuracy of the approaches described in the previous section.
+
+[{\bf this seems to be a duplicate of the previous section }]
+
+\subsection{Application to simulated data} 
 \label{sec:simulations}
 
-Here we assemble and analyse a number of datasets to assess the precision and accuracy of the approaches described in the previous section.
+To be added
+
+\subsection{Observations}
+
+A fundamental limitation with testing these approaches on observational data is the relative scarcity of independent measures of the system parameters.
+
+Separately we have assessed how well the observed burst peak flux measurements can constrain the distance (W\"orpel et al. 2020, in preparation).
+
+Here we provide a number of case studies intended to test or at least demonstrate these approaches.
+
+\subsubsection{Application to reference bursts}
+
+Here apply to the data on 4U 1820-30, 1808 and 1826-24, with comparisons to \cite{goodwin19c} and \cite{johnston20}.
+
+\section{Discussion}
+
+Summarise the results, such as they are
 
@@
 \acknowledgments
-
-To be added
 
@@
 
+% comment on the inability to measure distance/mass/radius? Or can we actually do that?
+% I think you can get distance by comparing the ignition column with the fluence
+
+\subsection{Inferring distance} 
+\label{subsec:distance}
+
+The observed burst energy permits estimation of the ignition depth, assuming complete fuel consumption.
+For a burst fluence $E_b$, the total burst energy release (taking into account the anisotropy effect) is
+\begin{equation}
+E_{\rm burst} = 4\pi d^2 \xi_b E_b 
+\end{equation}
+The ignition depth is given by 
+\begin{eqnarray}
+y_{\rm ign} & = & E_{\rm burst}\frac{1 + z}{4\pi R_{\rm NS}^2 Q_{\rm nuc}} \\
+  & = & E_b \frac{\xi_b (1 + z)}{ Q_{\rm nuc}} \left(\frac{d}{R_{\rm NS}}\right)^2 
+\end{eqnarray}
+Independently, we can estimate the accreted mass as simply $\Delta M = ...$ and thus estimate the distance...?
+
 % End of file `sample61.tex'.

### 2021-09-14 concord/diskmodel.py

@@ -4,6 +4,7 @@ import matplotlib.pyplot as plt
 import astropy.io.ascii as ascii
 from scipy.interpolate import interp1d
 import astropy.units as u
+import astropy.uncertainty as unc
 import os
 
 import pkg_resources
@@ -70,11 +71,17 @@ def anisotropy(inclination, model='he16_a', test=False):
 
         plt.show()
 
-# Calculate the values for the passed quantity, and return
-# numpy cos will correctly treat the units, so no need to do a conversion
-# in that case
+    # Calculate the values for the passed quantity, and return
+    # numpy cos will correctly treat the units, so no need to do a conversion
+    # in that case
 
-    theta = inclination
+    # Ensure the routine works with astropy distributions as well as arrays
+    if hasattr(inclination,'distribution'):
+        theta = inclination.distribution
+        scalar = False
+    else:
+        theta = inclination
+        scalar = (np.shape(theta) == ()) or (np.shape(theta) == (1,))
     if (hasattr(inclination,'unit') == False):
         print ("** WARNING ** assuming inclination in degrees")
         theta *= u.degree
@@ -83,7 +90,11 @@ def anisotropy(inclination, model='he16_a', test=False):
         xi_b = 1./(0.5+abs(np.cos(theta)))
         xi_p = 0.5/abs(np.cos(theta))
 
-        return xi_b, xi_p
+        if scalar:
+            return xi_b, xi_p
+        else:
+            return unc.Distribution(xi_b*u.dimensionless_unscaled), \
+                   unc.Distribution(xi_p*u.dimensionless_unscaled)
 
 
     elif model in he16_models:
@@ -103,7 +114,11 @@ def anisotropy(inclination, model='he16_a', test=False):
             xi_b = 1./(inv_xi_d+inv_xi_r)
             xi_p = 1./inv_xi_p
 
-        return xi_b, xi_p
+        if scalar:
+            return xi_b, xi_p
+        else:
+            return unc.Distribution(xi_b*u.dimensionless_unscaled), \
+                   unc.Distribution(xi_p*u.dimensionless_unscaled)
 
 
     else:

### 2021-09-14 concord/utils.py

@@ -1,27 +1,38 @@
 # Various utilities moved from burstclass.py
 # Augmented 2019 Aug with routines from Inferring\ composition.ipynb
 #
+# def value_to_dist(_num, nsamp=NSAMP_DEF, unit=None):
+# def len_dist(d):
 # def g(M,R,Newt=False,units='cm/s^2'):
-# def opz(M,R):
+# def redshift(M,R):
 # def calc_mr(g,opz):
 # def solve_radius(M,R_Newt,eta=1e-6):
 # def decode_LaTeX(string):
 # def Q_nuc(Xbar, quadratic=False, old_relation=False, coeff=False)
+# def X_0(xbar, zcno, tdel, opz=1.259, debug=False, old_relation=False):
+# def alpha(_tdel, _fluen, _fper, _c_bol=1.0, nsamp=NSAMP_DEF):
 # def hfrac(alpha, tdel, opz=1.259, zcno=0.02, old_relation=False, ...)
 # def iso_dist(nsamp=1000, imin=0., imax=IMAX_NDIP)
 # def dist(F_pk, F_pk_err, nsamp=10000, X=0.0, empirical=False, ...)
 # def mdot(F_per, F_per_err, dist, nsamp=10000, M=M_NS, R=R_NS, ...)
 # def yign(E_b, E_b_err, dist, R_NS=10., Xbar=0.7, opz=1.259, ...)
+# def luminosity(F_X, F_X_err=0.0, dist=8*u.kpc, dist_err=None, nsamp=NSAMP_DEF, ...)
 # def L_Edd(F_pk, F_pk_err, dist, dist_err=0.0, nsamp=10000, dip=False, ...)
 
 import astropy.units as u
+import astropy.uncertainty as unc
+from astropy.visualization import quantity_support
 
 # Some defaults
 
 CONF=68.        # default 68% confidence intervals
+NSAMP_DEF=1000  # default 1000 samples
 IMAX_NDIP=75.   # [degrees] maximum possible inclination angle for non-dipper
 M_NS=1.4 * u.M_sun  # default neutron-star mass
 R_NS=11.2 * u.km    # default neutron-star radius
+OPZ=1.259       # default redshift
+
+ETA=1.e-6       # generic fraction to detect deviation from default values
 
 # This parameter sets the prefactor for the time to burn all H via hot-CNO; see
[…]

### 2021-09-17 doc/ms.tex
Lots of small changes to improve robustness, consistency
Introduced intvl_to_errors to convert the percentiles output from functions to proper 2-sided errors as used in input.
Corrected the way value_to_dist deals with scalars, as well as how it treats units with or without prefactors.
Changed the warning, error messages to use the logger instead
Updates to several functions to improve consistency of behaviour

@@
 
-Thermonuclear (type-I) bursts are observed from neutron stars accreting from low-mass ($\lesssim 1\ M_\odot$) binary companions \cite[e.g.][]{gal17b}. ...
+Thermonuclear (type-I) bursts are observed from neutron stars accreting from low-mass ($\lesssim 1\ M_\odot$) binary companions \cite[e.g.][]{gal21a}. ...
 
@@
 
-[{\bf insert table with key burst parameters here? }]
+\todo{insert table with key burst parameters here? }
 
@@
 % \label{subsec:anisotropy}
-[{\bf bit more description here }]
+\todo{ bit more description here }
 
@@
 
-From data taken prior to the burst ignition we can measure the persistent flux $F_{\rm per}$, a measure of the accretion rate; [{\bf add expression here}]
+From data taken prior to the burst ignition we can measure the persistent flux $F_{\rm per}$, a measure of the accretion rate; \todo{ add expression here; see eq. \ref{eq:fper}}
 % and provided more than one burst is observed, we may also measure (or estimate) the recurrence time, $\Delta t$. 
@@
 %
-[{\bf mention the average method, as used in the rates paper }]
+\todo{mention the average method, as used in the rates paper }
 
@@
 \frac{\dot{m} Q_{\rm grav}}{4\pi d^2 (1+z)\xi_p c_{\rm bol}}
+\label{eq:fper}
 \end{equation}
@@
 Describe how the code is written and how it works, the astropy.distribution thingy, MCMC for error determination etc.
+\begin{itemize}
+  \item Approach for calculations via astropy distributions
+  \item 7(?) different methods for defining observational/input quantities
+\end{itemize}
 
@@
 %
-We focus on adopting a suitable approach given the availability of data, which will differ for different sources. 
-%
-We also describe how the {\sc concord} code can be used to fully account for the astrophysical uncertainties, and as a companion to this section (and the code itself) we provide a companion {\sc jupyter} notebook [{\bf ref}] which replicates the analyses described below.
+We describe how the {\sc concord} code can be used to fully account for the astrophysical uncertainties, and as a companion to this section (and the code itself) we provide a companion {\sc jupyter} notebook [{\bf ref}] which replicates the analyses described below.
+
+% simple example
+
+As a simple example of the code utility, we describe here how the confidence range for peak luminosity may be estimated for a burst from a source with independently known distance.
+We consider the PRE burst observed from 4U~0513$-$40 with the {\it Rossi X-ray Timing Explorer} ({\it RXTE}) PCA on MJD~54043.68857 \cite[observation ID 92403-01-15-04; MINBAR ID \#3443;][]{minbar}.
+According to the MINBAR analysis this burst reached a peak flux of $(21.7\pm0.6)\times10^{-9}\ \epcs$.
+The source is located in the globular cluster NGC~1851, for which the distance is estimated at $(10.32^{+0.20}_{-0.24})$~kpc \cite[]{WatkinsEtAl2015}.
+
+The isotropic luminosity can be calculated from the peak flux and distance using the {\tt luminosity} method, with {\tt isotropic=True} as $2.77\times10^{38}\ \epcs$. 
+%
+However, a more accurate calculation should take into account the uncertainty in the flux; the (asymmetric) uncertainty in the distance; and the likely effect of the anisotropy in the burst emission, based on the estimated system inclination $i>80^\circ$ \cite[]{fiocchi11}.
+This can be achieved by passing the flux and distance uncertainties to the routine; {\sc astropy} distributions are generated for both these parameters, with the required properties (e.g.
+Figure \ref{fig:burstlum}, panel (a)).
+We also pass the inclination limits, and an array of inclination values is generated assuming an isotropic distribution of inclinations within the limits.
+We calculate the anisotropy factor based on the modelling of \cite{he16}, and the user can select different models to quantify the variation introduced.
+%
+The resulting luminosity is significantly higher, $(4.85_{-0.42}^{+0.46})\times10^{38}\ \eps$ (Fig.
+\ref{fig:burstlum}, panel (b)); the discrepancy is primarily the result of the significant attenuation of the burst flux from the viewing angle close to the plane of the accretion disk.
+The corresponding anisotropy value $\xi_b$ is in the range 1.56--2 (Fig.
+\ref{fig:burstlum}, panel (c)).
+The dependence of the luminosity on the assumed inclination is illustrated in (Fig.
+\ref{fig:burstlum}, panel (d)).
+
+The output from the function is by default the median value and the lower and upper uncertainties, at the required confidence level (68\% by default).
+Using the {\tt fulldist=True} flag, the function instead returns a dictionary including the full distributions for the luminosity, distance, inclination, anisotropy factor, as well as the anisotropy model identifier.
+% and the confidence level
+Each of the distribution arrays can be accessed via the {\tt distribution} attribute.
+This object can be written to a file or used for input to subsequent analyses.
+
+\begin{figure}[ht!]
+\plotone{fig0.pdf}
+\caption{Illustration of calculation of burst peak luminosity using {\sc concord}, incorporating measurement uncertainties and the modeled effect of the burst anisotropy.
+The adopted distance PDF is shown in panel (a); note the slight skew in the distribution.
+The resulting luminosity PDF in panel (b) is substantially higher than the isotropic value ({\it red dashed line}).
+The $1\sigma$ confidence intervals are indicated by the green lines.
+The PDF for the anisotropy factor $\xi_b$ is shown in panel (c); the dependence of the luminosity on the adopted system inclination is shown in panel (d).
+\label{fig:burstlum}}
+\end{figure}
+
+[{\bf comment here perhaps on the additional uncertainties that might contribute; e.g. for $M$, $R$; code is flexible enough to incorporate them }]
+
+In the sections below we describe other examples,
+%
+focussing on adopting approaches tailored to the availability of data, which will differ for different sources.
 
@@
 
-The distance can be estimated from the peak flux of a PRE burst based on equation \ref{disteq}.
+The distance can be estimated from the peak flux of a PRE burst based on equation \ref{disteq}, using the {\tt dist} method.
 The challenge is taking into account the uncertainties in the various parameters that go into this calculation.
@@
 
-\cite{kuiper20} estimated the distance to the source based on the empirical value of the Eddington luminosity, and incorporating the expected anisotropy for an inclination range of 
+Following \cite{kuiper20}, we estimate the distance to the source based on the empirical value of the Eddington luminosity, and incorporating the expected anisotropy for an inclination range of 
 % kuiper20 uses slightly more restrictive 28-30
@@
 %
-Based on the distance constraints, the estimated accretion rate  is within the range $(2.0_{-0.4}^{+0.6})\times10^3\ \rm{g\,cm^{-2}\,s^{-1}}$ for the usual assumed range for neutron star mass and radius (and hence redshift).
+Based on the distance constraints, the estimated accretion rate  is within the range $(2.0_{-0.4}^{+0.6})\times10^3\ \rm{g\,cm^{-2}\,s^{-1}}$ for the usual assumed values for neutron star mass and radius (and hence redshift).
 
@@
 
-Instead, we can estimate the ignition column using equation \ref{eq:column}, given an estimate of the nuclear burning yield, $\qnuc$.
-Since this quantity is unkown, we assume a set of randomly generated values of \Xb\
-to estimate $\yign$ from equation \ref{eq:column}
-%
-We choose a uniform distribution for \Xb\ in the range 0--0.7; this distribution is probably not realistic for a wide sample of bursts, but it is difficult to do better.
+Instead, we can constrain the fuel composition via the ignition column (equation \ref{eq:column}), based on an estimate of the nuclear burning yield, $\qnuc$.
+Since this quantity is unknown, we assume a set of random uniformly-distributed\footnote{Such distribution may not be realistic, but it serves to demonstrate the code capabilities.
+The largest (and most unlikely) values of \Xb\ are suppressed by the limit on $X_0$ imposed during this process.} values of \Xb\ in the range 0--0.7
+to estimate $\yign$ from equation \ref{eq:column}.
+%
+We then estimate the recurrence time for each pair of $\yign$, $\dot{m}$ values as
+\begin{equation}
+\Delta t_{\rm rec} = (1+z)\yign/\dot{m}
+\end{equation}
+And finally calculate the fuel composition, $X_0$, for each set of \Xb, $\yign$, $\Delta t$ from equation \ref{eq:X_0}, again assuming a uniform distribution for $Z_{\rm CNO}$ between 0--0.02.
+%
+Because we have chosen uninformative prior distributions for \Xb\ and $Z_{\rm CNO}$, some of the $X_0$ values are unrealistic, well in excess of the maximum expected for such sources, 0.75 or so.
+We thus limit the distributions of each of the parameters to those such that $X_0\leq0.75$, and quote the $1\sigma$ parameter ranges for each.
+
+% We choose a uniform distribution for \Xb\ in the range 0--0.7; this distribution is % probably not realistic for a wide sample of bursts, but it is difficult to do better.
 % text copied from Kuiper et al. 2020 down to ********
 % For the single burst observed from IGR~J17591$-$2342 
-The inferred ignition column is $2.2_{-0.8}^{+1.4}\times10^8\,{\rm g\,cm^{-2}}$, and the average expected recurrence time is $1.8\pm0.7$~d. 
+% Original values; couldn't replicate in 2021 Sep
+% The inferred ignition column is $2.2_{-0.8}^{+1.4}\times10^8\,{\rm g\,cm^{-2}}$, and the average expected recurrence time is $1.8\pm0.7$~d. 
+% 2021 Sep values
+The inferred ignition column is $1.63_{-0.58}^{+1.05}\times10^8\,{\rm g\,cm^{-2}}$, and the average expected recurrence time is $1.19_{-0.42}^{+0.79}$~d. 
 % Such long recurrence times likely guarantee that the hydrogen in the accreted fuel is exhausted at the base prior to ignition, and explaining the fast rise and short duration of the burst. 
-, and $\Delta t_{\rm rec} = \yign/\dot{m}$.
 % not yet edited (need to modify sufficiently to avoid plagiarism issues
-Given each pair of (\Xb,  $\Delta t_{\rm rec}$) we can infer the accreted H-fraction $X_0$, based on 
-% the expectation that hydrogen burns via the hot-CNO cycle and will be exhausted in a time $t_{\rm CNO} =9.8(X_0/0.7)(Z_{\rm CNO}/0.02)^{-1}$ \citep[e.g.][]{lampe16}.
-equation \ref{eq:X_0}.
-%
-Since $Z_{\rm CNO}$ is also unknown, we similarly draw random values from a uniform distribution in the range 0.0--0.02.
-We then select only those $X_0$ values within a physically realistic range $X_0<0.75$, and compute the confidence limits on the remaining parameters. 
 %

### 2021-09-17 concord/utils.py

@@ -64,7 +64,7 @@ def create_logger():
     if not logger.handlers: # Check if created before, otherwise a reload will add handlers
         logger.setLevel(logging.INFO)
         handler = logging.StreamHandler()
-        formatter = logging.Formatter('%(levelname)s:%(name)s:%(message)s')
+        formatter = logging.Formatter('%(levelname)s:%(funcName)s:%(message)s')
         handler.setFormatter(formatter)
         logger.addHandler(handler)
     return logger
@@ -100,26 +100,40 @@ def value_to_dist(_num, nsamp=NSAMP_DEF, unit=None):
     # Check here if it's already a distribution; don't want to run this twice
 
     if hasattr(_num, 'distribution'):
-        print ('** WARNING ** this quantity is already a distribution')
+        logger.warning('this quantity is already a distribution')
         return _num
 
+    # Want to permit all these combinations:
+    # val (if unit is omitted as for the first 3, assumed in MINBAR units)
+    # (val, err)
+    # (val, loerr, hierr)
+    # val*unit
+    # (val, err)*unit
+    # (val, loerr, hierr)*unit
     num_unit = u.dimensionless_unscaled
     if hasattr(_num, 'unit'):
         # strip off the unit here to make the code which follows, work
         num_unit = _num.unit
-        assert (unit is None) or (unit == num_unit)
+        if hasattr(unit, 'unit'):
+            assert (unit.unit == num_unit) # supplied unit might have prefactor
+        else:
+            assert (unit is None) or (unit == num_unit)  # supplied unit might have prefactor
         num = _num.value
     else:
+        num = _num
         if unit is not None:
             num_unit = unit
-        num = _num
 
     if np.shape(num) == () or np.shape(num) == (1,):
         # scalar, with no errors; just replicate that value
-        return unc.Distribution(np.full(nsamp, num) * num_unit)
+        # return unc.Distribution(np.full(nsamp, num) * num_unit)
+        # since we can combine distributions with scalars, no need to make a
+        # distribution out of this scalar
+        return num * num_unit
     elif np.shape(num) == (2,):
         if num[1] is None:
-            return unc.Distribution(np.full(nsamp, num[0]) * num_unit)
+            # return unc.Distribution(np.full(nsamp, num[0]) * num_unit)
+            return num[0] * num_unit
         # value and error
[…]

### 2021-09-20 doc/ms.tex
Swapped sense of quadratic flag in Q_nuc

@@
 
-\subsection{Burst energetics}
+\subsection{Burst energetics and fuel composition}
 \label{subsec:energetics}
@@
 The fluence (and the shape of the burst profile) is determined by the amount of accumulated fuel, and its composition.
-For accretion of mixed H/He with hydrogen mass fraction $X_0$, the composition at ignition may be modified substantially by $\beta$-limited CNO burning.
+For accretion of mixed H/He with hydrogen mass fraction $X_0$, the composition at ignition may be modified substantially by $\beta$-limited CNO burning, ``catalysed'' by CNO nuclei, with mass fraction $Z_{\rm CNO}$.
 In extreme cases the recurrence time is sufficiently long that the accreted hydrogen at the base of the fuel layer is exhausted, and ignition of intense, fast He-rich bursts occur.
 
-From data taken prior to the burst ignition we can measure the persistent flux $F_{\rm per}$, a measure of the accretion rate; \todo{ add expression here; see eq. \ref{eq:fper}}
-% and provided more than one burst is observed, we may also measure (or estimate) the recurrence time, $\Delta t$. 
-the persistent spectra of burst sources are  more complex than the bursts themselves, and typically exhibit thermal as well as non-thermal components [{\bf ref}].
-%
-These parameters may be combined to estimate the $\alpha$-value, the (observed) ratio of the burst to persistent flux: 
-\begin{equation}
-\alpha = \frac{\Delta t F_{\rm per} c_{\rm bol}}{E_B}
-\end{equation}
-where $c_{\rm bol}$ is the bolometric correction giving the inverse fraction of total persistent flux emitted in the instrumental band.
-%
-The $\alpha$ parameter is an important diagnostic of the burst fuel and ignition conditions.
-%
-\todo{mention the average method, as used in the rates paper }
-
-Given the measured $\alpha$-value, the nuclear energy generation rate $Q_{\rm nuc}$ can be estimated, as follows:
-\begin{eqnarray}
-\alpha & = & \frac{\Delta t F_{\rm per} c_{\rm bol}}{E_B} \\
-& = & \frac{Q_{\rm grav}}{Q_{\rm nuc}}\frac{\xi_b}{\xi_p}(1+z)
-\end{eqnarray}
-where $Q_{\rm grav} = c^2z/(1+z) \approx GM_{\rm NS}/R_{\rm NS}$ is the gravitational energy release per gram.
-Note the dependence of the observational $\alpha$-value on both the surface redshift and the anisotropy parameters.
-The nuclear energy generation rate is thus
-\begin{equation}
-Q_{\rm nuc} = \frac{c^2 z}{\alpha}\frac{\xi_b}{\xi_p} \label{qnuc}
-\end{equation}
-Thus, given an estimate of $z$ and the inclination, the burst measurements can be used to infer $Q_{\rm nuc}$.
-
-Given an estimate of $Q_{\rm nuc}$, the average hydrogen fraction $\langle X\rangle$ in the fuel layer at ignition has been estimated to date using the relation $Q_{\rm nuc}=1.6 + 4\langle X\rangle$~MeV/nucleon \cite[e.g.][and references therein]{gal03d}.
-This expression includes $\approx35$\% losses
-attributed to neutrino emission \cite[]{fuji87}.
-However, the 35\% value applies only to $\beta$-decays in the much more extensive {\it rp}-process burning chain, and recent {\sc kepler} simulations indicate instead an approximation of
-% $Q{\rm nuc}=0.96+6.6\langle X\rangle$ 
-$Q{\rm nuc}=1.35+6.05\langle X\rangle$ 
-% (Goodwin et al. 2018, in preparation). 
-\cite[]{goodwin19a}.
-[{\it comment on improved quadratic expression}]
-%
-The average H-fraction at ignition $\langle X\rangle$ is a lower limit for the H-fraction in the accreted fuel, $X_0$, as for most sources the H-fraction will be reduced by steady burning prior to burst ignition.
-%
-The recurrence time for the burst can further be used to estimate $X_0$, although the burning rate also depends on the CNO mass fraction, $Z_{\rm CNO}$.
-
-additional data analysis comments
-\begin{itemize}
-\item correcting lightcurve for variations e.g. from dips, or achromatic variations
-\item using different spectral models than blackbody
-\item allowing the persistent flux to vary
-\item maybe a table with the parameters?
-\end{itemize}
-
-For a fuel layer consisting of mixed H/He, $\qnuc$ depends on the mean hydrogen fraction at ignition, \Xb:
-% The nuclear energy generation rate $\qnuc$ depends on the composition; 
-studies with the 1-D numerical code {\sc kepler} suggest that 
-\begin{eqnarray}
-Q_{\rm nuc} &= & 1.31+6.95\Xb-1.92\Xb^2 \nonumber \\
-& \approx & 1.35 + 6.05 \Xb
-\label{eq:qnuc}
-\end{eqnarray}
-\cite[]{goodwin19a}.
-% where $\Xb$ is the average H-fraction of the fuel layer at ignition. 
-For fuel with solar composition (i.e. $X=0.7$), $Q_{\rm nuc}=5.22$~MeV/nucleon.
-
-{\it Perhaps put in here some estimates of the fuel composition as a function of $X_0$ and recurrence time; could perhaps develop a ``prior'' for $\Xb$ }
-
-Provided an estimate of $\Xb$ is available, we can estimate the burst column that has ignited as
+The burst ignition column can be estimated from the fluence $E_b$ as
 % (1e-6*u.erg/u.cm**2*(10.*u.kpc).to('cm')**2/(5.22*u.MeV/const.m_p)*1.259
@@
 % {\it need to update prefactor, denominator in $Q_{\rm nuc}$ term here } - done
+where $\qnuc$ is the energy generation rate for the burst.
+
+For a fuel layer consisting of mixed H/He, $\qnuc$ depends on the mean hydrogen fraction at ignition, \Xb. 
+%
+% Given an estimate of $Q_{\rm nuc}$, the average hydrogen fraction $\langle X\rangle$ in the fuel layer at ignition has been estimated to date 
+In the absence of other information, $\qnuc$ (and hence $y$) has in the past been estimated by adopting a value for 
+% the average H-fraction at ignition, 
+$\Xb$,
+%
+and using the relation $Q_{\rm nuc}=1.6 + 4\langle X\rangle$~MeV/nucleon \cite[e.g.][and references therein]{gal03d}. This expression includes $\approx35$\% losses
+attributed to neutrino emission \cite[]{fuji87}. However, the 35\% value applies only to $\beta$-decays in the much more extensive {\it rp}-process burning chain.
+% The nuclear energy generation rate $\qnuc$ depends on the composition; 
+Recent studies with the 1-D numerical code {\sc kepler} suggest instead that 
+\begin{eqnarray}
+Q_{\rm nuc} &= & 1.31+6.95\Xb-1.92\Xb^2 \nonumber \\
+& \approx & 1.35 + 6.05 \Xb
+\label{eq:qnuc}
+\end{eqnarray}
+\cite[]{goodwin19a}.
+% where $\Xb$ is the average H-fraction of the fuel layer at ignition. 
+For fuel with solar composition (i.e. $X=0.7$), $Q_{\rm nuc}=5.22$~MeV/nucleon.
+%
+The average H-fraction at ignition $\Xb$ is a lower limit for the H-fraction in the accreted fuel, $X_0$, as for most sources the H-fraction will be reduced by steady burning prior to burst ignition.
+
+Provided the burst recurrence time is known, $\qnuc$ can be estimated via the $\alpha$-parameter, the (observed) ratio of the burst to persistent flux.
+%
+From data taken prior to the burst ignition we can measure the persistent flux $F_{\rm per}$, a measure of the accretion rate; \todo{ add expression here; see eq. \ref{eq:fper}}
+% and provided more than one burst is observed, we may also measure (or estimate) the recurrence time, $\Delta t$. 
+the persistent spectra of burst sources are  more complex than the bursts themselves, and typically exhibit thermal as well as non-thermal components [{\bf ref}].
+%
+These parameters may be combined to estimate the $\alpha$-value: 
+\begin{equation}
+\alpha = \frac{\Delta t F_{\rm per} c_{\rm bol}}{E_b}
+\label{eq:alpha}
+\end{equation}
+where $c_{\rm bol}$ is the bolometric correction giving the inverse fraction of total persistent flux emitted in the instrumental band.
+%
+% \todo{mention the average method, as used in the rates paper }
+As for the burst rate, we may estimate the average $\alpha$ for a set of $N$ bursts with  
+% average fluence $\langle E_b\rangle$, 
+fluences $E_{b,i}$, 
+%
+observed in $n$ low-duty cycle observations with persistent flux $F_j$ and exposure $T_j$ as
+\begin{equation}
+\langle \alpha\rangle = \frac{c_{\rm bol}\sum^n T_jF_j}{\sum^N E_b,i}
+% {\langle E_b\rangle}
+\label{eq:alphaav}
+\end{equation}
+assuming a common bolometric correction $c_{\rm bol}$ for each observation.
+
+The $\alpha$ parameter is an important diagnostic of the burst fuel and ignition conditions.
+%
+The ``observed'' value given by equations \ref{eq:alpha} and \ref{eq:alphaav} must be interpreted with a little caution, as they likely incorporate the effects of anisotropy of the burst and persistent emission.
+%
+We can also write $\alpha$ as
+% Given the measured $\alpha$-value, the nuclear energy generation rate $Q_{\rm nuc}$ can be estimated, as follows:
+\begin{equation}
+\alpha % & = & \frac{\Delta t F_{\rm per} c_{\rm bol}}{E_B} \\
+ =  \frac{Q_{\rm grav}}{Q_{\rm nuc}}\frac{\xi_b}{\xi_p}(1+z)
+ \label{eq:alphath}
+\end{equation}
+where $Q_{\rm grav} = c^2z/(1+z) \approx GM_{\rm NS}/R_{\rm NS}$ is the gravitational energy release per gram.
+Note the implicit dependence of the ``observational'' $\alpha$-value (equations \ref{eq:alpha} and \ref{eq:alphaav}) on both the surface redshift and the anisotropy parameters, illustrated in equation \ref{eq:alphath}.
+%
+They expected values for $Q_{\rm grav}$ are $\approx190$~MeV~nucleon$^{-1}$ (for $M_{\rm NS}=1.4\ M_\odot$ and $R_{\rm NS}=11.2$~km, giving $1+z=1.259$).
+Thus, excluding the anisotropy effects, $\alpha$ should be $\approx40$
+% (const.c**2*(cd.OPZ-1)/cd.OPZ).to('MeV g-1')*const.m_p.to('g')/(5.22*u.MeV)
+% 36.977056, more precisely
+for bursts burning H-rich fuel, or 140 for pure He.
+
+The nuclear energy generation rate is thus
+\begin{equation}
+Q_{\rm nuc} = \frac{c^2 z}{\alpha}\frac{\xi_b}{\xi_p} \label{qnuc}
+\end{equation}
+%
+%Thus, given an estimate of $z$ and the inclination, the burst measurements can be used to infer $Q_{\rm nuc}$. 
+%
+% Thus, 
+By substituting in equation \ref{eq:qnuc} we can estimate 
+\begin{equation}
+%\langle X\rangle 
+\Xb
+% In [42]: c.c**2/(6.0455*u.MeV/c.m_p).to("m**2/s**2")                             
+% Out[41]: <Quantity 155.20173377>
+%% = \frac{c^2z/\alpha}{6.6\ {\rm Mev/nucleon}}\frac{\xi_b}{\xi_p} - 0.145
+% = z\frac{142}{\alpha}\frac{\xi_b}{\xi_p} - 0.145 \label{xbar}
+ = z\frac{155}{\alpha}\frac{\xi_b}{\xi_p} - 0.223 \label{xbar}
+\end{equation}
+[{\it comment on the degree of error introduced by ignoring anisotropy etc.}] 
+%
+Clearly this expression is only applicable for $\alpha$-values up to some limit, which we calculate as
+\begin{equation}
+%In [43]: c.c**2/(1.*u.MeV/c.m_p).to("m**2/s**2")/1.3455        
+% Out[43]: <Quantity 697.34082607>
+% \alpha \leq 980\, z \frac{\xi_b}{\xi_p}
+\alpha \leq 697\, z \frac{\xi_b}{\xi_p}
+\end{equation}
+Interestingly, this limit is 
+% $\approx70$\% 
+%  c.c**2/(1.*u.MeV/c.m_p).to("m**2/s**2")/1.6                 
+% Out[44]: <Quantity 586.42005092>
+% In [45]: 697./586.                                                               
+% Out[45]: 1.189
+$\approx20$\%
+%
+larger than the equivalent value derived for the old expression for $Q_{\rm nuc}$.
+This result suggests that larger observed values of $\alpha$ may be accommodated without resorting to explanations including incomplete burning of burst fuel \cite[e.g.][]{bcatalog}.
+
+% upper limit on xbar
+Conversely, for low values of $\alpha$, we may apply the constraint that $\Xb \lesssim 0.77$, corresponding to the expected maximum possible for accreted fuel with primordial abundances.
+
+The hydrogen fraction at ignition $\Xb$ and the burst recurrence time may then be used to estimate the hydrogen fraction in the fuel, $X_0$.
+The H-fraction at the base of the fuel layer is reduced steadily by $\beta$-limited hot CNO burning, and will completely exhaust the accreted hydrogen in a time \cite[]{lampe16}:
+\begin{equation}
+t_{\rm CNO} =9.8\left(\frac{X_0}{0.7}\right) \left(\frac{Z_{\rm CNO}}{0.02}\right)^{-1}
+\end{equation}
+measured in the neutron star frame.
+% updated 2019 Aug 26 with exhaustion case
+The hydrogen fraction at the base will thus be $X_0[1-\Delta t/(1+z)t_{\rm CNO}]$, 
+% [{\it should this expression include time dilation?}] -- yes
+%
+provided that $\Delta t < (1+z)t_{\rm CNO}$.
+Once $\Delta t$ exceeds the time to burn all the hydrogen at the base, the abundance there will be zero, and a growing layer of pure He fuel will develop.
+The average H-fraction in the layer for these two cases will be
+%so that the average H-fraction in the layer will be 
+\begin{equation}
+\Xb = \begin{cases}
+%  X_0\left[1-0.5\frac{\Delta t}{(1+z)t_{\rm CNO}}\right] & \Delta t \leq (1+z)t_{\rm CNO} \\
+%  0.5X_0\frac{(1+z)t_{\rm CNO}}{\Delta t} & \Delta t > (1+z)t_{\rm CNO}
+  X_0(1-0.5f_{\rm burn}) & \Delta t \leq (1+z)t_{\rm CNO} \\
+  0.5X_0/f_{\rm burn} & \Delta t > (1+z)t_{\rm CNO}
+  \end{cases}
+\end{equation}
+where $f_{\rm burn} = \frac{\Delta t}{(1+z)t_{\rm CNO}}$ is the ratio of the recurrence time to the time to burn all the H.
+%
+We can combine this expression with equation \ref{xbar} to give
+\begin{equation}
+X_0 = \begin{cases}
+%        & \langle X\rangle +0.35\left[\frac{\Delta t}{(1+z)9.8\ {\rm hr}}\right]
+%                            \left(\frac{Z_{\rm CNO}}{0.02}\right) \nonumber \\
+         z\frac{142}{\alpha}\frac{\xi_b}{\xi_p} - 0.145 
+                           +\left[\frac{\Delta t}{(1+z)28\,{\rm hr}}\right]
+                       \left(\frac{Z_{\rm CNO}}{0.02}\right) & f_{\rm burn} \leq 1\\
+         \sqrt{ \frac{\Delta t}{(1+z)7\,{\rm hr}}\frac{Z_{\rm CNO}}{0.02}
+                        \left(z\frac{142}{\alpha}\frac{\xi_b}{\xi_p} - 0.145\right) }
+                        & f_{\rm burn} > 1
+        \end{cases}
+    \label{eq:X_0}
+\end{equation}
+Practically, the issue with these expressions is that calculating $f_{\rm burn}$ requires knowledge of $X_0$, which is the unknown we are trying to constrain.
+One approach is to adopt a trial value of $X_0$, calculate $f_{\rm burn}$ and hence an updated estimate of $X_0$ via equation \ref{eq:X_0}, and iterate until no further change in the estimate arises.
+
+\begin{figure}[ht!]
+\plotone{xbar.pdf}
+\caption{Hydrogen mass fraction of the accreted fuel, $X_0$, inferred from the average H-fraction of the burst at ignition, $\overline{X}$, as a function of burst recurrence time $\Delta t$.
+$X_0$ is calculated from equation \ref{eq:X_0}, assuming $\zcno=0.001$ (blue line) or 0.02 (other lines).
+%
+Note how for some combinations of $\overline{X}$, $\Delta t$ the inferred fuel H-fraction is $>0.77$ (red dashed line).
+Such values are physically implausible, and may be resolved if the CNO mass fraction $\zcno$ is smaller than the adopted value.
+% 
+The top $x$-axis shows the corresponding $\qnuc$ values.
+\label{fig:xbar}}
+\end{figure}
+
+Clearly this estimate of $X_0$ will incorporate systematic errors arising from the assumed values of $z$, the inclination, and $Z_{\rm CNO}$. We explore the typical range of these errors in 
+% the next section.
+\S\ref{sec:simulations}
+
+\todo{rationalise older/redundant text below to end of section}
+
+The recurrence time for the burst can further be used to estimate $X_0$, although the burning rate also depends on the CNO mass fraction, $Z_{\rm CNO}$.
+
+additional data analysis comments
+\begin{itemize}
+\item correcting lightcurve for variations e.g. from dips, or achromatic variations
+\item using different spectral models than blackbody
+\item allowing the persistent flux to vary
+\item maybe a table with the parameters?
+\end{itemize}
+
+{\it Perhaps put in here some estimates of the fuel composition as a function of $X_0$ and recurrence time; could perhaps develop a ``prior'' for $\Xb$ }
+
+% \todo{move later sections here on determining the fuel composition?}
+
+Given a minimally complete set of burst observations as described in \S\ref{subsec:data}, a common approach \cite[e.g][]{falanga11} is to estimate the fuel composition at ignition, and hence the accreted fuel composition, based on simple analytic estimates of the burst energy production $Q_{\rm nuc}$.
+
+One issue with this approach 
+% (as for analyses based on one burst, as described in \S\ref{subsec:1burst}) 
+is that the effects of anisotropy are frequently neglected.
+%
+A second issue which has emerged recently is that the commonly used estimates for the nuclear energy generation rate $Q_{\rm nuc}$ are not supported by recent {\sc kepler} models.
+
+% describe plausible assumptions for ration of $\xi_b/\xi_p$.
+The system inclination for LMXBs is notoriously difficult to measure.
+For non-dipping sources, the likely range is up to $72^\circ$, while dipping sources are likely $i\gtrsim79.3^\circ$ \cite[]{gal16a}.
+% 
+% 5, 16, 50, 68, 95 th percentiles:
+% [ 0.58815149  0.72522467  1.12987208  1.49149624  1.59287602]
+% min, max
+% 0.525409532417 1.63715782279
+For an isotropically-distributed population of non-dipping sources, the ratio $\xi_b/\xi_p$ is roughly uniformly-distributed between 0.53--1.64, rising a little towards the high end.
+The median value is 1.13, while the 95\% confidence interval is 0.73--1.49.
+
+\begin{itemize}
+\item describe plausible assumptions for $z$, e.g. based on recent analysis of bursts
+\end{itemize}
 
@@
 
-Given a minimally complete set of burst observations as described in \S\ref{subsec:data}, a common approach \cite[e.g][]{falanga11} is to estimate the fuel composition at ignition, and hence the accreted fuel composition, based on simple analytic estimates of the burst energy production $Q_{\rm nuc}$.
-
-One issue with this approach (as for analyses based on one burst, as described in \S\ref{subsec:1burst}) is that the effects of anisotropy are frequently neglected.
-%
-A second issue which has emerged recently is that the commonly used estimates for the nuclear energy generation rate $Q_{\rm nuc}$ are not supported by recent {\sc kepler} models. 
-%
-We describe the general approach below including the effects of these two issues.
-
-% describe plausible assumptions for ration of $\xi_b/\xi_p$.
-The system inclination for LMXBs is notoriously difficult to measure.
-For non-dipping sources, the likely range is up to $72^\circ$, while dipping sources are likely $i\gtrsim79.3^\circ$ \cite[]{gal16a}.
-% 
-% 5, 16, 50, 68, 95 th percentiles:
-% [ 0.58815149  0.72522467  1.12987208  1.49149624  1.59287602]
-% min, max
-% 0.525409532417 1.63715782279
-For an isotropically-distributed population of non-dipping sources, the ratio $\xi_b/\xi_p$ is roughly uniformly-distributed between 0.53--1.64, rising a little towards the high end.
-The median value is 1.13, while the 95\% confidence interval is 0.73--1.49.
-
-\begin{itemize}
-\item describe plausible assumptions for $z$, e.g. based on recent analysis of bursts
-\end{itemize}
-
-Thus, by substituting in equation \ref{qnuc} we can estimate 
-\begin{equation}
-\langle X\rangle 
-% In [42]: c.c**2/(6.0455*u.MeV/c.m_p).to("m**2/s**2")                             
-% Out[41]: <Quantity 155.20173377>
-%% = \frac{c^2z/\alpha}{6.6\ {\rm Mev/nucleon}}\frac{\xi_b}{\xi_p} - 0.145
-% = z\frac{142}{\alpha}\frac{\xi_b}{\xi_p} - 0.145 \label{xbar}
- = z\frac{155}{\alpha}\frac{\xi_b}{\xi_p} - 0.223 \label{xbar}
-\end{equation}
-[{\it comment on the degree of error introduced by ignoring anisotropy etc.}] 
-%
-Clearly this expression is only applicable for $\alpha$-values up to some limit, which we calculate as
-\begin{equation}
-%In [43]: c.c**2/(1.*u.MeV/c.m_p).to("m**2/s**2")/1.3455        
-% Out[43]: <Quantity 697.34082607>
-% \alpha \leq 980\, z \frac{\xi_b}{\xi_p}
-\alpha \leq 697\, z \frac{\xi_b}{\xi_p}
-\end{equation}
-Interestingly, this limit is 
-% $\approx70$\% 
-%  c.c**2/(1.*u.MeV/c.m_p).to("m**2/s**2")/1.6                 
-% Out[44]: <Quantity 586.42005092>
-% In [45]: 697./586.                                                               
-% Out[45]: 1.189
-$\approx20$\%
-%
-larger than the equivalent value derived for the old expression for $Q_{\rm nuc}$.
-This result suggests that larger observed values of $\alpha$ may be accommodated without resorting to explanations including incomplete burning of burst fuel \cite[e.g.][]{bcatalog}.
-
-% upper limit on xbar
-Conversely, for low values of $\alpha$, we may apply the constraint that $\langle X\rangle \lesssim 0.77$, corresponding to the expected maximum possible for accreted fuel with primordial abundances.
-
-The hydrogen fraction at ignition $\langle X\rangle$ and the burst recurrence time may then be used to estimate the hydrogen fraction in the fuel, $X_0$.
-The H-fraction at the base of the fuel layer is reduced steadily by $\beta$-limited hot CNO burning, and will completely exhaust the accreted hydrogen in a time \cite[]{lampe16}:
-\begin{equation}
-t_{\rm CNO} =9.8\left(\frac{X_0}{0.7}\right) \left(\frac{Z_{\rm CNO}}{0.02}\right)^{-1}
-\end{equation}
-measured in the neutron star frame.
-% updated 2019 Aug 26 with exhaustion case
-The hydrogen fraction at the base will thus be $X_0[1-\Delta t/(1+z)t_{\rm CNO}]$, 
-% [{\it should this expression include time dilation?}] -- yes
-%
-provided that $\Delta t < (1+z)t_{\rm CNO}$.
-Once $\Delta t$ exceeds the time to burn all the hydrogen at the base, the abundance there will be zero, and a growing layer of pure He fuel will develop.
-The average H-fraction in the layer for these two cases will be
-%so that the average H-fraction in the layer will be 
-\begin{equation}
-\langle X \rangle = \begin{cases}
-%  X_0\left[1-0.5\frac{\Delta t}{(1+z)t_{\rm CNO}}\right] & \Delta t \leq (1+z)t_{\rm CNO} \\
-%  0.5X_0\frac{(1+z)t_{\rm CNO}}{\Delta t} & \Delta t > (1+z)t_{\rm CNO}
-  X_0(1-0.5f_{\rm burn}) & \Delta t \leq (1+z)t_{\rm CNO} \\
-  0.5X_0/f_{\rm burn} & \Delta t > (1+z)t_{\rm CNO}
-  \end{cases}
-\end{equation}
-where $f_{\rm burn} = \frac{\Delta t}{(1+z)t_{\rm CNO}}$ is the ratio of the recurrence time to the time to burn all the H.
-%
-We can combine this expression with equation \ref{xbar} to give
-\begin{equation}
-X_0 = \begin{cases}
-%        & \langle X\rangle +0.35\left[\frac{\Delta t}{(1+z)9.8\ {\rm hr}}\right]
-%                            \left(\frac{Z_{\rm CNO}}{0.02}\right) \nonumber \\
-         z\frac{142}{\alpha}\frac{\xi_b}{\xi_p} - 0.145 
-                           +\left[\frac{\Delta t}{(1+z)28\,{\rm hr}}\right]
-                       \left(\frac{Z_{\rm CNO}}{0.02}\right) & f_{\rm burn} \leq 1\\
-         \sqrt{ \frac{\Delta t}{(1+z)7\,{\rm hr}}\frac{Z_{\rm CNO}}{0.02}
-                        \left(z\frac{142}{\alpha}\frac{\xi_b}{\xi_p} - 0.145\right) }
-                        & f_{\rm burn} > 1
-        \end{cases}
-    \label{eq:X_0}
-\end{equation}
-Practically, the issue with these expressions is that calculating $f_{\rm burn}$ requires knowledge of $X_0$, which is the unknown we are trying to constrain.
-One approach is to adopt a trial value of $X_0$, calculate $f_{\rm burn}$ and hence an updated estimate of $X_0$ via equation \ref{x0}, and iterate until no further change in the estimate arises.
-
-Clearly this estimate of $X_0$ will incorporate systematic errors arising from the assumed values of $z$, the inclination, and $Z_{\rm CNO}$. We explore the typical range of these errors in 
-% the next section.
-\S\ref{sec:simulations}
-
+We describe the general approach below including the effects of these two issues (anisotropy and qnuc).
+
+% original text here moved to the earlier "methods" section
 \begin{itemize}

### 2021-09-20 concord/utils.py

@@ -433,7 +433,7 @@ def Q_nuc(Xbar, quadratic=False, old_relation=False, coeff=False):
 
     q = [1.3050, 6.9511, -1.9218]
 
-    if quadratic:
+    if not quadratic:
         q = [1.3455, 6.0455]
 
     if old_relation:

### 2021-09-23 concord/utils.py
Updated hfrac to provide better flexibility, consistency with other functions
Added tdel_dist function to estimate PDF of recurrence time given burst count, exposure

@@ -2,9 +2,13 @@
 # Augmented 2019 Aug with routines from Inferring\ composition.ipynb
 #
 # def value_to_dist(_num, nsamp=NSAMP_DEF, unit=None):
+# def homogenize_params(theta, nsamp=None):
 # def len_dist(d):
+# def asym_norm(m, sigm=None, sigp=None, nsamp=NSAMP_DEF, positive=False, model=1):
+# def intvl_to_errors(perc):
 # def g(M,R,Newt=False,units='cm/s^2'):
 # def redshift(M,R):
+# def check_M_R_opz(M, R, opz):
 # def calc_mr(g,opz):
 # def solve_radius(M,R_Newt,eta=1e-6):
 # def decode_LaTeX(string):
@@ -14,14 +18,15 @@
 # def hfrac(alpha, tdel, opz=1.259, zcno=0.02, old_relation=False, ...)
 # def iso_dist(nsamp=1000, imin=0., imax=IMAX_NDIP)
 # def dist(F_pk, F_pk_err, nsamp=10000, X=0.0, empirical=False, ...)
+# def luminosity(F_X, F_X_err=0.0, dist=8*u.kpc, dist_err=None, nsamp=NSAMP_DEF, ...)
 # def mdot(F_per, F_per_err, dist, nsamp=10000, M=M_NS, R=R_NS, ...)
 # def yign(E_b, E_b_err, dist, R_NS=10., Xbar=0.7, opz=1.259, ...)
-# def luminosity(F_X, F_X_err=0.0, dist=8*u.kpc, dist_err=None, nsamp=NSAMP_DEF, ...)
 # def L_Edd(F_pk, F_pk_err, dist, dist_err=0.0, nsamp=10000, dip=False, ...)
 
 import astropy.units as u
 import astropy.uncertainty as unc
 from astropy.visualization import quantity_support
+from scipy.stats import poisson
 
 # Some defaults
 
@@ -411,6 +416,40 @@ def decode_LaTeX(string, delim='pm'):
 
 # ------- --------- --------- --------- --------- --------- --------- ---------
 
+def tdel_dist(nburst, exp, nsamp=NSAMP_DEF):
+    """
+    Function to generate a synthetic distribution of recurrence times, given
+    nburst observed events over a total exposure of exp. The resulting
+    distribution may approximate the PDF of the underlying burst rate, but
+    does assume that the bursts are independent, which is obviously not the
+    case
+    :param nburst: number of events detected
+    :param exp: total exposure time
[…]
